### 2024-03-25 outputs/paper/paper.Rmd
qmd and bib file structured

---
title: 
subtitle: 
author: "Ziheng Zhong"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: pdf_document
abstract: TBD
thanks: 'Code and data are available at: https://github.com/iJustinn/House_Price.git'
toc: no
bibliography: references.bib
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(kableExtra)
library(countrycode)
library(chron)
library(ds4psy)
library(scales)
library(DiagrammeR)
library(knitr)
library(tinytex)
library(haven)
library(ggplot2)

masterdata <- read_csv("masterdata.csv")
masterdata_clean <- read.csv("masterdata_clean.csv")

```

# Introduction

Wordle is an online word-guessing game bearing resemblance to historically popular word puzzles such as the New York Times crossword puzzle and the board game Mastermind. Players are given a five-letter word each day and are asked to guess the word within six tries. If the word is not guessed in earlier attempts, the letters for your word will become either grey, yellow, or green: grey indicating that the letter was not found anywhere in the word, yellow indicating the presence of the letter but in the wrong position, and green indicating both letter and position are correct. The goal is to guess five green letters and therefore the correct word. The game was created by Josh Wardle, a software engineer who created the game for himself and his partner to pass the time during the 2021 COVID-19 pandemic. The meteoric rise in popularity of his creation has been attributed to a feature where players can easily copy and paste their results to share with their friends and family. Wardle's game found international spread in late December of 2021 thanks to multiple factors, including the ease in sharing results and the daily nature of the game, as well as its simplicity. Wordle's popularity can be seen to its fullest extent on Twitter, where millions of users share their results every day by pasting their scores via the website's built-in feature.

This paper aims to analyze several metrics of Wordle's popularity on Twitter. As a result of millions of users sharing their scores each day, analysis can and has been done on variables including average score- a recent study found that Sweden was the country with the lowest average guesses with 3.72 average guesses out of a maximum of 6 [@citewordtips]. Here, some key findings we wish to seek include attempting to model whether certain characteristics of a word correlate with its average difficulty, and the average score of Twitter users in relation to their geographical location, such as their country of residence. These metrics were not discussed deeply in prior papers and may be of interest to future puzzle makers, linguists, and game theorists.

Using the Twitter API, tweets pertaining to Wordle were gathered and aggregated into multiple datasets used for analysis. These tweets were gathered during intervals from April 4-10, 2022, and April 17-23, 2022). User scores were collected by searching for common templates that are used by the website domain to share scores in their tweets. Along with the score the used reached, the time of posting, geographical location, and other data provided by Twitter were also obtained. The results were turned into graphical data and analyzed according to our key missions.

In the Data section, we go over the data gathered from the Twitter API and how it was cleaned and arranged for proper statistical analysis. We discuss variables used, methodology, and graphs showing distributions of data. In the Model section, we discuss our model and its implications for how we interpret our results going forward. In the Results section, graphical data is shown and used to present our story according to the key goals and findings we discover during our analysis. Lastly, we discuss what has been done, its significance, and weaknesses of the paper in our Discussion section.

# Data

## Dataset

To accomplish the goals set out in this paper, the data used consisted of a sample of tweets from the social media website Twitter. Twitter hosts and stores an immense number of messages which can be accessed via the website's API; however, access to tweet data is restricted to the past 7 days during access unless given permission by the site. The raw data collected in this paper contains 504000 tweets and 90 variables regarding various properties of the tweet and its sender, such as the message, date and time of posting, details of the sender's account, popularity metrics, and direct links to the tweet online. Tweets were searched for and collected over two timeframes: tweets from April 4-10, and from April 17-23. This was done using the programming software R [@citeR] and the tweet collecting package rtweet [@citertweet]. Complete details as to how the data was obtained and cleaned can be found in the datasheet in the appendix.

## Variables

The dataset consisted of 90 variables which needed to be trimmed down to a manageable list of relevant data. These variables included unique tweet identifiers and the exact text included, details regarding whether the tweet was original or sent as a reply or response to another, the number of likes and retweets on the initial or response posts, the device used for sending posts, and links to the posts and accounts of each observation. We seeked to only utilize the score the user had achieved in the game, their country/location, the date of the message, and the presence of 'hard mode' while playing, which adds the condition that you must use hints you are given from prior guesses. To accomplish this, variables needed to be changed and manipulated. From the raw data, which used Unix time to store chronological data, the tidyverse [@citetidyverse] and lubridate [@citelubridate] packages were used to create new variables. From the Unix time, the date was extracted from the raw data and stored in a YYYY-MM-DD format. An indicator variable corresponding to whether the user used hard mode was created and each observation was given a classification based on the contents of their tweet. The final dataframe contained 4 variables, those variables being the date, country, score, and gamemode for each tweet instance provided.

```{r preview, tab.cap = "Sample of Cleaned Twitter Dataset", tab.width=5, echo=FALSE}
masterdata_clean %>%
  select(date_tweeted, country, score, hardmode) %>%
  arrange(date_tweeted) %>%
  slice(1:10) %>%
  kable(
    caption = "First ten rows of a dataset showing Wordle-related tweets",
    col.names = c("Date", "Country", "Wordle Score", "Hardmode"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  ) %>%
  kable_styling(latex_options = c("striped", "scale_down"))

```
## Missing Data

There are limitations to the data gathered in this paper which will have impacts on the conclusions we make going forward. Due to regulations of Twitter and limitations of the API provided, historical data was not available for analysis, since the website only allows for mass collection of tweets from a week in the past. The paper regrets that long-term historical analysis was not possible to conduct. As well, chronological data was unavailable to be processed since the rtweet package can only collect recent tweets in bulk. The result of this was that data was unable to be gathered consistently over a long time interval due to hardware limitations and feasibility. While the data should not be negatively impacted by this, it becomes impossible to properly conduct chronological analysis, and there remains the possibility of time interval bias from a majority of the observations being taken at one timeframe.

There is also inherent bias included in the dataset as a result of the nature of tweets. Not all players of Wordle report their results online, and a smaller subset on Twitter, leaving the possibility that the conclusions drawn do not apply to the general population of all Wordle players. At best, we can confirm our conclusions from a sample of Twitter users and attempt to extrapolate to a far larger population. As Wordle is a word based off the English language, it is likely that the geographical data will also be biased in favor of English-speaking countries, as countries that are not primarily English-speaking likely have far less Wordle players, which has a magnified effect considering the nature of the dataset. Further bias exists where users who experience lower scores (higher average guesses to guess the word) or fail to complete the game may not post about these results on Twitter. All data collected is by necessity self-reported and lends itself to nonresponse bias. Users may also lie or provide false data by changing their reported score or by parodying the original Wordle message, which may affect the shape of our distribution.

## Plots

Plots were created corresponding to the chosen variables in our cleaned dataset in order to familiarize with the obtained data and provide motivation for further analysis. To begin with, we wish to plot the frequency of each result/score across our entire dataset. We compare the frequency of each result and plot it as both the raw frequency and the percentage that each score appears in our dataset. Note that failures, where the user was unable to guess the word in the maximum 6 tries, is plotted as a score of 7. We observe that a majority of Twitter users guess complete the game within 4 tries, and that the distribution of guesses is sensibly unimodal. The distribution is slightly right-skewed, and the dataset shows that the number of failures is over three times as much as the number of people who guess the word on their first try, which is expected due to the low probability of guessing the exact word with no prior information compared to the chance of failure. This is also reflected in the proportions between scores of 2 and 6. 
```{r freq, fig.cap = "Scores of Twitter Wordle Players", echo = FALSE}
masterdata_clean %>%
  ggplot(mapping = aes(x = score)) + geom_bar(fill = '#6aaa64') + xlab("Score") + ylab("Frequency") +
  geom_text(aes(label = ..count..), vjust = -0.5, stat = "count")

# https://stackoverflow.com/questions/40249943/adding-percentage-labels-to-a-bar-chart-in-ggplot2
```

```{r perc, fig.cap = "Proportions of Scores of Twitter Wordle Players", echo = FALSE}
masterdata_clean %>% 
    count(Score = factor(score)) %>% 
    mutate(pct = prop.table(n)) %>% 
    ggplot(aes(x = Score, y = pct, label = scales::percent(pct))) + 
    geom_col(position = 'dodge', fill = '#6aaa64') + 
    geom_text(position = position_dodge(width = .9),
              vjust = -0.5,
              size = 3) + 
    scale_y_continuous(labels = scales::percent)

```
Next, we look to compare the distributions between users who use hard mode and users who do not. It must be noted that the amount of people who do not enable hard mode dwarf the number of people who do. However, it is possible that a non-insignificant majority of people who do not report using hard mode are self-imposing hard mode conditions onto themselves, whether intentionally or unintentionally. Regardless, we plot the proportion of hard mode users to normal users and see that 
94% of users in our sample do not use hardmode, which can be explained by the fact that hard mode is opt-in and users may not wish to bother with enabling it for multiple reasons. We then repeat the distribution of scores in the previous set of graphs but separate the results by the mode that users play on. From our graphs, we see that our sample of players on hard mode more commonly reach scores of 3 and 4, but also have higher failure rate than normal players. A strategy used by normal players is to disregard hints in the first 2 or 3 words by trying to 'use' as many common letters as possible to reveal as many letters in the word as possible before making a realistic guess, while hard mode players must use the hints they are given which reduces the efficiency of the strategy but means that each word they use has a relatively higher chance of being correct. However, this runs into problems when it is difficult to use the hints provided or the word shares a structure with other common words (for example, Wordle #284 was 'stove': a player could easily waste guesses on 'stone', 'store', 'stoke', etc), which results in a higher failure rate.  
```{r mode, fig.cap = "Proportions of Gamemodes of Twitter Wordle Players", echo = FALSE}
masterdata_clean %>% 
    count(count = factor(hardmode)) %>% 
    mutate(pct = prop.table(n)) %>% 
    ggplot(aes(x = count, y = pct, label = scales::percent(pct))) + 
    geom_col(position = 'dodge', fill = '#6aaa64') + 
    geom_text(position = position_dodge(width = .9),
              vjust = -0.5,
              size = 3) + 
    scale_y_continuous(labels = scales::percent) + xlab('Proportion of normal to hardmode players') +
  ylab('Percent')
```

```{r percgamemode, fig.cap = "Proportions of Scores of Twitter Wordle Players by Gamemode", echo = FALSE}
# Distributions

masterdata_clean %>%
  filter(hardmode == 'Off') %>%
    count(Score = factor(score)) %>% 
    mutate(pct = prop.table(n)) %>% 
    ggplot(aes(x = Score, y = pct, label = scales::percent(pct))) + 
    geom_col(position = 'dodge', fill = '#6aaa64') + 
    geom_text(position = position_dodge(width = .9),
              vjust = -0.5,
              size = 3) + 
    scale_y_continuous(labels = scales::percent) + xlab('Score (Normal)') +
  ylab('Percent')

masterdata_clean %>%
  filter(hardmode == 'On') %>%
    count(Score = factor(score)) %>% 
    mutate(pct = prop.table(n)) %>% 
    ggplot(aes(x = Score, y = pct, label = scales::percent(pct))) + 
    geom_col(position = 'dodge', fill = '#6aaa64') + 
    geom_text(position = position_dodge(width = .9),
              vjust = -0.5,
              size = 3) + 
    scale_y_continuous(labels = scales::percent) + xlab('Score (Hardmode)') +
  ylab('Percent')

```

We also show the frequency that countries occur in our dataset, using the geographical data provided in the tweets. We display the top 10 countries found in our dataset and observe that the Americas and the UK make up a majority of the tweets found in the data. 
```{r country, fig.cap = "Top 10 Frequent Wordle Countries", echo = FALSE}
countryfreq <- table(masterdata_clean$country)
countryfreq %>%
  as.data.frame() %>%
  arrange(desc(Freq)) %>%
  slice(1:10) %>%
  kable(
    caption = "Top 10 most common countries found in dataset",
    col.names = c("Country", "Frequency"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  ) %>% kable_styling(latex_options = c("striped", "scale_down"))
```
To conduct further analysis, we seek to plot data based on the average score of users on different dates and different countries. Shown is a table displaying the date of the daily word being given, the average score of Twitter users across our dataset, and the word of the day. Also plotted is the distribution of the 'easiest' and 'hardest' words determined by average score, displayed in percentages and grouped by the mode used.
```{r wordinfo, fig.cap = "Score and Word by Date", echo = FALSE}

average_score_df <- masterdata_clean %>%
  group_by(date_tweeted) %>%
  transmute(average_score = mean(as.numeric(score))) %>%
  unique() %>%
  arrange(date_tweeted, desc(date_tweeted))
average_score_df <- average_score_df[-c(8, 9, 17, 18), ]

average_score_df$daily_word <- c("shawl", "natal", "comma", "foray", "scare", "stair", "black",
                                 "ample", "flair", "foyer", "cargo", "oxide", "plant", "olive")

#average_score_df$daily_word %>%
  #count_chars()

average_score_df %>% 
  kable(
    caption = "Average score of Twitter Wordle players",
    col.names = c("Date", "Average Score", "Daily Word"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  ) %>% 
  kable_styling(latex_options = c("striped", "scale_down"))

```

```{r somestats, fig.cap = "Examples of Wordle Scores Grouped By Gamemode", echo = FALSE}

masterdata_clean %>%
    filter(date_tweeted == "2022-04-09") %>%
    count(Score = factor(score)) %>% 
    mutate(pct = prop.table(n)) %>% 
    ggplot(aes(x = Score, y = pct, label = scales::percent(pct))) + 
    geom_col(position = 'dodge', fill = '#6aaa64') + 
    geom_text(position = position_dodge(width = .9),
              vjust = -0.5,
              size = 3) + 
    scale_y_continuous(labels = scales::percent)

masterdata_clean %>%
    filter(date_tweeted == "2022-04-19") %>%
    count(Score = factor(score)) %>% 
    mutate(pct = prop.table(n)) %>% 
    ggplot(aes(x = Score, y = pct, label = scales::percent(pct))) + 
    geom_col(position = 'dodge', fill = '#6aaa64') + 
    geom_text(position = position_dodge(width = .9),
              vjust = -0.5,
              size = 3) + 
    scale_y_continuous(labels = scales::percent)

# Hardmode

masterdata_clean %>%
    group_by(hardmode) %>%
    filter(date_tweeted == "2022-04-09") %>%
    count(Score = factor(score)) %>% 
    mutate(pct = prop.table(n)) %>% 
    ggplot(aes(x = Score, y = pct, fill = hardmode, label = scales::percent(pct, 0.1))) + 
    geom_col(position = 'dodge') + 
    geom_text(position = position_dodge(width = 1),
              vjust = -0.5,
              size = 3) + 
    scale_y_continuous(labels = scales::percent) + xlab('Score (April 9 "Chair")') +
  ylab('Percent') + scale_fill_manual(values=c("#6aaa64", "#c9b458"))

masterdata_clean %>%
    group_by(hardmode) %>%
    filter(date_tweeted == "2022-04-19") %>%
    count(Score = factor(score)) %>% 
    mutate(pct = prop.table(n)) %>% 
    ggplot(aes(x = Score, y = pct, fill = hardmode, label = scales::percent(pct, 0.1))) + 
    geom_col(position = 'dodge') + 
    geom_text(position = position_dodge(width = 1),
              vjust = -0.5,
              size = 3) + 
    scale_y_continuous(labels = scales::percent) + xlab('Score (April 19 "Foyer")') +
  ylab('Percent') + scale_fill_manual(values=c("#6aaa64", "#c9b458"))

```

Finally, we conduct analysis on the words used in the game themselves. The following table shows the date the word was used, the average score users scored on it, the word of the day, the frequency of the word in texts according to the Corpus of Contemporary American English (COCA) [@citecoca], and the number of orthographic neighbors to the word according to MCWord [@citemcword]. Orthographic neighbors are defined as words which are the same length but differ by one letter (for example, scare and stare).

```{r ling, fig.cap = "Linguistic Properties of Wordle Words", echo = FALSE}
average_score_df$rank <- c("11938", "26852", "11931", "10467", "3837", "2880", "253",
                                 "6374", "11502", "9811", "4953", "10008", "623", "6242")
average_score_df$neighbournum <- c(2, 3, 1, 1, 10, 2, 5, 3, 1, 2, 1, 0, 5, 1)

average_score_df %>% 
  kable(
    caption = "Linguistic Properties of Daily Wordle Words",
    col.names = c("Date", "Average Score", "Daily Word", "Word Frequency Rank", "Orthographic Neighbours"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  ) %>% 
  kable_styling(latex_options = c("striped", "scale_down"))

```

# Model

In this paper, we seek to use regression models to determine if there were relationships between the average score that Twitter users scored daily and properties of the given word for the day. We begin by constructing directed acyclic graphs to visualize the variables we wish to discuss and model, and clearly show the relationships we believe exist between them. Using the DAG as a visual, the paper aims to show that the frequency of the word in literature, the number of orthographic neighbours, and the game mode of the player all have significance in predicting the average score of a player.

```{r, echo=FALSE}
# Thanks to https://www.tellingstorieswithdata.com/causality.html#causality .

DiagrammeR::grViz("
digraph {
  graph [ranksep = 0.2]
  node [shape = plaintext, fontsize = 10, fontname = Helvetica]
    'Average score'
    'Frequency of word'
    '# Orth neighbours'
    'Game mode'
  edge [minlen = 2, arrowhead = vee]
    'Game mode'->'Average score'
    'Frequency of word'->'Average score'
    '# Orth neighbours'->'Average score'
{ rank = same;     'Frequency of word';
    '# Orth neighbours';
    'Game mode'; }
}
", height = 200)

DiagrammeR::grViz("
digraph {
  graph [ranksep = 0.2]
  node [shape = plaintext, fontsize = 10, fontname = Helvetica]
    'Failure rate'
    'Frequency of word'
    '# Orth neighbours'
    'Game mode'
  edge [minlen = 2, arrowhead = vee]
    'Game mode'->'Failure rate'
    'Frequency of word'->'Failure rate'
    '# Orth neighbours'->'Failure rate'
{ rank = same;     'Frequency of word';
    '# Orth neighbours';
    'Game mode'; }
}
", height = 200)
```
The paper predicts that the linguistic properties of words such as the number of orthographic neighbors will have a different effect on the average score of a player based on their game mode; players on hard mode are hypothesized to have higher (worse) scores when the daily word shares similar structures to other possible guesses. In order to attempt to model this effect, we create multiple regression models with 2 predictors each as follows:

$$Y_{1} = \beta_{0} + \beta_{1}X_{1} + \beta_{2}X_{2}  + \epsilon$$

In our first multiple regression model, $Y_{1}$ represents the average number of guesses it takes a normal player to complete a given Wordle puzzle. $X_{1}$ represents the frequency rank of a given word used in the game, $X_{2}$ represents the number of orthographic neighbors that the word has, $\beta_{0}$ represents the intercept or average score when the frequency rank and number of neighbors is 0, and $\beta_{1}$ and $\beta_{2}$ represent the regression coefficients. For every one increase in frequency rank and every one increased neighbour, $Y_{1}$ increases by $\beta_{2}$ and $\beta_{2}$ respectively.

$$Y_{2} = \beta_{0} + \beta_{1}X_{1} + \beta_{2}X_{2}  + \epsilon$$

The second multiple regression model is similar. $Y_{2}$ represents the average number of guesses it takes a player on hard mode to complete a given puzzle and the other predictors and coefficients share the same meaning. We then wish to model the probability of failure to complete the game based on the same groupings of game mode and linguistic properties. The models are as follows:

$$Y_{3} = {Pr}(y_{i}=1)={logit}^{-1}(\beta_{0H} + \beta_{1H}X_{1H} + \beta_{2H}X_{2H})$$

$$Y_{4} = {Pr}(y_{i}=1)={logit}^{-1}(\beta_{0H} + \beta_{1H}X_{1H} + \beta_{2H}X_{2H})$$

$Y_{3}$ and $Y_{4}$ respectively represent the average probability that a normal Wordle player and a Wordle player on hard mode will fail in guessing the daily word within 6 guesses. The other predictors and coefficients share the same meaning as the prior models.

```{r, echo=FALSE, include=FALSE}
masterdata_clean_modelmain <- cbind(masterdata$text, masterdata_clean)%>%
  mutate(text = masterdata$text)%>%
  subset(select = -c(1))%>%
  mutate(text = sub(".*Wordle ", "", text ))%>%
  mutate(text = sub(" .*", "", text))

# Thanks to https://stackoverflow.com/questions/25276155/remove-the-rows-that-have-non-numeric-characters-in-one-column-in-r

masterdata_clean_modelmain <- masterdata_clean_modelmain[!is.na(as.numeric(as.character(masterdata_clean_modelmain$text))),]

masterdata_clean_modelmain <- masterdata_clean_modelmain%>%
  filter(between(text, 289,310) == TRUE)

masterdata_clean_modelmain <- masterdata_clean_modelmain %>%
  mutate(rank = ifelse(
    text==289, 11938, ifelse(
    text==290, 26852, ifelse(
    text==291, 11931, ifelse(
    text==292, 10467, ifelse(
    text==293, 3837,  ifelse(
    text==294, 2880,  ifelse(
    text==295, 253,   ifelse(
    text==302, 6374,  ifelse( 
    text==303, 11502, ifelse(
    text==304, 9811,  ifelse(
    text==305, 4953,  ifelse(
    text==306, 10008, ifelse(
    text==307, 623, 6242
 ))))))))))))))

masterdata_clean_modelmain <- masterdata_clean_modelmain %>%
  mutate(neighbournum = ifelse(
    text==289, 2, ifelse(
    text==290, 3, ifelse(
    text==291, 1, ifelse(
    text==292, 1, ifelse(
    text==293, 10,ifelse(
    text==294, 2, ifelse(
    text==295, 5, ifelse(
    text==302, 3, ifelse( 
    text==303, 1, ifelse(
    text==304, 2, ifelse(
    text==305, 1, ifelse(
    text==306, 0, ifelse(
    text==307, 5, 1
 ))))))))))))))

masterdata_clean_modelmain_norm <-  masterdata_clean_modelmain%>%
  filter(hardmode == "Off")
masterdata_clean_modelmain_hard <-  masterdata_clean_modelmain%>%
  filter(hardmode == "On")

model1 <- lm(score ~ rank + neighbournum, data = masterdata_clean_modelmain_norm)

model2 <- lm(score ~ rank + neighbournum, data = masterdata_clean_modelmain_hard)

masterdata_clean_modelmain_norm <- masterdata_clean_modelmain_norm%>%
  mutate(failed = (score == 7))
masterdata_clean_modelmain_hard <- masterdata_clean_modelmain_hard%>%
  mutate(failed = (score == 7))

model3 <- lm(failed ~ rank + neighbournum, data = masterdata_clean_modelmain_norm)
model4 <- lm(failed ~ rank + neighbournum, data = masterdata_clean_modelmain_hard)

```
## Features

We are interested in the linguistic properties of words and how they may translate to the difficulty of guessing it in the context of Wordle, so the variables we decide to include for our model include strictly predictors related to our goal. The frequency rank of a word directly correlates to its use in literature, speech and academia as reported by the COCA, and is used here to represent how common a word is and how likely one is to have encountered it in their lives. It is believed that more familiar words will be easier to recall and guess, and therefore have a lower score, while less common words will take longer to guess. We also look at the number of orthographical neighbors as a predictor: structurally similar words can serve to 'remind' players of a word's presence and lead them to the correct guess, but they can also confuse and discourage players who continually guess the wrong neighbor. The lack of neighbors can also assist the player, since a lack of structural similarity to other words is representative of 'uniqueness', and therefore cannot be confused for other words: for example, if one is asked to fill in '_nique', it becomes apparent that only 'unique' completes the word, and in general unique has 0 neighbors and is entirely distinct in structure. This may serve to avoid confusion in players and reduce the amount of guesses required to win the game.

## Model Concerns

The most concerning element of modelling these relationships is the sample size of the amount of words in our database. Due to Twitter limitations and constraints, a very small number of words were collected which is not nearly enough to form an accurate model of what we wish to predict. The number of data points per day is suitable and reduces variance for each word, but the small number of words means that each day shows great significance in the data and does not lead to an accurate conclusion. The model would benefit from a dataset spanning a much longer timerange in order for more words to be analyzed and fit into the model.

# Results

Two variations of two models were created and analyzed for a total of four completed models, and interpreted by analyzing the values of the predictor coefficients and their implications on the overall model and the value which they served to predict. We continue by sharing the results of each type of model and how their findings can be compared to each other.

## Modelling Average Score

The first form of model was created to predict the average score of a Wordle player based on the frequency of the word and the number of orthographic neighbours the word had. Our results show that if a player is playing normally, i.e. not using hard mode, their average number of guesses will increase by 0.00004126 for each rank the word is from 1. This corresponds to the value of $Y_{1}$ = 0.00004126 in our model equation. For example, a word around the ranking 10000 would increase the average number of guesses by 0.4126. Furthermore, the value of $Y_{2}$ was determined to be -0.04117. This can be interpreted as: For every orthographical neighbour the word of the day has, the average number of guesses it takes a normal Wordle player will be reduced by 0.04117 guesses. Lastly, the intercept was found to be $\beta_{0}$ = 3.954. The interpretation of this result is that a word at ranking '0' and with 0 neighbours will be guessed in an average of 3.954 tries; it represents the average number of guesses when the other two predictors are held constant.

```{r, echo=FALSE}
summary(model1)
```
The second model used the same predictor variables but used a dataset consisting of only players using Wordle's hard mode. This means that the interpretations of the variables remain the same, but they have the effect of changing the average number of guesses for a hard mode player and not a normal player. This has an effect on the values of the coefficients, which are reflected in the model. In the secondary model, the intercept has a value of 3.843, the ranking coefficient has a value of 0.00004232, and the neighbour coefficient has a value of -0.04185. In general, there is not a drastic change in the values of the coefficients between each model. The second model features a lower intercept, implying that hard mode players finish the game slightly faster, but their results are very slightly more impacted by the frequency of the word each day.

```{r, echo=FALSE}
summary(model2)
```
## Modelling failure rate

The second form of model was created to predict the failure rate of players across both game modes using the same metrics as the first model. The results of these models have similar structure due to the simplicity of the models and the small number of predictor variables, but the values and their interpretations are different and reveal separate results from the previous models. The third model, which predicts failure rate of a normal Wordle player, shows that the intercept has a value of $\beta_{0}$ = 0.02647. Because this model is predicting a percentage, the interpretation of the intercept shows that if a word has 0 rank and 0 orthographical neighbours, a normal Wordle player will have a 2.647% chance of failure on average according to the model. The value of the rank coefficient is $Y_{1}$ = 0.000001314. For every rank a word has, the percentage chance that a normal will fail to guess the word within 6 tries increases by 0.0001314. If a word had a rank of 10000, then the chance that a player would fail increases by 1.314 from the intercept value. The neighbour coefficient is recorded as being $Y_{2}$ = -0.0004534, and shows that for every neighbour a word has, the chances of failure drop by 0.04534%.

```{r, echo=FALSE}
summary(model3)
```
The final model which predicts the failure rate of hard mode players shows significantly different results than the previous model predicting failure rates of normal players. We observe an intercept of 0.0301, a rank coefficient of 0.000001598, and a neighbour coefficient of -0.00134. A table shows the two comparisons side by side along with the difference by value and percentage.
```{r, echo=FALSE}
summary(model4)
```

# Discussion

## Data and Model Findings

In the time of writing of this paper, numerous reports and articles have been written about various aspects of Wordle, thanks to its meteoric rise in popularity since late 2021 as a result of the COVID-19 pandemic. This paper attempts to go a step further into analysis and cover the linguistic aspects of the popular word game and if any significant conclusions can be drawn from the results of the author's findings. The models constructed in this paper lead to both conclusions that are both reflected in the data and which are surprising at first glance but can be understood. It was the paper's hypothesis that the number of orthographic neighbours a word has would negatively impact the performance of the player; the failure rate and the number of guesses taken would both increase the similar a word was to others. This was concluded from the author's own experiences and the thought that words sharing high similarity would lead to many incorrect guesses used on the neighbours. However, the four models constructed show that while the less common a word is, the harder it is to guess, as predicted, the number of neighbours has a negative correlation with the average number of guesses as well as the failure rate of both types of players. Some explanations could include that having more neighbours means that the word's structure is easier to recall and thus provides the player with an early lead if guessed, or that the common structure of the word is correlated with words that players like to begin with as their first or second words, and therefore are easier to obtain hints on at the beginning of the game.

The data had shown that on average, players on hard mode have lower average guesses than normal players but exhibit a higher failure rate overall. This is reflected in the paper's models- while the model of the hard mode dataset showed a lower intercept value in the first type of model, there was a significant difference in the failure rates of the two types of players as reflected in the intercepts of the third and fourth model. It was expected that players on hard mode would have lower guesses on average, for multiple reasons. One reason hypothesized in the paper was that always utilizing hints given by the game results in continual progress and eventual deduction of the word. Another reason is that players on hard mode may take the game more seriously or competitively, and therefore may have lower averages on virtue of using techniques or strategies that other players are indifferent to.

## Weaknesses and next steps

Weaknesses of the paper largely pertain to the dataset used and the significance and validation of the model used to predict gameplay. The dataset used in this paper was gathered from the author and is largely incomplete in observations. While Wordle has been released and its rise in popularity present for months as of writing, the dataset only covers two weeks and there is a gap in observations in between them. This results in an extremely small sample size which is not suitable for modelling and analysis of sample statistics. Prior articles referenced in the paper had analyzed Wordle games over several weeks or months, which would be a more appropriate timeframe in order to gather enough data to conduct a proper analysis on average gameplay statistics. While the dataset is lacking in sample size regarding days, the volume of tweets for each day recorded is sufficient to gauge public opinion on the recorded days accurately such that statistics such as geographical and gamemode proportions should hold. However, more data is required on both the amount of words and scores in order to make sure that the results in this paper hold.

The models created in the paper also struggle to hold as a result of the dataset and the lack of predictor variables. It is likely that the model does not hold when compared to the full population of Wordle games and words, as we have only taken a small sample of words which is prone to large amounts of variance. The models have not been tested for validation or regression coefficients which may result in an unfounded and inaccurate model. Future work may involve refinement of the models and addition of more predictor variables in order to accurately model the linguistic relationships between the chosen words and player performance.

It is the author's hope that the results from this paper serve as a start and as inspiration for continual research into word-based games and puzzle games in general. As a viral game, Wordle is a popular topic of discussion internationally and online, and already many spin-offs of the game have seen their own share of success. Future game designers and puzzle game makers may seek to use the results and impacts of Wordle to tailor their own works to provide a fun challenge for players, and research should be continued to determine how one can find a balance between difficulty and frustration as players of Wordle frequently experience. Researchers of linguistics may also find interest in Wordle as a method of gathering public knowledge and opinion of the English language. As the language continually evolves and words take on new meanings or become obsolete and forgotten, Wordle may be useful in teaching and reminding the general population of literacy much like the crossword puzzles of the past. Popular trends can not only find success among the general population and provide a source of bonding and discussion, but can also be a source of research in order to learn from these trends and utilize their implications for further purposes.

\newpage

\appendix

# Appendix {-}

## Datasheet

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to conduct analysis on Twitter trends regarding the popular online game Wordle. Twitter is a convenient website to gather large amounts of player statistics as users regularily and daily share their scores in an easy to analyze format.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The dataset was created by the author of this paper to serve the paper's purposes.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - No monetary cost was required in the creation of the dataset; it was created free of cost using the rtweet [@citertweet] package and the programming software R [@citeR].
4. *Any other comments?*
    - Due to data gathering errors, there is a gap in the temporal data of the dataset. A large majority of the tweets are gathered at a general timepoint correlating to when the tweets were obtained. As such, the tweets are not evenly distributed timewise.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- Each instance of an observation represents a message known as a 'tweet'. This message is akin to a text message which may contain emoticons, pictures, and GIFs attached with the message. Tweets can also be sent as replies to other tweets, or as a quote tweet which is standalone but is connected to an initial message.
2. *How many instances are there in total (of each type, if appropriate)?*
	- In total, there are 508000 instances and 508000 gathered tweets.
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- The dataset does not contain all possible instances. Billions of messages are sent and processed by Twitter daily and it is impossible to process and filter such a colossal stream of data to obtain only relevant data. The dataset is a sample of tweets relevant to Wordle and the larger set would consist of all relevant tweets made each day for the period of data gathering. The sample is not chronologically representative of the entire set because of limitations with data gathering. rtweet is limited in that it can only gather 'recent' tweets related to when the request is sent. Due to this, evenly distributed chronological coverage was not able to be gathered. 
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each instance consists of a message sent in text and unicode characters, which may have a picture or GIF attached. There are also properties of the tweet and sender account such as chronological data in the dataset.
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- Each tweet has a status id which is shown in the raw data. This status id is unique and directly corresponds to the pertaining message in the Twitter databse. 
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- There is no information missing that was blocked by Twitter or our data collecting programs. Information such as the location of the sender of the tweet or the account information may be intentionally withheld due to privacy concerns of the user.  
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- Relationships between invididual instances are made explicit. Part of the raw data contains information about if the instance was a reply or 'quote tweet' in response to an initial tweet. These initial tweets may or may not be in the dataset, and this can be verified by using the status id of the initial message and checking if it is in the dataset.
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- A datasplit of 80/20 for the training and testing datasets will be used to create and test models for the data. These splits will consist of randomly sampled partitions of the data and will be done so to mitigate bias from the models created.
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- Some sources of noise may include tweets which matched our criteria for relevance and were obtained but do not provide any usable information. There may also be 'joke' results which parody the data we wish to obtain but is itself not accurate information. 
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- The dataset relies on external resources, as we are looking at tweets. There are no guarantees that they will exist in the future as Twitter and its users have the right to delete and hide tweets at any moment. However, archival services exist which would allow for the messages to be preserved. As well, deleted tweets are stored in Twitter's databases, although unavailible to the public.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- The dataset may contain confidential data if the user explicitly and voluntarily mentioned this in their tweet. 
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- The dataset may contain offensive, insulting, and malicious text if the user chose to include such text in their original tweet.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- The dataset does not identify any sub-populations as this data is not obtainible from tweets.
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- It would be possible to identify individuals from the dataset if there were enough information in their account details and their message, as this data is included and gathered alongside the tweet instance.
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- The data contains the geographical location of where the tweet was sent and the geographical location of the user. There is no other sensitive information included.
16. *Any other comments?*
	- None

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data was directly observable from each tweet and user, as tweets are sent as raw text and other details are directly obtainible from the timestamp and self-provided user information. 
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- The data was collected using the rtweet package found in R. These procedures were validated by manually checking that the program was gathering valid tweets using the status id and referencing of tweets and accounts. 
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- The sampling strategy was to take recent tweets matching the search criteria as of the time of request. Recent tweets are shown in order of time posted.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- Only the author of the paper was involved in data collection.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The data was collected over two timeframes- on April 10 and on April 23, 2022. The timeframe roughly matches the creation timeframe of the data: The creation timeframes are roughly from April 4-10 and April 17-23, 2022. These times are when daily tweets are collected.
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- None
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- The data was collected from other sources, using the rtweet package.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- The individuals in question were not notified about the data collection.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- The individuals in question consented to the collection and use of their data in the Twitter Terms of Service which all users must agree to for site access. The link can be found here: https://twitter.com/en/tos

10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- Consenting individuals may choose to deactivate their account which would delete all messages and tweets they had sent from the platform.
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- None
12. *Any other comments?*
	- None

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- Cleaning of the data was done to obtain more relevance in our dataset. After the raw data was collected, observations were filtered by specific message criteria. The relevant data (Wordle scores) are contained in messages through specific phrases which are copied and pasted by users from the game's website. Scores take the form of the phrase "Wordle ___ _/6" where the number of the daily word and the number of guesses taken by the player are displayed. Tweets were filtered according to whether this phrase was included in their tweet. Once this was done, the numerical data of the number of guesses the user took was extracted from the message text and stored as a new variable. Games in which the user did not successfully guess the word in 6 tries, and therefore failed, were counted as taking 7 guesses to complete. After this was done, the date and time that the tweets were sent was extracted from the Unix time provided in the raw data using the lubridate [@citelubridate] package. A binary variable was also created indicating if the user had 'hard mode' enabled in their game, which adds additional conditions to the game making it more challenging. This indication was availible in the raw text; messages from hard mode players would contain a  "_/6*" in their tweet, the star indicating the use of hard mode. Once this was done, the data was aggregated into a single dataframe. Partitions were also made according to the creation timeframe of the tweets for potential future use.

2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- The "raw" data is contained in the repository that this paper is contained in.
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- The software used to clean the data consists of R packages, which have been cited and credited.
4. *Any other comments?*
	- None

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- None
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- Yes. Code and data are available at: https://github.com/Ivannoar/Twilight
3. *What (other) tasks could the dataset be used for?*
	- Other tasks might consist of more analysis pertaining to Wordle, as well as using the dataset to draw conclusions based on Twitter in general.
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- Tweet geographical and chronological data, as well as the status and description of the accounts used could be used to unfairly treat the individuals involved. A dataset consumer might avoid making these conclusions and taking the data at more of face value.
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- The dataset should not be used for any malicious or illegal activity.
6. *Any other comments?*
	- None

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- The dataset will not be explicitly distributed to third parties, but will be made availible publically online.
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- The dataset will be distributed on GitHub via this paper's repository.
3. *When will the dataset be distributed?*
	- The dataset will be distributed on April 27, 2022.
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- None
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- None
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- None
7. *Any other comments?*
	- None

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- The dataset will be hosted by the author of the paper.
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- The owner of the dataset can be reached via the GitHub account the repository is hosted on.
3. *Is there an erratum? If so, please provide a link or other access point.*
	- None
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- The dataset will not be updated.
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- There are no limits on the retention of the data associated with the instances.
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- Older versions of the dataset may be hosted by the author of the paper on their local files. The obsolescence will be communicated from the repository the paper is hosted on.
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- Others can contribute to the dataset via GitHub's built-in collaboration features, which will be verified and finalized by the author.
8. *Any other comments?*
	- None

\newpage

# References

### 2024-03-25 outputs/paper/paper.qmd
qmd and bib file structured

---
title: ""
subtitle: ""
author: 
  - Ziheng Zhong
thanks: "Code and data are available at: https://github.com/iJustinn/House_Price.git"
date: today
date-format: long
abstract: ""
format: pdf
toc: true
toc-depth: 3
number-sections: true
bibliography: references.bib
---

```{r}
#| message: false
#| echo: false

# load packages
packages <- c("tidyverse", "boot", "broom.mixed", "collapse", "dataverse", "gutenbergr", "janitor", "knitr", "marginaleffects", "modelsummary", "rstanarm", "tidybayes", "loo")

# Function to check and install missing packages
install_and_load <- function(packages) {
  for (package in packages) {
    if (!require(package, character.only = TRUE)) {
      install.packages(package)
      library(package, character.only = TRUE)
    }
  }
}

# Call the function with the list of packages
install_and_load(packages)
```

```{r}
#| message: false
#| echo: false

# read in data
file_path_0 <- "../../inputs/data/raw_data.csv"
file_path_1 <- "../../outputs/data/cleaned_data.csv"
file_path_2 <- "../../outputs/data/top_5_cause_data.csv"
file_path_3 <- "../../outputs/data/summarized_top_5_cause_data.csv"

raw_data <- read_csv(file_path_0, show_col_types = FALSE)
cleaned_data <- read_csv(file_path_1, show_col_types = FALSE)
top_5_cause_data <- read_csv(file_path_2, show_col_types = FALSE)
summarized_top_5_cause_data <- read_csv(file_path_3, show_col_types = FALSE)
```

# Introduction

# Data {#sec-data}

Data used in this paper was cleaned, processed and tested with the programming language R [@citeR]. Also with support of additional packages in R: `tidyverse` [@citeTidyverse], `ggplot2` [@citeGgplot], `janitor` [@citeJanitor], `readr` [@citeReadr], `knitr` [@citeKnitr], `rstanarm` [@citeRstanarm], `modelsummary` [@citeModelsummary], `tidybayes` [@citeTidybayes], `loo` [@citeLoo], `testthat` [@citeTestthat].

## Source

## Method

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: tbl-causes
#| tbl-cap: top 8 major causes in 2019

```

# Results {#sec-res}

```{r}
#| include: false

```

## Data Trend

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: fig-trends-1
#| fig-cap: Trend for Alberta Mortality-Top 5 Causes

```

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: fig-trends-2
#| fig-cap: Trend for Alberta Mortality-Sum of Causes

```

## Modeling

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: tbl-summary
#| tbl-cap: summary of modeling

```

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: fig-models
#| fig-cap: Model Results
#| fig-subcap: ["Poisson Model","Negative Binomial Model"]
#| layout-ncol: 2
#| layout-nrow: 1

```

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: tbl-cross_validation
#| tbl-cap: cross validation of modeling

```

# Discussion {#sec-dis}

## Demographic Shifts

## Health-related Behaviors

## Government Policies

## Environmental Changes

## Possible Improvements

# Conclusion {#sec-con}

\newpage

\newpage

# References

### 2024-03-25 scripts/00-data_simulation.R
simulation data added

@@ -1,13 +1,49 @@
-set.seed(1234)
+#### Preamble ####
+# Purpose: 
+# Author: Ziheng Zhong
+# Date: 25 March 2024
+# Contact: ziheng.zhong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: none
 
-simulated_data <-
-  tibble(
-    'Date' = sample(seq(as.Date('2022/04/05'), as.Date('2022/04/23'), by="day"), 10, replace = TRUE),
-         
-    'Time' = sample(format(seq(as.POSIXct('00:00:00', format = "%T"), 
-                    as.POSIXct('23:59:59', format = "%T"), by = "sec"), "%T"), 10, replace = TRUE),
-    'Location' = sample(countrycode::codelist$country.name.en, 10),     
-    'Score' = sample(x = c('1/6', '2/6', '3/6', '4/6', '5/6', '6/6', 'DNF'), 10, replace = TRUE),
-    'Hardmode' = sample(c(TRUE, FALSE), 10, replace = TRUE)
+
+
+#### Workspace setup ####
+library(dplyr) # for data manipulation
+library(ggplot2) # for data visualization, part of the tidyverse
+
+
+
+#### Sumulation ####
+set.seed(123)  # for reproducibility
+
+simulate_dataset <- function(n) { # Function to simulate the dataset
+  longitude <- rnorm(n, mean = -100, sd = 1)
+  latitude <- rnorm(n, mean = 30, sd = 1)
+  housing_median_age <- sample(1:50, n, replace = TRUE)
+  total_rooms <- rpois(n, lambda = 100)
+  households <- rpois(n, lambda = 100)
+  median_income <- rlnorm(n, meanlog = 10, sdlog = 0.5) / 10000
+  median_house_value <- rlnorm(n, meanlog = 15, sdlog = 1)
+  ocean_proximity <- factor(
+    sample(c("<1H OCEAN", "INLAND", "NEAR OCEAN", "NEAR BAY", "ISLAND"), n, replace = TRUE),
+    levels = c("<1H OCEAN", "INLAND", "NEAR OCEAN", "NEAR BAY", "ISLAND")
   )
-view(simulated_data)
+  
+  return(data.frame(
+    longitude,
+    latitude,
+    housing_median_age,
+    total_rooms,
+    total_bedrooms = total_rooms * runif(n, 0.5, 1), # assuming bedrooms are 50-100% of total rooms
[…]

### 2024-03-25 scripts/01-data_cleaning.R

@@ -1,18 +1,14 @@
 #### Preamble ####
-# Purpose: Clean the survey data downloaded from [...UPDATE ME!!!!!]
-# Author: Ivan Li
-# Date: April 3 2022
-# Contact: ivann.li@mail.utoronto.ca 
+# Purpose: 
+# Author: Ziheng Zhong
+# Date: 25 March 2024
+# Contact: ziheng.zhong@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 
-# - Need to have downloaded the ACS data and saved it to inputs/data
-# - Don't forget to gitignore it!
-# - Change these to yours
-# Any other information needed?
+# Pre-requisites: none
+
 
 
 #### Workspace setup ####
-# Use R Projects, not setwd().
 
 library(haven)
 library(tidyverse)
@@ -20,292 +16,3 @@ library(rtweet)
 library(lubridate)
 
 
-
-
-# Read in the raw data.
-
-consumer_key <- '***'
-consumer_secret <- '***'
-access_token <- '***'
-access_secret <- '***'
-
-# https://github.com/sarikayamehmet/GetOldTweets-R
-
-get_token()
-create_token(app = "uoftwitter", consumer_key, consumer_secret, access_token, access_secret)
-setup_twitter_oauth(consumer_key, consumer_secret, access_token, access_secret)
-get_favorites(user = "RohanAlexander")
-
-# Obtaining old tweets
-april4 <- search_tweets(q = "Wordle 289", n = 36000, retryonratelimit = TRUE, type = "mixed")
-  df <- apply(april4,2,as.character)
-  write.csv(x=df, file = "inputs/data/april_4.csv")
-april5 <- search_tweets(q = "Wordle 290", n = 36000, retryonratelimit = TRUE, type = "mixed")
-  df <- apply(april5,2,as.character)
-  write.csv(x=df, file = "inputs/data/april_5.csv")
-april6 <- search_tweets(q = "Wordle 291", n = 36000, retryonratelimit = TRUE, type = "mixed")
-  df <- apply(april6,2,as.character)
-  write.csv(x=df, file = "inputs/data/april_6.csv")
-april7 <- search_tweets(q = "Wordle 292", n = 36000, retryonratelimit = TRUE, type = "mixed")
-  df <- apply(april7,2,as.character)
-  write.csv(x=df, file = "inputs/data/april_7.csv")
[…]

### 2024-03-29 outputs/paper/paper.qmd
clean data script updated, clean data added, paper updated accordingly

@@
 # load packages
-packages <- c("tidyverse", "boot", "broom.mixed", "collapse", "dataverse", "gutenbergr", "janitor", "knitr", "marginaleffects", "modelsummary", "rstanarm", "tidybayes", "loo")
+packages <- c("tidyverse", "boot", "broom.mixed", "collapse", "dataverse", "gutenbergr", "janitor", "knitr", "marginaleffects", "modelsummary", "rstanarm", "tidybayes", "loo", "kableExtra")
 
@@
 # read in data
-file_path_0 <- "../../inputs/data/raw_data.csv"
-file_path_1 <- "../../outputs/data/cleaned_data.csv"
-file_path_2 <- "../../outputs/data/top_5_cause_data.csv"
-file_path_3 <- "../../outputs/data/summarized_top_5_cause_data.csv"
+file_path_0 <- "../../inputs/data/cali_house_price.csv"
+file_path_1 <- "../../outputs/data/cali_house_price_clean.csv"
+file_path_2 <- "../../outputs/data/.csv"
+file_path_3 <- "../../outputs/data/.csv"
 
-raw_data <- read_csv(file_path_0, show_col_types = FALSE)
-cleaned_data <- read_csv(file_path_1, show_col_types = FALSE)
-top_5_cause_data <- read_csv(file_path_2, show_col_types = FALSE)
-summarized_top_5_cause_data <- read_csv(file_path_3, show_col_types = FALSE)
+cali_house_price <- read_csv(file_path_0, show_col_types = FALSE)
+cali_house_price_clean<- read_csv(file_path_1, show_col_types = FALSE)
+
 ```
@@
 Data used in this paper was cleaned, processed and tested with the programming language R [@citeR].
-Also with support of additional packages in R: `tidyverse` [@citeTidyverse], `ggplot2` [@citeGgplot], `janitor` [@citeJanitor], `readr` [@citeReadr], `knitr` [@citeKnitr], `rstanarm` [@citeRstanarm], `modelsummary` [@citeModelsummary], `tidybayes` [@citeTidybayes], `loo` [@citeLoo], `testthat` [@citeTestthat].
+Also with support of additional packages in R: `tidyverse` [@citeTidyverse], `ggplot2` [@citeGgplot], `janitor` [@citeJanitor], `readr` [@citeReadr], `knitr` [@citeKnitr], `rstanarm` [@citeRstanarm], `modelsummary` [@citeModelsummary], `tidybayes` [@citeTidybayes], `loo` [@citeLoo], `testthat` [@citeTestthat], `KableExtra` [@citeKableEx].
 
@@
 #| fig-pos: 'h'
-#| label: tbl-causes
-#| tbl-cap: top 8 major causes in 2019
+#| label: tbl-sum
+#| tbl-cap: Summary statistics of the California housing dataset
+#| layout-ncol: 3
+#| layout-nrow: 3
 
+longitude_summary <- summary(cali_house_price %>% select(longitude))
+latitude_summary <- summary(cali_house_price %>% select(latitude))
+housing_median_age_summary <- summary(cali_house_price %>% select(housing_median_age))
+total_rooms_summary <- summary(cali_house_price %>% select(total_rooms))
+total_bedrooms_summary <- summary(cali_house_price %>% select(total_bedrooms))
+population_summary <- summary(cali_house_price %>% select(population))
+households_summary <- summary(cali_house_price %>% select(households))
+median_income_summary <- summary(cali_house_price %>% select(median_income))
+median_house_value_summary <- summary(cali_house_price %>% select(median_house_value))
+
+kable(longitude_summary)
+kable(latitude_summary)
+kable(housing_median_age_summary)
+kable(total_rooms_summary)
+kable(total_bedrooms_summary)
+kable(population_summary)
+kable(households_summary)
+kable(median_income_summary)
+kable(median_house_value_summary)
+
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-pos: 'h'
+#| label: tbl-null-raw
+#| tbl-cap: Count of missing values for each variable
+
+null_values <- cali_house_price %>%
+  summarise_all(~sum(is.na(.))) %>%
+  gather(key = "Variables", value = "MissingValues")
+
+kable(null_values)
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-pos: 'h'
+#| label: tbl-null-clean
+#| tbl-cap: Count of missing values for each variable after cleaning
+
+null_values <- cali_house_price_clean %>%
+  summarise_all(~sum(is.na(.))) %>%
+  gather(key = "Variables", value = "MissingValues")
+
+kable(null_values)
 ```

### 2024-03-29 scripts/01-data_cleaning.R

@@ -10,9 +10,28 @@
 
 #### Workspace setup ####
 
-library(haven)
 library(tidyverse)
-library(rtweet)
-library(lubridate)
+library(dplyr)
+
+
+
+#### load data ####
+cali_house_price <- read_csv("inputs/data/cali_house_price.csv")
+
+
+
+#### Cleaning ####
+total_bedrooms_mean_value <- mean(cali_house_price$total_bedrooms, na.rm = TRUE) # Calculate the mean of the total_bedrooms column, excluding NA values
+
+cali_house_price_clean <- cali_house_price %>%
+  mutate(total_bedrooms = ifelse(is.na(total_bedrooms), total_bedrooms_mean_value, total_bedrooms)) # Replace NA values with the mean value in the total_bedrooms column
+
+
+
+#### save data ####
+write_csv(
+  x = cali_house_price_clean,
+  file = "outputs/data/cali_house_price_clean.csv"
+)
 
 

### 2024-03-29 scripts/02-modeling.R

@@ -1,0 +1,22 @@
+#### Preamble ####
+# Purpose: 
+# Author: Ziheng Zhong
+# Date: 25 March 2024
+# Contact: ziheng.zhong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: none
+
+
+
+#### Workspace setup ####
+
+
+
+#### load data ####
+top_5_cause_data <- read_csv("outputs/data/top_5_cause_data.csv")
+
+
+
+#### poisson model ####
+
+

### 2024-03-29 scripts/03-tests.R

@@ -1,0 +1,25 @@
+#### Preamble ####
+# Purpose: The code defines a test suite for validating data integrity, including checks for positive calendar years, four-digit year format, non-empty cause entries, and non-negative death counts across multiple datasets within a research project.
+# Author: Ziheng Zhong
+# Date: 25 March 2024
+# Contact: ziheng.zhong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: none
+
+
+
+#### Workspace setup ####
+
+
+
+#### load data ####
+setwd("/Users/ijustin/Library/CloudStorage/Dropbox/Code/STA302/Research_Papers/Mortality_Rate_in_Alberta")
+raw_data <- readr::read_csv("inputs/data/raw_data.csv")
+top_5_cause_data <- read_csv("outputs/data/top_5_cause_data.csv")
+summarized_top_5_cause_data <- read_csv("outputs/data/summarized_top_5_cause_data.csv")
+
+
+
+#### raw_data testing ####
+
+

### 2024-04-02 scripts/01-data_cleaning.R
data cleaning script updated, ocean_proximity_frequency data added

@@ -28,10 +28,35 @@ cali_house_price_clean <- cali_house_price %>%
 
 
 
+cali_house_price_clean %>% 
+  summarise(across(where(is.numeric), ~mean(.x, na.rm = TRUE))) %>% 
+  walk2(names(.), ~cat("The mean value of", .y, ":", .x, "\n")) # Calculating and printing mean of each numeric column
+
+
+
+columns_needed <- c("ocean_proximity") # Specify the needed columns
+
+ocean_proximity_frequency <- data.frame(Variable = character(), Frequency = integer()) # Initialize an empty data frame for storing frequencies
+
+for (j in columns_needed) {
+  temp <- as.data.frame(table(cali_house_price_clean[[j]]))  # Calculate frequency using table and convert to data frame
+  
+  colnames(temp) <- c("Variable", "Frequency")  # Rename the columns accordingly
+  
+  ocean_proximity_frequency <- bind_rows(ocean_proximity_frequency, temp)  # Combine with the main dataset
+}
+
+print(ocean_proximity_frequency) # Show the resulting dataset
+
+
+
 #### save data ####
 write_csv(
   x = cali_house_price_clean,
   file = "outputs/data/cali_house_price_clean.csv"
 )
 
-
+write_csv(
+  x = ocean_proximity_frequency,
+  file = "outputs/data/ocean_proximity_frequency.csv"
+)

### 2024-04-02 outputs/paper/paper.qmd
paper updated

@@
 ---
-title: ""
-subtitle: ""
+title: "TBD"
+subtitle: "TBD"
 author: 
@@
 date-format: long
-abstract: ""
+abstract: "TBD"
 format: pdf
@@
 file_path_1 <- "../../outputs/data/cali_house_price_clean.csv"
-file_path_2 <- "../../outputs/data/.csv"
+file_path_2 <- "../../outputs/data/ocean_proximity_frequency.csv"
 file_path_3 <- "../../outputs/data/.csv"
@@
 cali_house_price_clean<- read_csv(file_path_1, show_col_types = FALSE)
+ocean_proximity_frequency<- read_csv(file_path_2, show_col_types = FALSE)
 
@@
 #| fig-pos: 'h'
-#| label: fig-trends-1
-#| fig-cap: Trend for Alberta Mortality-Top 5 Causes
+#| label: fig-ocean_proximity_frequency
+#| fig-cap: Frequency of Ocean Proximity
+
+ggplot(ocean_proximity_frequency, aes(x = Variable, y = Frequency)) +
+  geom_bar(stat = "identity") +
+  theme_minimal() +
+  labs(title = "Frequency of Categories in Ocean Proximity",
+       x = "Ocean Proximity",
+       y = "Frequency") +
+  coord_flip()
 
@@
 #| fig-pos: 'h'
-#| label: fig-trends-2
-#| fig-cap: Trend for Alberta Mortality-Sum of Causes
+#| label: fig-TBD_1
+#| fig-cap: TBD
 
@@
 #| fig-pos: 'h'
-#| label: tbl-summary
-#| tbl-cap: summary of modeling
+#| label: tbl-TBD_2
+#| tbl-cap: TBD
 
@@
 #| fig-pos: 'h'
-#| label: fig-models
-#| fig-cap: Model Results
-#| fig-subcap: ["Poisson Model","Negative Binomial Model"]
-#| layout-ncol: 2
-#| layout-nrow: 1
+#| label: fig-TBD_3
+#| fig-cap: TBD
 
@@
 #| fig-pos: 'h'
-#| label: tbl-cross_validation
-#| tbl-cap: cross validation of modeling
+#| label: tbl-TBD_4
+#| tbl-cap: TBD
 

### 2024-04-02 scripts/03-tests.R
test script updated

@@ -9,17 +9,91 @@
 
 
 #### Workspace setup ####
+library(readr)
+library(testthat)
+library(tidyverse)
 
 
 
-#### load data ####
-setwd("/Users/ijustin/Library/CloudStorage/Dropbox/Code/STA302/Research_Papers/Mortality_Rate_in_Alberta")
-raw_data <- readr::read_csv("inputs/data/raw_data.csv")
-top_5_cause_data <- read_csv("outputs/data/top_5_cause_data.csv")
-summarized_top_5_cause_data <- read_csv("outputs/data/summarized_top_5_cause_data.csv")
+#### cali_house_price_clean testing ####
+test_that("cali_house_price_clean testing", {
+  cali_house_price_clean<- readr::read_csv("../../House_Price/outputs/data/cali_house_price_clean.csv")
+  
+  # Test if 'longitude' values are within expected range
+  expect_true(all(cali_house_price_clean$longitude >= -180 & cali_house_price_clean$longitude <= 180), 
+              info = "Longitude values are within expected range")
+  
+  # Test if 'latitude' values are within expected range
+  expect_true(all(cali_house_price_clean$latitude >= -90 & cali_house_price_clean$latitude <= 90),
+              info = "Latitude values are within expected range")
+  
+  # Test if 'housing_median_age' values are positive
+  expect_true(all(cali_house_price_clean$housing_median_age > 0),
+              info = "Housing median age values are positive")
+  
+  # Test if 'total_rooms' values are positive
+  expect_true(all(cali_house_price_clean$total_rooms > 0),
+              info = "Total rooms values are positive")
+  
+  # Test if 'total_bedrooms' values are positive
+  expect_true(all(cali_house_price_clean$total_bedrooms > 0),
+              info = "Total bedrooms values are positive")
+  
+  # Test if 'population' values are positive
+  expect_true(all(cali_house_price_clean$population > 0),
+              info = "Population values are positive")
+  
+  # Test if 'households' values are positive
+  expect_true(all(cali_house_price_clean$households > 0),
+              info = "Households values are positive")
+  
+  # Test if 'median_income' values are positive
+  expect_true(all(cali_house_price_clean$median_income > 0),
+              info = "Median income values are positive")
+  
+  # Test if 'median_house_value' values are positive
+  expect_true(all(cali_house_price_clean$median_house_value > 0),
+              info = "Median house value values are positive")
+  
[…]

### 2024-04-06 outputs/paper/paper.qmd
dataset changed, rearrange files

@@
 
+library(maps)
+library(dplyr)
+library(readr)
+library(viridis)
+library(ggplot2)
+
+# Generating fake data for average house prices for each US state
+set.seed(42) # For reproducibility
+states <- map_data("state")$region %>% unique()
+average_house_prices <- runif(n = length(states), min = 100000, max = 700000)
+state_average_prices <- data.frame(State = states, AverageHousePrice = average_house_prices)
+
+# Getting map data
+us_states_map <- map_data("state")
+
+# Merging the map data with the fake average house prices
+map_data_merged <- merge(us_states_map, state_average_prices, by.x = "region", by.y = "State")
+
+# Creating the heatmap
+ggplot() +
+  geom_polygon(data = map_data_merged, aes(x = long, y = lat, group = group, fill = AverageHousePrice), color = "white") +
+  coord_fixed(1.3) +
+  viridis::scale_fill_viridis(name = "Avg House Price", option = "plasma") +
+  theme_minimal() +
+  labs(title = "Average House Price by State in the US (Fake Data)",
+       subtitle = "Data: Generated",
+       caption = "Visualization by Example") +
+  theme(legend.position = "right")
+
 ```

### 2024-04-07 scripts/00-data_simulation.R
data cleaning scripts updated, processed data added

@@ -9,41 +9,33 @@
 
 
 #### Workspace setup ####
-library(dplyr) # for data manipulation
-library(ggplot2) # for data visualization, part of the tidyverse
+library(maps)
+library(dplyr)
+library(ggplot2)
 
 
 
 #### Sumulation ####
 set.seed(123)  # for reproducibility
 
-simulate_dataset <- function(n) { # Function to simulate the dataset
-  longitude <- rnorm(n, mean = -100, sd = 1)
-  latitude <- rnorm(n, mean = 30, sd = 1)
-  housing_median_age <- sample(1:50, n, replace = TRUE)
-  total_rooms <- rpois(n, lambda = 100)
-  households <- rpois(n, lambda = 100)
-  median_income <- rlnorm(n, meanlog = 10, sdlog = 0.5) / 10000
-  median_house_value <- rlnorm(n, meanlog = 15, sdlog = 1)
-  ocean_proximity <- factor(
-    sample(c("<1H OCEAN", "INLAND", "NEAR OCEAN", "NEAR BAY", "ISLAND"), n, replace = TRUE),
-    levels = c("<1H OCEAN", "INLAND", "NEAR OCEAN", "NEAR BAY", "ISLAND")
-  )
-  
-  return(data.frame(
-    longitude,
-    latitude,
-    housing_median_age,
-    total_rooms,
-    total_bedrooms = total_rooms * runif(n, 0.5, 1), # assuming bedrooms are 50-100% of total rooms
-    population = households * sample(2:5, n, replace = TRUE), # assuming each household has 2-5 members
-    households,
-    median_income,
-    median_house_value,
-    ocean_proximity
-  ))
-}
+average_house_prices <- runif(n = length(states), min = 100000, max = 700000) # Simulating data for average house prices for each US state
+state_average_prices <- data.frame(State = states, AverageHousePrice = average_house_prices) 
+simulation_data <- merge(us_states_map, state_average_prices, by.x = "region", by.y = "State")
 
-simulation_data <- simulate_dataset(100) # Generate the simulation_data
+head(simulation_data) # preview simulation data
+tail(simulation_data)
 
-head(simulation_data) # Print the first few rows of the simulation_data
+
+
+#### Simulation Chart ####
+us_states_map <- map_data("state") # Getting map data
+
+ggplot() + # Creating the heatmap
+  geom_polygon(data = simulation_data, aes(x = long, y = lat, group = group, fill = AverageHousePrice), color = "white") +
[…]

### 2024-04-07 scripts/01-data_cleaning.R

@@ -9,54 +9,55 @@
 
 
 #### Workspace setup ####
-
-library(tidyverse)
 library(dplyr)
+library(tidyverse)
+library(lubridate)
 
 
 
 #### load data ####
-cali_house_price <- read_csv("inputs/data/cali_house_price.csv")
+raw_data <- read_csv("inputs/data/all_house_price.csv")
 
 
 
 #### Cleaning ####
-total_bedrooms_mean_value <- mean(cali_house_price$total_bedrooms, na.rm = TRUE) # Calculate the mean of the total_bedrooms column, excluding NA values
+clean_house_price <- raw_data %>% # Delete the first row since it cause issues in further analysis (we focus on state but the first row is data for U.S. as a whole)
+  slice(-1)
 
-cali_house_price_clean <- cali_house_price %>%
-  mutate(total_bedrooms = ifelse(is.na(total_bedrooms), total_bedrooms_mean_value, total_bedrooms)) # Replace NA values with the mean value in the total_bedrooms column
+clean_house_price <- clean_house_price %>% # Remove specific columns
+  select(-c(SizeRank, RegionID, RegionType, RegionName))
 
+#### Further Cleaning ####
+# By State - monthly
+state_house_price_by_month <- clean_house_price %>%
+  group_by(StateName) %>%
+  summarise(across(where(is.numeric), mean, na.rm = TRUE)) # Calculating means house price of each state excluding NA
 
-
-cali_house_price_clean %>% 
-  summarise(across(where(is.numeric), ~mean(.x, na.rm = TRUE))) %>% 
-  walk2(names(.), ~cat("The mean value of", .y, ":", .x, "\n")) # Calculating and printing mean of each numeric column
-
-
-
-columns_needed <- c("ocean_proximity") # Specify the needed columns
-
-ocean_proximity_frequency <- data.frame(Variable = character(), Frequency = integer()) # Initialize an empty data frame for storing frequencies
-
-for (j in columns_needed) {
-  temp <- as.data.frame(table(cali_house_price_clean[[j]]))  # Calculate frequency using table and convert to data frame
-  
-  colnames(temp) <- c("Variable", "Frequency")  # Rename the columns accordingly
-  
-  ocean_proximity_frequency <- bind_rows(ocean_proximity_frequency, temp)  # Combine with the main dataset
-}
-
-print(ocean_proximity_frequency) # Show the resulting dataset
+# By State - yearly
[…]

### 2024-04-07 outputs/paper/paper.qmd
paper updated

@@
 # Generating fake data for average house prices for each US state
-set.seed(42) # For reproducibility
+set.seed(123) # For reproducibility
 states <- map_data("state")$region %>% unique()

### 2024-04-07 scripts/01_01-data_cleaning.R
more data added, associate scripts added

@@ -1,0 +1,63 @@
+#### Preamble ####
+# Purpose: 
+# Author: Ziheng Zhong
+# Date: 25 March 2024
+# Contact: ziheng.zhong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: none
+
+
+
+#### Workspace setup ####
+library(dplyr)
+library(tidyverse)
+library(lubridate)
+
+
+
+#### load data ####
+raw_data <- read_csv("inputs/data/1b_house_price.csv")
+
+
+
+#### Cleaning ####
+clean_house_price <- raw_data %>% # Delete the first row since it cause issues in further analysis (we focus on state but the first row is data for U.S. as a whole)
+  slice(-1)
+
+clean_house_price <- clean_house_price %>% # Remove specific columns
+  select(-c(SizeRank, RegionID, RegionType, RegionName))
+
+#### Further Cleaning ####
+# By State - monthly
+state_house_price_by_month <- clean_house_price %>%
+  group_by(StateName) %>%
+  summarise(across(where(is.numeric), mean, na.rm = TRUE)) # Calculating means house price of each state excluding NA
+
+# By State - yearly
+state_house_price_by_year <- state_house_price_by_month %>%
+  pivot_longer( # Gather the monthly columns into key-value pairs, convert the keys to Dates, extract the year, and then calculate the mean
+    cols = starts_with("20"), # Assuming all your date columns start with '20'
+    names_to = "Month",
+    values_to = "HousePrice"
+  ) %>%
+  mutate(Year = year(ymd(Month))) %>%
+  group_by(StateName, Year) %>%
+  summarise(AvgHousePrice = mean(HousePrice, na.rm = TRUE), .groups = 'drop') # Calculating mean house price by year for each state
+
+
+
+#### save data ####
+write_csv(
+  x = clean_house_price,
+  file = "outputs/data/clean_1b_house_price.csv"
+)
+
+write_csv(
+  x = state_house_price_by_month,
+  file = "outputs/data/state_1b_house_price_by_month.csv"
+)
+
+write_csv(
+  x = state_house_price_by_year,
+  file = "outputs/data/state_1b_house_price_by_year.csv"
+)

### 2024-04-07 scripts/01_02-data_cleaning.R

@@ -1,0 +1,63 @@
+#### Preamble ####
+# Purpose: 
+# Author: Ziheng Zhong
+# Date: 25 March 2024
+# Contact: ziheng.zhong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: none
+
+
+
+#### Workspace setup ####
+library(dplyr)
+library(tidyverse)
+library(lubridate)
+
+
+
+#### load data ####
+raw_data <- read_csv("inputs/data/2b_house_price.csv")
+
+
+
+#### Cleaning ####
+clean_house_price <- raw_data %>% # Delete the first row since it cause issues in further analysis (we focus on state but the first row is data for U.S. as a whole)
+  slice(-1)
+
+clean_house_price <- clean_house_price %>% # Remove specific columns
+  select(-c(SizeRank, RegionID, RegionType, RegionName))
+
+#### Further Cleaning ####
+# By State - monthly
+state_house_price_by_month <- clean_house_price %>%
+  group_by(StateName) %>%
+  summarise(across(where(is.numeric), mean, na.rm = TRUE)) # Calculating means house price of each state excluding NA
+
+# By State - yearly
+state_house_price_by_year <- state_house_price_by_month %>%
+  pivot_longer( # Gather the monthly columns into key-value pairs, convert the keys to Dates, extract the year, and then calculate the mean
+    cols = starts_with("20"), # Assuming all your date columns start with '20'
+    names_to = "Month",
+    values_to = "HousePrice"
+  ) %>%
+  mutate(Year = year(ymd(Month))) %>%
+  group_by(StateName, Year) %>%
+  summarise(AvgHousePrice = mean(HousePrice, na.rm = TRUE), .groups = 'drop') # Calculating mean house price by year for each state
+
+
+
+#### save data ####
+write_csv(
+  x = clean_house_price,
+  file = "outputs/data/clean_2b_house_price.csv"
+)
+
+write_csv(
+  x = state_house_price_by_month,
+  file = "outputs/data/state_2b_house_price_by_month.csv"
+)
+
+write_csv(
+  x = state_house_price_by_year,
+  file = "outputs/data/state_2b_house_price_by_year.csv"
+)

### 2024-04-07 scripts/01_03-data_cleaning.R

@@ -1,0 +1,63 @@
+#### Preamble ####
+# Purpose: 
+# Author: Ziheng Zhong
+# Date: 25 March 2024
+# Contact: ziheng.zhong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: none
+
+
+
+#### Workspace setup ####
+library(dplyr)
+library(tidyverse)
+library(lubridate)
+
+
+
+#### load data ####
+raw_data <- read_csv("inputs/data/3b_house_price.csv")
+
+
+
+#### Cleaning ####
+clean_house_price <- raw_data %>% # Delete the first row since it cause issues in further analysis (we focus on state but the first row is data for U.S. as a whole)
+  slice(-1)
+
+clean_house_price <- clean_house_price %>% # Remove specific columns
+  select(-c(SizeRank, RegionID, RegionType, RegionName))
+
+#### Further Cleaning ####
+# By State - monthly
+state_house_price_by_month <- clean_house_price %>%
+  group_by(StateName) %>%
+  summarise(across(where(is.numeric), mean, na.rm = TRUE)) # Calculating means house price of each state excluding NA
+
+# By State - yearly
+state_house_price_by_year <- state_house_price_by_month %>%
+  pivot_longer( # Gather the monthly columns into key-value pairs, convert the keys to Dates, extract the year, and then calculate the mean
+    cols = starts_with("20"), # Assuming all your date columns start with '20'
+    names_to = "Month",
+    values_to = "HousePrice"
+  ) %>%
+  mutate(Year = year(ymd(Month))) %>%
+  group_by(StateName, Year) %>%
+  summarise(AvgHousePrice = mean(HousePrice, na.rm = TRUE), .groups = 'drop') # Calculating mean house price by year for each state
+
+
+
+#### save data ####
+write_csv(
+  x = clean_house_price,
+  file = "outputs/data/clean_3b_house_price.csv"
+)
+
+write_csv(
+  x = state_house_price_by_month,
+  file = "outputs/data/state_3b_house_price_by_month.csv"
+)
+
+write_csv(
+  x = state_house_price_by_year,
+  file = "outputs/data/state_3b_house_price_by_year.csv"
+)

### 2024-04-07 scripts/01_04-data_cleaning.R

@@ -1,0 +1,63 @@
+#### Preamble ####
+# Purpose: 
+# Author: Ziheng Zhong
+# Date: 25 March 2024
+# Contact: ziheng.zhong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: none
+
+
+
+#### Workspace setup ####
+library(dplyr)
+library(tidyverse)
+library(lubridate)
+
+
+
+#### load data ####
+raw_data <- read_csv("inputs/data/4b_house_price.csv")
+
+
+
+#### Cleaning ####
+clean_house_price <- raw_data %>% # Delete the first row since it cause issues in further analysis (we focus on state but the first row is data for U.S. as a whole)
+  slice(-1)
+
+clean_house_price <- clean_house_price %>% # Remove specific columns
+  select(-c(SizeRank, RegionID, RegionType, RegionName))
+
+#### Further Cleaning ####
+# By State - monthly
+state_house_price_by_month <- clean_house_price %>%
+  group_by(StateName) %>%
+  summarise(across(where(is.numeric), mean, na.rm = TRUE)) # Calculating means house price of each state excluding NA
+
+# By State - yearly
+state_house_price_by_year <- state_house_price_by_month %>%
+  pivot_longer( # Gather the monthly columns into key-value pairs, convert the keys to Dates, extract the year, and then calculate the mean
+    cols = starts_with("20"), # Assuming all your date columns start with '20'
+    names_to = "Month",
+    values_to = "HousePrice"
+  ) %>%
+  mutate(Year = year(ymd(Month))) %>%
+  group_by(StateName, Year) %>%
+  summarise(AvgHousePrice = mean(HousePrice, na.rm = TRUE), .groups = 'drop') # Calculating mean house price by year for each state
+
+
+
+#### save data ####
+write_csv(
+  x = clean_house_price,
+  file = "outputs/data/clean_4b_house_price.csv"
+)
+
+write_csv(
+  x = state_house_price_by_month,
+  file = "outputs/data/state_4b_house_price_by_month.csv"
+)
+
+write_csv(
+  x = state_house_price_by_year,
+  file = "outputs/data/state_4b_house_price_by_year.csv"
+)

### 2024-04-07 scripts/01_05-data_cleaning.R

@@ -1,0 +1,63 @@
+#### Preamble ####
+# Purpose: 
+# Author: Ziheng Zhong
+# Date: 25 March 2024
+# Contact: ziheng.zhong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: none
+
+
+
+#### Workspace setup ####
+library(dplyr)
+library(tidyverse)
+library(lubridate)
+
+
+
+#### load data ####
+raw_data <- read_csv("inputs/data/5bplus_house_price.csv")
+
+
+
+#### Cleaning ####
+clean_house_price <- raw_data %>% # Delete the first row since it cause issues in further analysis (we focus on state but the first row is data for U.S. as a whole)
+  slice(-1)
+
+clean_house_price <- clean_house_price %>% # Remove specific columns
+  select(-c(SizeRank, RegionID, RegionType, RegionName))
+
+#### Further Cleaning ####
+# By State - monthly
+state_house_price_by_month <- clean_house_price %>%
+  group_by(StateName) %>%
+  summarise(across(where(is.numeric), mean, na.rm = TRUE)) # Calculating means house price of each state excluding NA
+
+# By State - yearly
+state_house_price_by_year <- state_house_price_by_month %>%
+  pivot_longer( # Gather the monthly columns into key-value pairs, convert the keys to Dates, extract the year, and then calculate the mean
+    cols = starts_with("20"), # Assuming all your date columns start with '20'
+    names_to = "Month",
+    values_to = "HousePrice"
+  ) %>%
+  mutate(Year = year(ymd(Month))) %>%
+  group_by(StateName, Year) %>%
+  summarise(AvgHousePrice = mean(HousePrice, na.rm = TRUE), .groups = 'drop') # Calculating mean house price by year for each state
+
+
+
+#### save data ####
+write_csv(
+  x = clean_house_price,
+  file = "outputs/data/clean_5bplus_house_price.csv"
+)
+
+write_csv(
+  x = state_house_price_by_month,
+  file = "outputs/data/state_5bplus_house_price_by_month.csv"
+)
+
+write_csv(
+  x = state_house_price_by_year,
+  file = "outputs/data/state_5bplus_house_price_by_year.csv"
+)

### 2024-04-07 scripts/02-data_merging.R

@@ -1,0 +1,61 @@
+#### Preamble ####
+# Purpose: 
+# Author: Ziheng Zhong
+# Date: 25 March 2024
+# Contact: ziheng.zhong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: none
+
+
+
+#### Workspace setup ####
+library(dplyr)
+library(tidyverse)
+library(lubridate)
+
+
+
+#### load data ####
+state_1b_house_price_by_year <- read_csv("outputs/data/state_1b_house_price_by_year.csv")
+state_2b_house_price_by_year <- read_csv("outputs/data/state_2b_house_price_by_year.csv")
+state_3b_house_price_by_year <- read_csv("outputs/data/state_3b_house_price_by_year.csv")
+state_4b_house_price_by_year <- read_csv("outputs/data/state_4b_house_price_by_year.csv")
+state_5bplus_house_price_by_year <- read_csv("outputs/data/state_5bplus_house_price_by_year.csv")
+
+
+
+#### merge data ####
+state_1b_house_price_by_year <- state_1b_house_price_by_year %>% # Rename AvgHousePrice column in state_1b_house_price_by_year
+  rename(AvgHousePrice_1b = AvgHousePrice)
+
+merge_data <- left_join(state_1b_house_price_by_year, # Merge state_2b_house_price_by_year
+                            state_2b_house_price_by_year %>% 
+                              select(StateName, Year, AvgHousePrice) %>%
+                              rename(AvgHousePrice_2b = AvgHousePrice), 
+                            by = c("StateName", "Year"))
+
+merge_data <- left_join(merge_data, # Merging state_3b_house_price_by_year
+                        state_3b_house_price_by_year %>%
+                          select(StateName, Year, AvgHousePrice) %>%
+                          rename(AvgHousePrice_3b = AvgHousePrice),
+                        by = c("StateName", "Year"))
+
+merge_data <- left_join(merge_data, # Merging state_4b_house_price_by_year
+                        state_4b_house_price_by_year %>%
+                          select(StateName, Year, AvgHousePrice) %>%
+                          rename(AvgHousePrice_4b = AvgHousePrice),
+                        by = c("StateName", "Year"))
+
+merge_data <- left_join(merge_data, # Merging state_5bplus_house_price_by_year
+                        state_5bplus_house_price_by_year %>%
+                          select(StateName, Year, AvgHousePrice) %>%
[…]

### 2024-04-07 scripts/03-modeling.R

@@ -17,6 +17,6 @@ top_5_cause_data <- read_csv("outputs/data/top_5_cause_data.csv")
 
 
 
-#### poisson model ####
+#### xxx model ####
 
 

### 2024-04-07 scripts/01_00-data_cleaning.R
scripts updated

@@ -16,7 +16,7 @@ library(lubridate)
 
 
 #### load data ####
-raw_data <- read_csv("inputs/data/all_house_price.csv")
+raw_data <- read_csv("../../inputs/data/all_house_price.csv")
 
 
 

### 2024-04-07 scripts/02-data_merging.R

@@ -16,43 +16,38 @@ library(lubridate)
 
 
 #### load data ####
-state_1b_house_price_by_year <- read_csv("outputs/data/state_1b_house_price_by_year.csv")
-state_2b_house_price_by_year <- read_csv("outputs/data/state_2b_house_price_by_year.csv")
-state_3b_house_price_by_year <- read_csv("outputs/data/state_3b_house_price_by_year.csv")
-state_4b_house_price_by_year <- read_csv("outputs/data/state_4b_house_price_by_year.csv")
-state_5bplus_house_price_by_year <- read_csv("outputs/data/state_5bplus_house_price_by_year.csv")
+state_1b_house_price_by_year <- read_csv("../../outputs/data/state_1b_house_price_by_year.csv")
+state_2b_house_price_by_year <- read_csv("../../outputs/data/state_2b_house_price_by_year.csv")
+state_3b_house_price_by_year <- read_csv("../../outputs/data/state_3b_house_price_by_year.csv")
+state_4b_house_price_by_year <- read_csv("../../outputs/data/state_4b_house_price_by_year.csv")
+state_5bplus_house_price_by_year <- read_csv("../../outputs/data/state_5bplus_house_price_by_year.csv")
 
 
 
 #### merge data ####
-state_1b_house_price_by_year <- state_1b_house_price_by_year %>% # Rename AvgHousePrice column in state_1b_house_price_by_year
-  rename(AvgHousePrice_1b = AvgHousePrice)
+# Change the column "StateName" to "NumBedroom" and set the value to 1 for all rows
+state_1b_house_price_by_year <- state_1b_house_price_by_year %>%
+  mutate(NumBedroom = 1) %>%
+  select(-StateName) # This removes the original StateName column
 
-merge_data <- left_join(state_1b_house_price_by_year, # Merge state_2b_house_price_by_year
-                            state_2b_house_price_by_year %>% 
-                              select(StateName, Year, AvgHousePrice) %>%
-                              rename(AvgHousePrice_2b = AvgHousePrice), 
-                            by = c("StateName", "Year"))
+state_2b_house_price_by_year <- state_2b_house_price_by_year %>%
+  mutate(NumBedroom = 2) %>%
+  select(-StateName) # This removes the original StateName column
 
-merge_data <- left_join(merge_data, # Merging state_3b_house_price_by_year
-                        state_3b_house_price_by_year %>%
-                          select(StateName, Year, AvgHousePrice) %>%
[…]

### 2024-04-07 scripts/03-modeling.R

@@ -9,14 +9,40 @@
 
 
 #### Workspace setup ####
+library(mgcv)
+library(readr)
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+library(modelsummary)
 
 
 
 #### load data ####
-top_5_cause_data <- read_csv("outputs/data/top_5_cause_data.csv")
+house_prices <- read_csv("../../outputs/data/merged_house_price.csv")
 
 
 
-#### xxx model ####
+#### Multiple Regression Model ####
+model_multi <- lm(AvgHousePrice ~ Year + NumBedroom, data = house_prices)
+summary(model_multi)
+modelsummary(model_multi) # Model Results
 
 
+
+#### Polynomial Regression Model ####
+model_poly <- lm(AvgHousePrice ~ poly(Year, 2) + poly(NumBedroom, 2), data = house_prices)
+summary(model_poly)
+modelsummary(model_poly) # Model Results
+
+
+
+#### Generalized Additive Models ####
+num_unique_years <- length(unique(house_prices$Year)) # Check the number of unique values for each variable
+num_unique_bedrooms <- length(unique(house_prices$NumBedroom))
+
+gam_model <- gam(AvgHousePrice ~ s(Year, k=num_unique_years-1) + s(NumBedroom, k=num_unique_bedrooms-1), data = house_prices) # Specify k as one less than the number of unique values
+summary(gam_model)
+modelsummary(gam_model) # Model Results
+
+plot(gam_model, pages=1) # Visualization of fit

### 2024-04-07 scripts/04_00-heatmaps.R

@@ -1,0 +1,180 @@
+#### Preamble ####
+# Purpose: 
+# Author: Ziheng Zhong
+# Date: 25 March 2024
+# Contact: ziheng.zhong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: none
+
+
+
+#### Workspace setup ####
+library(maps)
+library(mgcv)
+library(ggmap)
+library(readr)
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+library(viridis)
+
+
+
+#### load data ####
+state_house_price_by_year <- read_csv("outputs/data/state_house_price_by_year.csv")
+state_1b_house_price_by_year <- read_csv("outputs/data/state_1b_house_price_by_year.csv")
+state_2b_house_price_by_year <- read_csv("outputs/data/state_2b_house_price_by_year.csv")
+state_3b_house_price_by_year <- read_csv("outputs/data/state_3b_house_price_by_year.csv")
+state_4b_house_price_by_year <- read_csv("outputs/data/state_4b_house_price_by_year.csv")
+state_5bplus_house_price_by_year <- read_csv("outputs/data/state_5bplus_house_price_by_year.csv")
+
+
+
+#### Preparations ####
+us_states_map <- map_data("state") # Getting map data for US states
+
+state_abbrev_to_full <- c( # Define the mapping from state abbreviations to full names
+  AL = "alabama", AK = "alaska", AZ = "arizona", AR = "arkansas", CA = "california",
+  CO = "colorado", CT = "connecticut", DE = "delaware", FL = "florida", GA = "georgia",
+  HI = "hawaii", ID = "idaho", IL = "illinois", IN = "indiana", IA = "iowa",
+  KS = "kansas", KY = "kentucky", LA = "louisiana", ME = "maine", MD = "maryland",
+  MA = "massachusetts", MI = "michigan", MN = "minnesota", MS = "mississippi",
+  MO = "missouri", MT = "montana", NE = "nebraska", NV = "nevada", NH = "new hampshire",
+  NJ = "new jersey", NM = "new mexico", NY = "new york", NC = "north carolina",
+  ND = "north dakota", OH = "ohio", OK = "oklahoma", OR = "oregon", PA = "pennsylvania",
+  RI = "rhode island", SC = "south carolina", SD = "south dakota", TN = "tennessee",
+  TX = "texas", UT = "utah", VT = "vermont", VA = "virginia", WA = "washington",
[…]

### 2024-04-07 scripts/04_01-charts.R

@@ -1,0 +1,78 @@
+#### Preamble ####
+# Purpose: 
+# Author: Ziheng Zhong
+# Date: 25 March 2024
+# Contact: ziheng.zhong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: none
+
+
+
+#### Workspace setup ####
+library(maps)
+library(mgcv)
+library(ggmap)
+library(readr)
+library(tidyr)
+library(dplyr)
+library(ggplot2)
+library(viridis)
+library(lubridate)
+
+
+
+#### load data ####
+state_house_price_by_month <- read_csv("outputs/data/state_house_price_by_month.csv")
+state_1b_house_price_by_month <- read_csv("outputs/data/state_1b_house_price_by_month.csv")
+state_2b_house_price_by_month <- read_csv("outputs/data/state_2b_house_price_by_month.csv")
+state_3b_house_price_by_month <- read_csv("outputs/data/state_3b_house_price_by_month.csv")
+state_4b_house_price_by_month <- read_csv("outputs/data/state_4b_house_price_by_month.csv")
+state_5bplus_house_price_by_month <- read_csv("outputs/data/state_5bplus_house_price_by_month.csv")
+
+
+
+#### 5 Types of Houses Price Trend by Month ####
+process_data <- function(data, label) { # Function to process each dataset
+  data %>% 
+    select(-StateName) %>%
+    summarise_all(mean, na.rm = TRUE) %>%
+    pivot_longer(everything(), names_to = "Date", values_to = "AverageHousePrice") %>%
+    mutate(Trend = label,
+           Date = as.Date(Date))
+}
+
+data_list <- list( # Processing all datasets
+  one_bedroom = state_1b_house_price_by_month,
+  two_bedroom = state_2b_house_price_by_month,
+  three_bedroom = state_3b_house_price_by_month,
+  four_bedroom = state_4b_house_price_by_month,
+  five_plus_bedroom = state_5bplus_house_price_by_month
+)
+
+processed_data <- lapply(names(data_list), function(name) process_data(data_list[[name]], name))
+
+combined_data <- bind_rows(processed_data) # Combining all processed data into one dataframe
+
+ggplot(combined_data, aes(x = Date, y = AverageHousePrice, color = Trend)) + # Plotting
+  geom_line() + 
+  labs(title = "Trend of Average House Price from 2000 to 2024 by House Type",
+       x = "Year", y = "Average House Price ($)") +
[…]

### 2024-04-07 outputs/paper/paper.qmd
paper updated

--- 2024-04-02 outputs/paper/paper.qmd
@@
 # read in data
-file_path_0 <- "../../inputs/data/cali_house_price.csv"
-file_path_1 <- "../../outputs/data/cali_house_price_clean.csv"
-file_path_2 <- "../../outputs/data/ocean_proximity_frequency.csv"
-file_path_3 <- "../../outputs/data/.csv"
+state_house_price_by_year <- read_csv("outputs/data/state_house_price_by_year.csv")
+state_house_price_by_month <- read_csv("outputs/data/state_house_price_by_month.csv")
 
-cali_house_price <- read_csv(file_path_0, show_col_types = FALSE)
-cali_house_price_clean<- read_csv(file_path_1, show_col_types = FALSE)
-ocean_proximity_frequency<- read_csv(file_path_2, show_col_types = FALSE)
+state_1b_house_price_by_year <- read_csv("outputs/data/state_1b_house_price_by_year.csv")
+state_2b_house_price_by_year <- read_csv("outputs/data/state_2b_house_price_by_year.csv")
+state_3b_house_price_by_year <- read_csv("outputs/data/state_3b_house_price_by_year.csv")
+state_4b_house_price_by_year <- read_csv("outputs/data/state_4b_house_price_by_year.csv")
+state_5bplus_house_price_by_year <- read_csv("outputs/data/state_5bplus_house_price_by_year.csv")
+
+state_1b_house_price_by_month <- read_csv("outputs/data/state_1b_house_price_by_month.csv")
+state_2b_house_price_by_month <- read_csv("outputs/data/state_2b_house_price_by_month.csv")
+state_3b_house_price_by_month <- read_csv("outputs/data/state_3b_house_price_by_month.csv")
+state_4b_house_price_by_month <- read_csv("outputs/data/state_4b_house_price_by_month.csv")
+state_5bplus_house_price_by_month <- read_csv("outputs/data/state_5bplus_house_price_by_month.csv")
 
@@
 
-longitude_summary <- summary(cali_house_price %>% select(longitude))
-latitude_summary <- summary(cali_house_price %>% select(latitude))
-housing_median_age_summary <- summary(cali_house_price %>% select(housing_median_age))
-total_rooms_summary <- summary(cali_house_price %>% select(total_rooms))
-total_bedrooms_summary <- summary(cali_house_price %>% select(total_bedrooms))
-population_summary <- summary(cali_house_price %>% select(population))
-households_summary <- summary(cali_house_price %>% select(households))
-median_income_summary <- summary(cali_house_price %>% select(median_income))
-median_house_value_summary <- summary(cali_house_price %>% select(median_house_value))
-
-kable(longitude_summary)
-kable(latitude_summary)
-kable(housing_median_age_summary)
-kable(total_rooms_summary)
-kable(total_bedrooms_summary)
-kable(population_summary)
-kable(households_summary)
-kable(median_income_summary)
-kable(median_house_value_summary)
-
 ```
@@
 
-null_values <- cali_house_price %>%
-  summarise_all(~sum(is.na(.))) %>%
-  gather(key = "Variables", value = "MissingValues")
-
-kable(null_values)
 ```
@@
 
-null_values <- cali_house_price_clean %>%
-  summarise_all(~sum(is.na(.))) %>%
-  gather(key = "Variables", value = "MissingValues")
-
-kable(null_values)
 ```
@@
 
-ggplot(ocean_proximity_frequency, aes(x = Variable, y = Frequency)) +
-  geom_bar(stat = "identity") +
-  theme_minimal() +
-  labs(title = "Frequency of Categories in Ocean Proximity",
-       x = "Ocean Proximity",
-       y = "Frequency") +
-  coord_flip()
-
 ```
@@
 #| fig-cap: TBD
+
+library(maps)
+library(dplyr)
+library(readr)
+library(viridis)
+library(ggplot2)
 

### 2024-04-08 scripts/00-data_simulation.R
purpose of all scripts added, date updated

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: 
+# Purpose: This code simulates and visualizes average house prices across US states using R, combining randomly generated price data with geographical mapping to create a heat map representation of the prices by state.
 # Author: Ziheng Zhong
-# Date: 25 March 2024
+# Date: 02 April 2024
 # Contact: ziheng.zhong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: none

### 2024-04-08 scripts/01_00-data_cleaning.R

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: 
+# Purpose: This code is aimed at cleaning and restructuring a comprehensive dataset of house prices, focusing on aggregating the data by state both on a monthly and yearly basis, followed by saving these cleaned and aggregated datasets for further analysis or visualization.
 # Author: Ziheng Zhong
-# Date: 25 March 2024
+# Date: 02 April 2024
 # Contact: ziheng.zhong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: none

### 2024-04-08 scripts/02-data_merging.R

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: 
+# Purpose: The code merges yearly house price data for different types of houses based on the number of bedrooms from various datasets, normalizes their structure by adding a uniform 'NumBedroom' column while removing the 'StateName' column, omits any missing values, and then saves the merged and cleaned dataset to a new CSV file.
 # Author: Ziheng Zhong
-# Date: 25 March 2024
+# Date: 02 April 2024
 # Contact: ziheng.zhong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: none

### 2024-04-08 scripts/03-modeling.R

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: 
+# Purpose: The code is designed to analyze house price data through multiple linear regression, polynomial regression, and generalized additive models (GAMs) in R, summarizing and visualizing the relationships between average house prices and features such as year and number of bedrooms.
 # Author: Ziheng Zhong
-# Date: 25 March 2024
+# Date: 02 April 2024
 # Contact: ziheng.zhong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: none

### 2024-04-08 scripts/04_00-heatmaps.R

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: 
+# Purpose: The code utilizes R libraries for data manipulation, visualization, and mapping to generate heat maps of average house prices by state in the US, categorized by the number of bedrooms, using datasets on yearly house prices.
 # Author: Ziheng Zhong
-# Date: 25 March 2024
+# Date: 02 April 2024
 # Contact: ziheng.zhong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: none
@@ -17,6 +17,7 @@ library(tidyr)
 library(dplyr)
 library(ggplot2)
 library(viridis)
+library(lubridate)
 
 
 

### 2024-04-08 scripts/04_01-charts.R

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: 
+# Purpose: The code is designed to process and visualize trends in average house prices by month for different house types using R, integrating data manipulation and visualization libraries to prepare and plot the data.
 # Author: Ziheng Zhong
-# Date: 25 March 2024
+# Date: 02 April 2024
 # Contact: ziheng.zhong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: none

### 2024-04-08 scripts/05-tests.R

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: The code defines a test suite for validating data integrity, including checks for positive calendar years, four-digit year format, non-empty cause entries, and non-negative death counts across multiple datasets within a research project.
+# Purpose: The code is designed for validating the integrity and constraints of house price data across various datasets using R's testing frameworks.
 # Author: Ziheng Zhong
-# Date: 25 March 2024
+# Date: 02 April 2024
 # Contact: ziheng.zhong@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: none
@@ -15,85 +15,234 @@ library(tidyverse)
 
 
 
-#### cali_house_price_clean testing ####
-test_that("cali_house_price_clean testing", {
-  cali_house_price_clean<- readr::read_csv("../../House_Price/outputs/data/cali_house_price_clean.csv")
+#### merged_house_price testing ####
+test_that("merged_house_price testing", {
+  merged_house_price<- readr::read_csv("../../House_Price/outputs/data/merged_house_price.csv")
   
-  # Test if 'longitude' values are within expected range
-  expect_true(all(cali_house_price_clean$longitude >= -180 & cali_house_price_clean$longitude <= 180), 
-              info = "Longitude values are within expected range")
+  # Test if 'Year' values are positive
+  expect_true(all(merged_house_price$Year > 0),
+              info = "Year values are positive")
   
-  # Test if 'latitude' values are within expected range
-  expect_true(all(cali_house_price_clean$latitude >= -90 & cali_house_price_clean$latitude <= 90),
-              info = "Latitude values are within expected range")
+  # Test if 'AvgHousePrice' values are positive
+  expect_true(all(merged_house_price$AvgHousePrice > 0),
+              info = "AvgHousePrice values are positive")
   
-  # Test if 'housing_median_age' values are positive
-  expect_true(all(cali_house_price_clean$housing_median_age > 0),
-              info = "Housing median age values are positive")
+  # Test if 'NumBedroom' values are positive
+  expect_true(all(merged_house_price$NumBedroom > 0),
+              info = "NumBedroom values are positive")
   
-  # Test if 'total_rooms' values are positive
-  expect_true(all(cali_house_price_clean$total_rooms > 0),
-              info = "Total rooms values are positive")
+  # Test if 'NumBedroom' values are within range
[…]

### 2024-04-08 outputs/paper/paper.qmd
references updated, appendix formated

@@
 # read in data
-state_house_price_by_year <- read_csv("outputs/data/state_house_price_by_year.csv")
-state_house_price_by_month <- read_csv("outputs/data/state_house_price_by_month.csv")
+house_prices <- read_csv("../../outputs/data/merged_house_price.csv")
 
-state_1b_house_price_by_year <- read_csv("outputs/data/state_1b_house_price_by_year.csv")
-state_2b_house_price_by_year <- read_csv("outputs/data/state_2b_house_price_by_year.csv")
-state_3b_house_price_by_year <- read_csv("outputs/data/state_3b_house_price_by_year.csv")
-state_4b_house_price_by_year <- read_csv("outputs/data/state_4b_house_price_by_year.csv")
-state_5bplus_house_price_by_year <- read_csv("outputs/data/state_5bplus_house_price_by_year.csv")
+state_house_price_by_year <- read_csv("../../outputs/data/state_house_price_by_year.csv")
+state_1b_house_price_by_year <- read_csv("../../outputs/data/state_1b_house_price_by_year.csv")
+state_2b_house_price_by_year <- read_csv("../../outputs/data/state_2b_house_price_by_year.csv")
+state_3b_house_price_by_year <- read_csv("../../outputs/data/state_3b_house_price_by_year.csv")
+state_4b_house_price_by_year <- read_csv("../../outputs/data/state_4b_house_price_by_year.csv")
+state_5bplus_house_price_by_year <- read_csv("../../outputs/data/state_5bplus_house_price_by_year.csv")
 
-state_1b_house_price_by_month <- read_csv("outputs/data/state_1b_house_price_by_month.csv")
-state_2b_house_price_by_month <- read_csv("outputs/data/state_2b_house_price_by_month.csv")
-state_3b_house_price_by_month <- read_csv("outputs/data/state_3b_house_price_by_month.csv")
-state_4b_house_price_by_month <- read_csv("outputs/data/state_4b_house_price_by_month.csv")
-state_5bplus_house_price_by_month <- read_csv("outputs/data/state_5bplus_house_price_by_month.csv")
+state_house_price_by_month <- read_csv("../../outputs/data/state_house_price_by_month.csv")
+state_1b_house_price_by_month <- read_csv("../../outputs/data/state_1b_house_price_by_month.csv")
+state_2b_house_price_by_month <- read_csv("../../outputs/data/state_2b_house_price_by_month.csv")
+state_3b_house_price_by_month <- read_csv("../../outputs/data/state_3b_house_price_by_month.csv")
+state_4b_house_price_by_month <- read_csv("../../outputs/data/state_4b_house_price_by_month.csv")
+state_5bplus_house_price_by_month <- read_csv("../../outputs/data/state_5bplus_house_price_by_month.csv")
 
@@
 Data used in this paper was cleaned, processed and tested with the programming language R [@citeR].
-Also with support of additional packages in R: `tidyverse` [@citeTidyverse], `ggplot2` [@citeGgplot], `janitor` [@citeJanitor], `readr` [@citeReadr], `knitr` [@citeKnitr], `rstanarm` [@citeRstanarm], `modelsummary` [@citeModelsummary], `tidybayes` [@citeTidybayes], `loo` [@citeLoo], `testthat` [@citeTestthat], `KableExtra` [@citeKableEx].
+Also with support of additional packages in R: `tidyverse` [@citeTidyverse], `ggplot2` [@citeGgplot], `janitor` [@citeJanitor], `readr` [@citeReadr], `knitr` [@citeKnitr], `modelsummary` [@citeModelsummary], `testthat` [@citeTestthat], `KableExtra` [@citeKableEx], `viridis` [@citeViridis], `lubridate` [@citeLubridate], `maps` [@citeMaps], `mgcv` [@citeMgcv].
 
@@
 
+\appendix
+
+# Appendix
+
+## Datasheet
+
+**Motivation**
+
+**Composition**
+
+**Collection process**
+
+**Preprocessing/cleaning/labeling**
+
+**Uses**
+
+**Distribution**
+
+**Maintenance**
+
 \newpage

### 2024-04-08 outputs/paper/paper.qmd
charts, model results added to paper

@@
 # load packages
-packages <- c("tidyverse", "boot", "broom.mixed", "collapse", "dataverse", "gutenbergr", "janitor", "knitr", "marginaleffects", "modelsummary", "rstanarm", "tidybayes", "loo", "kableExtra")
+packages <- c("tidyverse", "ggplot2", "janitor", "readr", "knitr", "modelsummary", "testthat", "kableExtra", "viridis", "lubridate", "maps", "mgcv")
 
@@
 #| include: false
+# Models Setup
+
+#### Multiple Regression Model ####
+model_multi <- lm(AvgHousePrice ~ Year + NumBedroom, data = house_prices)
+
+#### Polynomial Regression Model ####
+model_poly <- lm(AvgHousePrice ~ poly(Year, 2) + poly(NumBedroom, 2), data = house_prices)
+
+#### Generalized Additive Models ####
+num_unique_years <- length(unique(house_prices$Year))
+num_unique_bedrooms <- length(unique(house_prices$NumBedroom))
+
+gam_model <- gam(AvgHousePrice ~ s(Year, k=num_unique_years-1) + s(NumBedroom, k=num_unique_bedrooms-1), data = house_prices)
 
@@
 #| fig-pos: 'h'
-#| label: fig-ocean_proximity_frequency
-#| fig-cap: Frequency of Ocean Proximity
+#| label: fig-trend_overall
+#| fig-cap: Trend of Average House Price from 2000 to 2024
+
+state_all <- state_house_price_by_month %>% select(-StateName) %>%  # Processing datasets
+  summarise_all(mean, na.rm = TRUE) %>%
+  pivot_longer(everything(), names_to = "Date", values_to = "AverageHousePrice")
+
+state_all$Date <- as.Date(state_all$Date) # Convert Date from character to Date type
+
+ggplot(state_all, aes(x = Date, y = AverageHousePrice)) + # Plotting
+  geom_line(color = "blue") + 
+  labs(x = "Year", y = "Average House Price ($)") + 
+  theme_minimal() + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
@@
 #| fig-pos: 'h'
-#| label: fig-TBD_1
-#| fig-cap: TBD
+#| label: fig-trend_by_type
+#| fig-cap: Trend of Average House Price from 2000 to 2024 by House Type
 
-library(maps)
-library(dplyr)
-library(readr)
-library(viridis)
-library(ggplot2)
+process_data <- function(data, label) { # Function to process each dataset
+  data %>% 
+    select(-StateName) %>%
+    summarise_all(mean, na.rm = TRUE) %>%
+    pivot_longer(everything(), names_to = "Date", values_to = "AverageHousePrice") %>%
+    mutate(Trend = label,
+           Date = as.Date(Date))
+}
+
+data_list <- list( # Processing all datasets
+  one_bedroom = state_1b_house_price_by_month,
+  two_bedroom = state_2b_house_price_by_month,
+  three_bedroom = state_3b_house_price_by_month,
+  four_bedroom = state_4b_house_price_by_month,
+  five_plus_bedroom = state_5bplus_house_price_by_month
+)
+
+processed_data <- lapply(names(data_list), function(name) process_data(data_list[[name]], name))
+
+combined_data <- bind_rows(processed_data) # Combining all processed data into one dataframe
+
+ggplot(combined_data, aes(x = Date, y = AverageHousePrice, color = Trend)) + # Plotting
+  geom_line() + 
+  labs(x = "Year", y = "Average House Price ($)") +
+  theme_minimal() + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_color_viridis_d(name = "House Type")
+
+```
+
+## Heat Maps
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-pos: 'h'
+#| label: fig-hm_overall
+#| fig-cap: Average Price by State in the US for All House Types
+
+us_states_map <- map_data("state") # Getting map data for US states
+
+state_abbrev_to_full <- c( # Define the mapping from state abbreviations to full names
+  AL = "alabama", AK = "alaska", AZ = "arizona", AR = "arkansas", CA = "california",
+  CO = "colorado", CT = "connecticut", DE = "delaware", FL = "florida", GA = "georgia",
+  HI = "hawaii", ID = "idaho", IL = "illinois", IN = "indiana", IA = "iowa",
+  KS = "kansas", KY = "kentucky", LA = "louisiana", ME = "maine", MD = "maryland",
+  MA = "massachusetts", MI = "michigan", MN = "minnesota", MS = "mississippi",
+  MO = "missouri", MT = "montana", NE = "nebraska", NV = "nevada", NH = "new hampshire",
+  NJ = "new jersey", NM = "new mexico", NY = "new york", NC = "north carolina",
+  ND = "north dakota", OH = "ohio", OK = "oklahoma", OR = "oregon", PA = "pennsylvania",
+  RI = "rhode island", SC = "south carolina", SD = "south dakota", TN = "tennessee",
+  TX = "texas", UT = "utah", VT = "vermont", VA = "virginia", WA = "washington",
+  WV = "west virginia", WI = "wisconsin", WY = "wyoming", DC = "district of columbia"
+)
+
+overall_hm_data <- na.omit(state_house_price_by_year) # Delete all NAs
+overall_hm_data$StateName <- tolower(state_abbrev_to_full[overall_hm_data$StateName]) # Convert StateName from abbreviation to full name
+
+overall_hm_data <- overall_hm_data %>% # Taking average for house price based on states
+  group_by(StateName) %>%
+  summarise(AvgHousePrice = mean(AvgHousePrice, na.rm = TRUE))
+
+overall_hm_map_merged_data <- merge(us_states_map, overall_hm_data, by.x = "region", by.y = "StateName") # Merge data ready for chart
+
+ggplot() + # Generating heat map
+  geom_polygon(data = overall_hm_map_merged_data, aes(x = long, y = lat, group = group, fill = AvgHousePrice), color = "white") +
+  coord_fixed(1.3) +
+  scale_fill_viridis(name = "Avg House Price", option = "plasma") +
+  theme_minimal() +
+  theme(legend.position = "right")
+
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-pos: 'h'
+#| label: fig-hm_by_type
+#| fig-cap: Average Price by State in the US for Different House Types
+#| fig-subcap: ["one bedroom","two bedrooms","three bedrooms","four bedrooms","five plus bedrooms"]
+#| layout-ncol: 2
+#| layout-nrow: 3
+
+us_states_map <- map_data("state") # Getting map data for US states
+
+state_abbrev_to_full <- c( # Define the mapping from state abbreviations to full names
+  AL = "alabama", AK = "alaska", AZ = "arizona", AR = "arkansas", CA = "california",
+  CO = "colorado", CT = "connecticut", DE = "delaware", FL = "florida", GA = "georgia",
+  HI = "hawaii", ID = "idaho", IL = "illinois", IN = "indiana", IA = "iowa",
+  KS = "kansas", KY = "kentucky", LA = "louisiana", ME = "maine", MD = "maryland",
+  MA = "massachusetts", MI = "michigan", MN = "minnesota", MS = "mississippi",
+  MO = "missouri", MT = "montana", NE = "nebraska", NV = "nevada", NH = "new hampshire",
+  NJ = "new jersey", NM = "new mexico", NY = "new york", NC = "north carolina",
+  ND = "north dakota", OH = "ohio", OK = "oklahoma", OR = "oregon", PA = "pennsylvania",
+  RI = "rhode island", SC = "south carolina", SD = "south dakota", TN = "tennessee",
+  TX = "texas", UT = "utah", VT = "vermont", VA = "virginia", WA = "washington",
+  WV = "west virginia", WI = "wisconsin", WY = "wyoming", DC = "district of columbia"
+)
+
+#### US States 1b House Price Heat Map ####
+oneb_hm_data <- na.omit(state_1b_house_price_by_year) # Delete all NAs
+oneb_hm_data$StateName <- tolower(state_abbrev_to_full[oneb_hm_data$StateName]) # Convert StateName from abbreviation to full name
+
+oneb_hm_data <- oneb_hm_data %>% # Taking average for house price based on states
+  group_by(StateName) %>%
+  summarise(AvgHousePrice = mean(AvgHousePrice, na.rm = TRUE))
+
+oneb_hm_map_merged_data <- merge(us_states_map, oneb_hm_data, by.x = "region", by.y = "StateName") # Merge data ready for chart
+
+ggplot() + # Generating heat map
+  geom_polygon(data = oneb_hm_map_merged_data, aes(x = long, y = lat, group = group, fill = AvgHousePrice), color = "white") +
+  coord_fixed(1.3) +
+  scale_fill_viridis(name = "Avg House Price", option = "plasma") +
+  theme_minimal() +
+  theme(legend.position = "right")
+
+#### US States 2b House Price Heat Map ####
+twob_hm_data <- na.omit(state_2b_house_price_by_year) # Delete all NAs
+twob_hm_data$StateName <- tolower(state_abbrev_to_full[twob_hm_data$StateName]) # Convert StateName from abbreviation to full name
+
+twob_hm_data <- twob_hm_data %>% # Taking average for house price based on states
+  group_by(StateName) %>%
+  summarise(AvgHousePrice = mean(AvgHousePrice, na.rm = TRUE))
+
+twob_hm_map_merged_data <- merge(us_states_map, twob_hm_data, by.x = "region", by.y = "StateName") # Merge data ready for chart
+
+ggplot() + # Generating heat map
+  geom_polygon(data = twob_hm_map_merged_data, aes(x = long, y = lat, group = group, fill = AvgHousePrice), color = "white") +
+  coord_fixed(1.3) +
+  scale_fill_viridis(name = "Avg House Price", option = "plasma") +
+  theme_minimal() +
+  theme(legend.position = "right")
+
+#### US States 3b House Price Heat Map ####
+threeb_hm_data <- na.omit(state_3b_house_price_by_year) # Delete all NAs
+threeb_hm_data$StateName <- tolower(state_abbrev_to_full[threeb_hm_data$StateName]) # Convert StateName from abbreviation to full name
+
+threeb_hm_data <- threeb_hm_data %>% # Taking average for house price based on states
+  group_by(StateName) %>%
+  summarise(AvgHousePrice = mean(AvgHousePrice, na.rm = TRUE))
+
+threeb_hm_map_merged_data <- merge(us_states_map, threeb_hm_data, by.x = "region", by.y = "StateName") # Merge data ready for chart
+
+ggplot() + # Generating heat map
+  geom_polygon(data = threeb_hm_map_merged_data, aes(x = long, y = lat, group = group, fill = AvgHousePrice), color = "white") +
+  coord_fixed(1.3) +
+  scale_fill_viridis(name = "Avg House Price", option = "plasma") +
+  theme_minimal() +
+  theme(legend.position = "right")
+
+#### US States 4b House Price Heat Map ####
+fourb_hm_data <- na.omit(state_4b_house_price_by_year) # Delete all NAs
+fourb_hm_data$StateName <- tolower(state_abbrev_to_full[fourb_hm_data$StateName]) # Convert StateName from abbreviation to full name
+
+fourb_hm_data <- fourb_hm_data %>% # Taking average for house price based on states
+  group_by(StateName) %>%
+  summarise(AvgHousePrice = mean(AvgHousePrice, na.rm = TRUE))
+
+fourb_hm_map_merged_data <- merge(us_states_map, fourb_hm_data, by.x = "region", by.y = "StateName") # Merge data ready for chart
+
+ggplot() + # Generating heat map
+  geom_polygon(data = fourb_hm_map_merged_data, aes(x = long, y = lat, group = group, fill = AvgHousePrice), color = "white") +
+  coord_fixed(1.3) +
+  scale_fill_viridis(name = "Avg House Price", option = "plasma") +
+  theme_minimal() +
+  theme(legend.position = "right")
+
+#### US States 5b+ House Price Heat Map ####
+fiveb_hm_data <- na.omit(state_5bplus_house_price_by_year) # Delete all NAs
+fiveb_hm_data$StateName <- tolower(state_abbrev_to_full[fiveb_hm_data$StateName]) # Convert StateName from abbreviation to full name
+
+fiveb_hm_data <- fiveb_hm_data %>% # Taking average for house price based on states
+  group_by(StateName) %>%
+  summarise(AvgHousePrice = mean(AvgHousePrice, na.rm = TRUE))
+
+fiveb_hm_map_merged_data <- merge(us_states_map, fiveb_hm_data, by.x = "region", by.y = "StateName") # Merge data ready for chart
+
+ggplot() + # Generating heat map
+  geom_polygon(data = fiveb_hm_map_merged_data, aes(x = long, y = lat, group = group, fill = AvgHousePrice), color = "white") +
+  coord_fixed(1.3) +
+  scale_fill_viridis(name = "Avg House Price", option = "plasma") +
+  theme_minimal() +
+  theme(legend.position = "right")
 
@@
 #| fig-pos: 'h'
-#| label: tbl-TBD_2
-#| tbl-cap: TBD
+#| label: tbl-linear_model_result
+#| tbl-cap: Modeling Results for Linear Models
+
+modelsummary(
+  list(
+    "Multiple Regression" = model_multi,
+    "Polynomial Regression" = model_poly
+  )
+)
 
@@
 #| fig-pos: 'h'
-#| label: fig-TBD_3
-#| fig-cap: TBD
+#| label: tbl-nonlinear_model_result
+#| fig-cap: Modeling Results for Non-linear Models
+
+modelsummary(
+  list(
+    "GAM Regression" = gam_model
+  )
+)
 
@@
 #| fig-pos: 'h'
-#| label: tbl-TBD_4
-#| tbl-cap: TBD
+#| label: fig-GAM
+#| tbl-cap: GAM Regression Fit Result
+
+plot(gam_model, pages=1)
 

### 2024-04-09 outputs/paper/paper.qmd
appendix data sheet added

@@
 
+1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
+    - The datasets are created to support a wide range of real estate analysis and research by Zillow Research group.
+These datasets typically serve purposes such as market trend analysis, price prediction, housing supply studies, and economic impact assessments.
+They are designed to fill the gap for comprehensive, accurate, and accessible real estate data for researchers, policymakers, and the general public interested in the housing market dynamics.
+
+2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
+    - The datasets available on the research page of Zillow, a leading real estate and rental marketplace, are typically created by Zillow's own economic research team.
+This team focuses on providing insights into the housing market and economic trends.
+
+3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
+    - The datasets on Zillow's research page are generally created and funded internally by Zillow Group, Inc.
+itself, without specific external grants.
+As a commercial entity with a vested interest in real estate markets, Zillow utilizes its resources to compile and analyze these datasets for public and internal use.
+
+4. *Any other comments?*
+    - None.
+
 **Composition**
+
+1.
+*What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)?
+Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)?
+Please provide a description.*
+	- In the dataset used for this paper, each instance representing a geographic region.
+These regions mainly include the country as a whole (United States), metropolitan statistical areas (e.g., New York, NY; Los Angeles, CA).
+The dataset contains a series of monthly average house prices, spanning from January 2000 to February 2024, as shown by the date-formatted columns.
+The data does not mix different types of instances (like movies, users, and ratings or people and interactions between them) but focuses solely on regional housing price data over time.
+Each row includes identifiers and names for the regions, the type of region (such as 'country' or 'msa' for metropolitan statistical area), and state names for regions within specific states.
+
+2. *How many instances are there in total (of each type, if appropriate)?*
+	- 50 States
+	- 895 Regions
+	- 290 Dates
+
+3.
+*Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set?
+If the dataset is a sample, then what is the larger set?
+Is the sample representative of the larger set (for example, geographic coverage)?
+If so, please describe how this representativeness was validated/verified.
+If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
+	- The dataset is likely a sample from a larger set of all possible geographic areas in the United States for which housing data could be collected.
+This larger set would include a comprehensive collection of all geographic regions in the United States, encompassing every city, town, rural area, and Metropolitan Statistical Area (MSA) with available housing market data.
+A clear example indicating that this is just a sample is the absence of data for Hawaii.
+Regarding representativeness, this data is representative, mainly because Zillow is one of the leading real estate trading companies in the United States, and its data covers almost all cities.
+Additionally, there is no special focus on regions that provide the most insights into the national housing market trends.
+
+4. *What data does each instance consist of? "Raw" data or features? In either case, please provide a description.*
+	- RegionID: A unique identifier for each geographic region.
+  - SizeRank: A rank based on the size or significance of the region, presumably in terms of population or housing market activity.
+  - RegionName: The name of the region, which can be a country (like the United States), a metropolitan statistical area (MSA), or possibly other region types.
+  - RegionType: The type of region, such as 'country' or 'msa', indicating the scope or level of the geographic area.
+  - StateName: The name of the state in which the region is located, applicable to regions within specific states.
+  - Monthly Average House Prices: A series of columns representing the average house prices for each month, spanning from January 2000 to February 2024.
+These columns are named by date (e.g., '2000-01-31', '2000-02-29', etc.), with each column representing the average housing price in that region for the given month.
+
+5. *Is there a label or target associated with each instance? If so, please provide a description.*
+	- RegionID: Unique ID associated with each region in RegionName.
+
+6.
+*Is any information missing from individual instances?
+If so, please provide a description, explaining why this information is missing (for example, because it was unavailable).
+This does not include intentionally removed information, but might include, for example, redacted text.*
+	- Information such as the latitude and longitude for each region is not included.
+The absence of geographic coordinates (latitude and longitude) means that while the dataset provides a comprehensive temporal view of housing price trends across different regions, it does not directly offer spatial data that would allow for mapping or spatial analysis of these trends.
+The reason for missing information is because of the dataset's primary focus on economic trends and prioritizes time-series analysis of housing prices.
+
+7. *Are relationships between individual instances made explicit? If so, please describe how these relationships are made explicit.*
+	- The relationships between individual instances (regions) are not explicitly defined in terms of direct interactions or connections between them.
+The dataset primarily focuses on time-series data for housing prices within various geographic regions without detailing explicit relationships like proximity, hierarchical structures (e.g., how states are composed of multiple cities), or economic interdependencies among these regions.
+
+8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
+	- There are detailed datasets for houses with specific number of bedrooms. Thus no need to split anything.
+
+9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
+	- None.
+
+10.
+*Is the dataset self-contained, or does it link to or otherwise rely on external resources?
+If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer?
+Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
+	- This is a self-contained dataset.
+
+11.
+*Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)?
+If so, please provide a description.*
+	- None.
+
+12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
+	- None.
+
+13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified.*
+	- The data are labled with StateName and RegionName. Therefore, data from each state or region coule be view as a subpopulations of this dataset.
+
+14.
+*Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset?
+If so, please describe how.*
+	- None.
+
+15.
+*Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)?
+If so, please provide a description.*
+	- None.
+
+16. *Any other comments?*
+	- None.
 
@@
 
+1.
+*How was the data associated with each instance acquired?
+Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)?
+If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified?
+If so, please describe how.
+How were these mechanisms or procedures validated?*
+	- The data in the Zillow dataset, specifically the housing prices for various regions, is primarily derived from Zillow's own listings and transactions data, along with public records and assessments.
+Here's a breakdown of the data acquisition methods:
+	  - Direct Observation and Public Records: Zillow aggregates housing price data from several sources, including direct listings on their platform, real estate transactions, and public property records such as sales and assessments.
+This means that much of the data is directly observable or comes from official records, making it a robust and reliable source of housing market information.
+	  - Zestimate: Some of the housing price data, particularly for times or areas where direct transaction data may be sparse, could be supplemented by Zillow's Zestimate® home values.
+The Zestimate is an estimated market value calculated using proprietary models that analyze public and user-submitted data.
+This would fall under data indirectly inferred/derived from other data.
+	  - Validation and Verification: For directly observed or recorded data (listings, transactions, public records), the validity comes from the data's official or commercial nature.
+These are factual records of housing sales and listings.
+For data like the Zestimate, Zillow continuously updates and refines its models based on new data, market trends, and feedback.
+The accuracy of Zestimates is evaluated by comparing estimated values with actual sale prices when they become available, and Zillow publishes accuracy metrics for its Zestimates, providing a form of validation.
+
+2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)?*
+	- Mainly direct obervation (using data from their platform, as mentioned in previous question).
+
+3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
+	- Not mentioned by data provider.
+
+4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
+	- Zillow’s Platform and Automated Systems: Much of the data collection is carried out through Zillow's own technological infrastructure, which aggregates and processes listings, sales data, and property information from across the United States.
+This process is automated, involving sophisticated software systems designed to handle large volumes of data.
+	- Real Estate Professionals: Realtors and other real estate professionals often use Zillow to list properties.
+While their primary motivation is to market properties to potential buyers, their contributions add to the dataset's comprehensiveness.
+Compensation for these professionals comes through the real estate transactions facilitated by their listings, not directly from Zillow for the data per se.
+
+5.
+*Over what timeframe was the data collected?
+Does this timeframe match the creation timeframe of the data associated with the instances?
+If not, please describe the timeframe in which the data associated with the instances was created.*
+	- The dataset provided spans from January 2000 to February 2024, indicating that the data collection covers this specific timeframe.
+This period reflects the creation timeframe of the data associated with each instance, which means each instance's housing price data was collected or estimated for these specific monthly intervals.
+
+6.
+*Were any ethical review processes conducted?
+If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
+	- For datasets like the one provided by Zillow, which compiles housing market data across various regions in the United States, specific ethical review processes might not be as prominently documented or required in the same manner as they would be for research involving human subjects directly.
+However, Zillow, like many data-providing entities, operates within a framework of legal and ethical considerations, especially regarding privacy, data accuracy, and transparency.
+
+7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
+	- Data was obtained from Zillow.
+
+8.
+*Were the individuals in question notified about the data collection?
+If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
+	- This is a publicly available information on housing prices and transactions.
+
+9.
+*Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted?
+If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
+	- None
+
+10. *Any other comments?*
+	- None.
+
 **Preprocessing/cleaning/labeling**
+
+1.
+*Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)?
+If so, please provide a description.
+If not, you may skip the remaining questions in this section.*
+	- Cleaning:
+	  - Columns (SizeRank, RegionID, RegionType, RegionName) are removed from the dataset, likely because they are not necessary for the subsequent analysis focused on state-level price trends.
+	- Further Processing:
+	  - Monthly Aggregation: The data is grouped by StateName, and the mean house price for each state is calculated for each month.
+This step involves aggregating all numeric columns (assumed to be monthly house price data) and computing their means, excluding missing values.
+	  - Yearly Aggregation: The monthly data is then transformed to calculate the average house price by year for each state.
+This involves converting month columns to a long format, extracting the year from each date, and then calculating the yearly mean house price for each state.
+
+2.
+*Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)?
+If so, please provide a link or other access point to the "raw" data.*
+	- The raw data was saved in the project folder [https://github.com/iJustinn/House_Price.git], specific location identified in the README section.
+
+3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
+	- Programming langugae R [https://cran.r-project.org/] in the IDE RStudio [https://www.rstudio.com/products/rstudio/download/].
+
+4. *Any other comments?*
+	- None.
 
@@
 
+1. *Has the dataset been used for any tasks already? If so, please provide a description.*
+	- Building modeling and creating charts for this project.
+
+2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
+	- GitHub Link[https://github.com/iJustinn/House_Price.git]
+
+3. *What (other) tasks could the dataset be used for?*
+	- None.
+
+4.
+*Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses?
+For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)?
+If so, please provide a description.
+Is there anything a dataset consumer could do to mitigate these risks or harms?*
+	- The processed dataset, centered around housing prices across various states and time periods, is limited to only use for similar tasks of this project. Hence it has impact its future use.
+
+5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
+	- Legal or Regulatory Compliance: Using aggregated and possibly anonymized data for purposes requiring detailed, verified information — such as legal compliance, zoning decisions, or adherence to housing regulations — might not be appropriate.
+Such applications typically require specific, case-by-case data rather than broad averages or trends.
+	- Short-term Investment Decisions: While historical data can highlight trends, using it for short-term real estate investment decisions without considering current market dynamics, economic indicators, and local factors could lead to misguided decisions.
+The dataset likely does not capture rapid market changes or short-term fluctuations.
+
+6. *Any other comments?*
+	- None.
+
 **Distribution**
 
+1.
+*Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created?
+If so, please provide a description.*
+	- None.
+
+2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
+	- GitHub, as mentioned in Uses section.
+
+3. *When will the dataset be distributed?*
+	- April, 2024. The time when this project is uploaded onto GitHub.
+
+4.
+*Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)?
+If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
+	- None.
+
+5.
+*Have any third parties imposed IP-based or other restrictions on the data associated with the instances?
+If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
+	- None.
+
+6.
+*Do any export controls or other regulatory restrictions apply to the dataset or to individual instances?
+If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
+	- None.
+
+7. *Any other comments?*
+	- None.
+
 **Maintenance**
+
+1. *Who will be supporting/hosting/maintaining the dataset?*
+	- The dataset will be held at server of GitHub, regulated by the owner of this project.
+
+2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
+	- The owner can be reached via GitHub account.
+
+3. *Is there an erratum? If so, please provide a link or other access point.*
+	- None.
+
+4.
+*Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)?
+If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
+	- Nothing about this project will updated after it is done.
+
+5.
+*If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)?
+If so, please describe these limits and explain how they will be enforced.*
+	- None.
+
+6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
+	- Only datasets on the GitHub will be hosted.
+
+7.
+*If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so?
+If so, please provide a description.
+Will these contributions be validated/verified?
+If so, please describe how.
+If not, why not?
+Is there a process for communicating/distributing these contributions to dataset consumers?
+If so, please provide a description.*
+	-  People other than the owner can extend the dataset via collaboration feature of GitHub, or directly contact owner for further work.
+
+8. *Any other comments?*
+	- None.
 

### 2024-04-14 outputs/paper/paper.qmd
introduction and source added

@@
 
+This paper will focus on exploring trends in U.S.
+house prices, a pertinent issue as real estate markets worldwide have experienced significant increases.
+Cities like Toronto, Canada, and Beijing, China have witnessed substantial escalations in housing costs, reflecting a global phenomenon.
+This study shifts its focus to the United States, one of the most economically stable and developed countries in the world, to determine if similar trends persist in its real estate market.
+By examining the U.S.
+housing market, this research aims to uncover the principal factors influencing house prices.
+
+The U.S.
+real estate market offers a unique case study due to its diverse economic landscapes and varied housing markets across states and metropolitan areas.
+This analysis will consider variables such as location, number of bedrooms, and other key factors that are believed to significantly impact house pricing dynamics.
+Understanding these factors is crucial, as the findings will provide insights into how policy adjustments and economic changes can affect housing affordability.
+Ultimately, this research aims to serve as a guide for policymakers, investors, and the public, helping them make informed decisions regarding housing investments and urban planning.
+
+Following this introduction, @sec-data (Methodology), outlines key reasons of the dataset choice, main processing techniques employed, emphasizing transparency and replicability.
+@sec-res (Results) presents the findings, specifically focusing on the dynamics of house prices across various U.S.
+states and metropolitan areas.
+@sec-dis (Discussion), analyzes these findings from the perspectives outlined earlier, integrating economic, policy, and regional variables.
+Finally, @sec-con (Conclusion), summarizes the key insights and implications of this research, offering recommendations for policymakers and stakeholders involved in the housing market.
+
 # Data {#sec-data}
@@
 ## Source
+
+Our study focuse on U.S.
+house price trends using the comprehensive datasets available from Zillow Research  [@citeZillowdata].
+This data repository offers an extensive array of real estate statistics, capturing the dynamics of house prices across various regions in the United States.
+As a cornerstone of transparency and public engagement, Zillow Research provides open access to these crucial real estate data, which are instrumental in understanding housing market behaviors.
+
+The dataset includes detailed information on house prices over several years, with breakdowns by state, city, and sometimes neighborhood levels, from median list prices to Zillow Home Value Index (ZHVI).
+This granularity allows for a nuanced analysis of house price trends and the identification of significant market dynamics, guiding potential real estate investments and policy formulations.
+
+We selected this dataset for its reliability, depth, and alignment with our study’s goals.
+As a trusted source in the real estate market, it provides accurate and timely data crucial for our analysis.
+Its comprehensive geographic and property-specific details allow for an in-depth examination of factors influencing house prices in the U.S.
+Additionally, the dataset’s structure supports efficient analysis and modeling of real estate trends.
+More details about this dataset can be found in the datasheet in @sec-app (Appendix).
 
@@
 
-## Demographic Shifts
+## 1
 
-## Health-related Behaviors
+## 2
 
-## Government Policies
+## 3
 
-## Environmental Changes
+## 4
 
@@
 
-# Appendix
+# Appendix {#sec-app}
 

### 2024-04-14 scripts/04_01-charts.R
scripts updated

@@ -41,25 +41,24 @@ process_data <- function(data, label) { # Function to process each dataset
            Date = as.Date(Date))
 }
 
-data_list <- list( # Processing all datasets
-  one_bedroom = state_1b_house_price_by_month,
-  two_bedroom = state_2b_house_price_by_month,
-  three_bedroom = state_3b_house_price_by_month,
-  four_bedroom = state_4b_house_price_by_month,
-  five_plus_bedroom = state_5bplus_house_price_by_month
+data_list <- list(
+  `1B` = state_1b_house_price_by_month,
+  `2B` = state_2b_house_price_by_month,
+  `3B` = state_3b_house_price_by_month,
+  `4B` = state_4b_house_price_by_month,
+  `5B+` = state_5bplus_house_price_by_month
 )
 
 processed_data <- lapply(names(data_list), function(name) process_data(data_list[[name]], name))
 
-combined_data <- bind_rows(processed_data) # Combining all processed data into one dataframe
+combined_data <- bind_rows(processed_data)
 
-ggplot(combined_data, aes(x = Date, y = AverageHousePrice, color = Trend)) + # Plotting
+ggplot(combined_data, aes(x = Date, y = AverageHousePrice, color = Trend)) +
   geom_line() + 
-  labs(title = "Trend of Average House Price from 2000 to 2024 by House Type",
-       x = "Year", y = "Average House Price ($)") +
+  labs(x = "Year", y = "Average House Price ($)") +
   theme_minimal() + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  scale_color_viridis_d(name = "House Type")
+  scale_color_viridis_d(name = "Number of Bedrooms")
 
 
 
@@ -70,8 +69,23 @@ state_all <- state_house_price_by_month %>% select(-StateName) %>%  # Processing
 
 state_all$Date <- as.Date(state_all$Date) # Convert Date from character to Date type
 
+extrema <- state_all %>%
+  mutate(is_peak = ifelse(lag(AverageHousePrice, 1, default = first(AverageHousePrice)) < AverageHousePrice & 
+                            lead(AverageHousePrice, 1, default = last(AverageHousePrice)) < AverageHousePrice,
+                          TRUE, FALSE),
+         is_trough = ifelse(lag(AverageHousePrice, 1, default = first(AverageHousePrice)) > AverageHousePrice & 
[…]

### 2024-04-14 outputs/paper/paper.qmd
paper updated till modeling section

@@
 # read in data
+all_house_price <- read_csv("../../inputs/data/all_house_price.csv")
 house_prices <- read_csv("../../outputs/data/merged_house_price.csv")
@@
 
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-pos: 'h'
-#| label: tbl-sum
-#| tbl-cap: Summary statistics of the California housing dataset
-#| layout-ncol: 3
-#| layout-nrow: 3
-
-```
+Our analysis commenced with the essential task of refining the raw dataset downloaded from Zillow.
+The initial phase involved understanding the dataset’s structure.
+As the @tbl-raw_data showing, this dataset typically consists of multiple attributes such as Region ID, Region Type, location (states), year of transaction, house price, etc.
+Each record represents a unique property, providing a comprehensive snapshot of its characteristics.
 
@@
 #| fig-pos: 'h'
-#| label: tbl-null-raw
-#| tbl-cap: Count of missing values for each variable
+#| label: tbl-raw_data
+#| tbl-cap: Raw Data Preview
+
+kable(head(all_house_price)[, 1:7])
 
 ```
+
+Then, we proceeded to clean and process the data as needed, setting the stage for deeper analysis.
+Our first step involved preprocessing the raw dataset from Zillow, which required several critical adjustments to ensure data quality and usability.
+Initially, we removed the first row of the dataset, which represented national averages rather than state-specific data, to focus our analysis on regional trends.
+This step was crucial as the inclusion of national data could skew the results when examining state-level dynamics.
+
+Next, we streamlined the dataset by removing less relevant columns such as 'SizeRank', 'RegionID', 'RegionType', and 'RegionName'.
+These columns were extraneous for our analysis, which aimed to concentrate on price fluctuations and trends across different states.
+By simplifying the dataset to whcih showing in @tbl-clean_data, we facilitated more focused and faster computations in subsequent steps.
 
@@
 #| fig-pos: 'h'
-#| label: tbl-null-clean
-#| tbl-cap: Count of missing values for each variable after cleaning
+#| label: tbl-clean_data
+#| tbl-cap: Clean Data Preview
+
+clean_house_price <- all_house_price %>%
+  select(-c(SizeRank, RegionID, RegionType, RegionName))
+
+kable(head(clean_house_price)[, 1:7])
 
 ```
+
+After the initial cleanup, we performed a more detailed grouping and summarization of the data by state on a monthly basis showing in left table of @tbl-processed.
+For each state, we calculated the mean house price, excluding any missing values to ensure the accuracy of our results.
+This aggregation provided a clearer view of the monthly pricing trends across different states, serving as a foundational analysis for identifying patterns and anomalies.
+
+The data was further processed to derive yearly price trends from the monthly data showing in right table of @tbl-processed.
+We transformed the dataset by pivoting the monthly columns into a single column of house prices, tagged by corresponding months.
+After converting these month labels into date formats and extracting the year, we calculated the average house price per state for each year.
+This yearly aggregation helped in examining longer-term trends and making more strategic conclusions about the housing market dynamics.
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-pos: 'h'
+#| label: tbl-processed
+#| tbl-cap: Processed Data Preview (by Month & by Year)
+#| layout-ncol: 2
+#| layout-nrow: 1
+
+state_house_price_by_month <- clean_house_price %>%
+  group_by(StateName) %>%
+  summarise(across(where(is.numeric), mean, na.rm = TRUE))
+
+state_house_price_by_year <- state_house_price_by_month %>%
+  pivot_longer(
+    cols = starts_with("20"),
+    names_to = "Month",
+    values_to = "HousePrice"
+  ) %>%
+  mutate(Year = year(ymd(Month))) %>%
+  group_by(StateName, Year) %>%
+  summarise(AvgHousePrice = mean(HousePrice, na.rm = TRUE), .groups = 'drop')
+
+kable(head(state_house_price_by_month)[, 1:3])
+
+kable(head(state_house_price_by_year))
+
+```
+
+Finally, as @tbl-hm_data shows, to create detailed heatmaps that visualize the distribution of average house prices across the United States, we enriched our dataset with additional geographic information.
+The subregion column are all NA since we do not need more detail than cities.
+This enhancement involved integrating region-specific data, along with precise longitude and latitude coordinates for each state.
+By incorporating these spatial dimensions, we were able to generate a more nuanced representation of the data, enabling us to pinpoint variations in housing prices with greater accuracy.
+
+```{r}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-pos: 'h'
+#| label: tbl-hm_data
+#| tbl-cap: Heatmap Data
+
+us_states_map <- map_data("state") # Getting map data for US states
+
+state_abbrev_to_full <- c( # Define the mapping from state abbreviations to full names
+  AL = "alabama", AK = "alaska", AZ = "arizona", AR = "arkansas", CA = "california",
+  CO = "colorado", CT = "connecticut", DE = "delaware", FL = "florida", GA = "georgia",
+  HI = "hawaii", ID = "idaho", IL = "illinois", IN = "indiana", IA = "iowa",
+  KS = "kansas", KY = "kentucky", LA = "louisiana", ME = "maine", MD = "maryland",
+  MA = "massachusetts", MI = "michigan", MN = "minnesota", MS = "mississippi",
+  MO = "missouri", MT = "montana", NE = "nebraska", NV = "nevada", NH = "new hampshire",
+  NJ = "new jersey", NM = "new mexico", NY = "new york", NC = "north carolina",
+  ND = "north dakota", OH = "ohio", OK = "oklahoma", OR = "oregon", PA = "pennsylvania",
+  RI = "rhode island", SC = "south carolina", SD = "south dakota", TN = "tennessee",
+  TX = "texas", UT = "utah", VT = "vermont", VA = "virginia", WA = "washington",
+  WV = "west virginia", WI = "wisconsin", WY = "wyoming", DC = "district of columbia"
+)
+
+overall_hm_data <- na.omit(state_house_price_by_year)
+overall_hm_data$StateName <- tolower(state_abbrev_to_full[overall_hm_data$StateName])
+
+overall_hm_data <- overall_hm_data %>%
+  group_by(StateName) %>%
+  summarise(AvgHousePrice = mean(AvgHousePrice, na.rm = TRUE))
+
+overall_hm_map_merged_data <- merge(us_states_map, overall_hm_data, by.x = "region", by.y = "StateName")
+
+kable(head(overall_hm_map_merged_data))
+
+```
+
+By meticulously cleaning and structuring the data through these steps, we ensured that our dataset was not only more manageable but also primed for high-quality, reliable analysis.
+This thorough preparation was instrumental in supporting our subsequent charting, statistical analyses and predictive modeling efforts.
 
@@
 #| include: false
-# Models Setup
 
@@
 #| fig-cap: Trend of Average House Price from 2000 to 2024
+#| fig-width: 7
+#| fig-height: 3.5
 
@@
 
+extrema <- state_all %>%
+  mutate(is_peak = ifelse(lag(AverageHousePrice, 1, default = first(AverageHousePrice)) < AverageHousePrice & 
+                            lead(AverageHousePrice, 1, default = last(AverageHousePrice)) < AverageHousePrice,
+                          TRUE, FALSE),
+         is_trough = ifelse(lag(AverageHousePrice, 1, default = first(AverageHousePrice)) > AverageHousePrice & 
+                              lead(AverageHousePrice, 1, default = last(AverageHousePrice)) > AverageHousePrice,
+                            TRUE, FALSE)) %>%
+  filter(is_peak | is_trough) %>%
+  select(Date, AverageHousePrice, is_peak, is_trough)
+
+extrema <- extrema %>%
+  filter(format(Date, "%Y") %in% c("2007", "2012", "2022"))
+
 ggplot(state_all, aes(x = Date, y = AverageHousePrice)) + # Plotting
-  geom_line(color = "blue") + 
+  geom_line(color = "black") + 
+  geom_text(data = extrema, aes(label = format(Date, "%Y")), nudge_y = 10000, check_overlap = TRUE, color = "black") +
+  geom_vline(data = extrema, aes(xintercept = as.numeric(Date)), linetype = "dashed", color = "red") +
   labs(x = "Year", y = "Average House Price ($)") + 
@@
 ```
+
+@fig-trend_overall illustrates the average house price trend from 2000 to 2024.
+This graph highlights several inflection points in housing prices over the span of nearly a quarter of a century.
+A significant peak occurs in 2006, followed by a notable dip culminating in a trough in 2012.
+Subsequently, prices climb, reaching another prominent peak in 2022.
+The chart clearly marks these fluctuations, drawing attention to the cyclic nature of the housing market over time.
+When we ignore these brief highs or lows, the overall trend of housing prices is a clear upward trend.
 
@@
 #| fig-cap: Trend of Average House Price from 2000 to 2024 by House Type
+#| fig-width: 7
+#| fig-height: 3.5
 
-process_data <- function(data, label) { # Function to process each dataset
+process_data <- function(data, label) {
   data %>% 
@@
 
-data_list <- list( # Processing all datasets
-  one_bedroom = state_1b_house_price_by_month,
-  two_bedroom = state_2b_house_price_by_month,
-  three_bedroom = state_3b_house_price_by_month,
-  four_bedroom = state_4b_house_price_by_month,
-  five_plus_bedroom = state_5bplus_house_price_by_month
+data_list <- list(
+  `1B` = state_1b_house_price_by_month,
+  `2B` = state_2b_house_price_by_month,
+  `3B` = state_3b_house_price_by_month,
+  `4B` = state_4b_house_price_by_month,
+  `5B+` = state_5bplus_house_price_by_month
 )
@@
 
-combined_data <- bind_rows(processed_data) # Combining all processed data into one dataframe
+combined_data <- bind_rows(processed_data)
 
-ggplot(combined_data, aes(x = Date, y = AverageHousePrice, color = Trend)) + # Plotting
+ggplot(combined_data, aes(x = Date, y = AverageHousePrice, color = Trend)) +
   geom_line() + 
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  scale_color_viridis_d(name = "House Type")
+  scale_color_viridis_d(name = "Number of Bedrooms")
 
@@
 
+@fig-trend_by_type segments the average house price trends from 2000 to 2024 by house type, from one-bedroom to houses with five or more bedrooms.
+The trajectory for each house type is distinct, with the five-plus-bedroom category reaching the highest average prices, particularly exhibiting a sharp rise after 2012.
+The four-bedroom category shows substantial growth, albeit less sharply than the largest homes.
+One and two-bedroom homes reveal a more gradual increase, suggesting a different progression in price trends compared to larger properties.
+The disparities among the trends of different house types highlight the varied performance within the housing market.
+The overall price trend of different bedroom types is also showing a clear upward trend.
+
 ## Heat Maps
+
+The heat map in @fig-hm_overall illustrates the average house prices across different states in the United States.
+The color gradient represents the average house price range, with the scale on the right indicating higher values in warmer colors (yellows and reds) and lower values in cooler colors (purples).
+The states colored with the warmer end of the spectrum exhibit higher average house prices, whereas those in the cooler end show lower average prices.
+There appears to be a geographical pattern where certain regions are predominantly one color, indicating a regional similarity in housing prices.
+States in the central part of the map are mostly in the mid-range of the color scale, neither the highest nor the lowest in average house price.
+The variation in color intensity among the states suggests a diversity in average house pricing across the country.
+Further, in the @fig-hm_overall, we can observe that there are a few states, particularly on the east and west coasts, such as California and New York, that are shaded in these most yellowish colour.
+These states stand out from the rest, suggesting they have notably higher average house prices compared to states with cooler colors (purples and dark blues).
 
@@
 
-overall_hm_data <- na.omit(state_house_price_by_year) # Delete all NAs
-overall_hm_data$StateName <- tolower(state_abbrev_to_full[overall_hm_data$StateName]) # Convert StateName from abbreviation to full name
+overall_hm_data <- na.omit(state_house_price_by_year)
+overall_hm_data$StateName <- tolower(state_abbrev_to_full[overall_hm_data$StateName])
 
-overall_hm_data <- overall_hm_data %>% # Taking average for house price based on states
+overall_hm_data <- overall_hm_data %>%
   group_by(StateName) %>%
@@
 
-overall_hm_map_merged_data <- merge(us_states_map, overall_hm_data, by.x = "region", by.y = "StateName") # Merge data ready for chart
+overall_hm_map_merged_data <- merge(us_states_map, overall_hm_data, by.x = "region", by.y = "StateName")
 
-ggplot() + # Generating heat map
+ggplot() +
   geom_polygon(data = overall_hm_map_merged_data, aes(x = long, y = lat, group = group, fill = AvgHousePrice), color = "white") +
@@
 ```
+
+@fig-hm_by_type comprises a collection of heatmaps (labeled a through e) that delineate the average house prices by state within the United States, segmented by the number of bedrooms.
+In the one-bedroom category @fig-hm_by_type-1, the color distribution is relatively homogeneous, suggesting minor variations in average prices across the states.
+Moving to the two-bedroom category @fig-hm_by_type-2, there is a discernible increase in the range of colors, which implies a modest expansion in the price range across different states.
+
+As the number of bedrooms increases to three @fig-hm_by_type-3, the color contrast intensifies, indicative of a wider disparity in average prices among the states.
+This trend continues with the four-bedroom heatmap @fig-hm_by_type-4, where a richer variety of colors emerges, particularly with several states displaying notably darker hues, signaling higher average prices.
+
+The heatmap for homes with five or more bedrooms @fig-hm_by_type-5 showcases the most striking contrast in color intensity, with several states marked by significantly darker shades.
+This pattern demonstrates that there is a broad spectrum of average prices for larger homes, with some states showing considerably higher average prices than others, as evidenced by the more intense coloring.
+Each heatmap distinctly represents the average price range for each house category, visually emphasizing the range from lower to higher average prices across the states.
 
@@
 #| layout-nrow: 3
+#| fig-width: 7
+#| fig-height: 3
 
-us_states_map <- map_data("state") # Getting map data for US states
+us_states_map <- map_data("state")
 
@@
 #### US States 1b House Price Heat Map ####
-oneb_hm_data <- na.omit(state_1b_house_price_by_year) # Delete all NAs
-oneb_hm_data$StateName <- tolower(state_abbrev_to_full[oneb_hm_data$StateName]) # Convert StateName from abbreviation to full name
+oneb_hm_data <- na.omit(state_1b_house_price_by_year)
+oneb_hm_data$StateName <- tolower(state_abbrev_to_full[oneb_hm_data$StateName])
 
-oneb_hm_data <- oneb_hm_data %>% # Taking average for house price based on states
+oneb_hm_data <- oneb_hm_data %>%
   group_by(StateName) %>%
@@
 
-oneb_hm_map_merged_data <- merge(us_states_map, oneb_hm_data, by.x = "region", by.y = "StateName") # Merge data ready for chart
+oneb_hm_map_merged_data <- merge(us_states_map, oneb_hm_data, by.x = "region", by.y = "StateName")
 
-ggplot() + # Generating heat map
+ggplot() +
   geom_polygon(data = oneb_hm_map_merged_data, aes(x = long, y = lat, group = group, fill = AvgHousePrice), color = "white") +

### 2024-04-14 outputs/paper/paper.qmd
paper updated

@@
 ---
-title: "TBD"
-subtitle: "TBD"
+title: "Analyzing Trends in U.S. Residential Real Estate Prices"
+subtitle: "A Data-Driven Approach to Understand the Market"
 author: 

### 2024-04-14 outputs/paper/paper.qmd
paper updated

--- 2024-04-14 outputs/paper/paper.qmd
@@
 date-format: long
-abstract: "TBD"
+abstract: "This research explores the dynamics of the U.S.
+housing market.
+It provides a comprehensive analysis of price trends across different states and housing types, revealing significant escalations in recent years.
+The study elucidates how location, property characteristics, and external economic factors influence house prices."
 format: pdf

### 2024-04-14 outputs/paper/paper.qmd
paper updated

@@
 
+### Model Setup
+
+The setup for the Multiple Regression model is straightforward, with direct relationships between the independent variables (Year and NumBedroom) and the dependent variable (house price).
+This model assumes that these relationships are linear and additive.
+
+$$
+TBD
+$$
+
+In contrast, the Polynomial Regression model requires a more intricate setup.
+It involves creating polynomial transformations of the independent variables to allow the model to fit non-linear trends.
+Specifically, the second-degree polynomial terms are created to model the curvature in the data, accounting for the possibility that the effect of time and the number of bedrooms on house prices changes at different rates.
+This setup acknowledges that increases in house prices may accelerate or decelerate over time and that the price premium for additional bedrooms may not be constant but could increase or decrease with the number of bedrooms.
+
+$$
+TBD
+$$
+
+The Generalized Additive Model (GAM) represents a flexible approach to modeling complex, non-linear relationships between the dependent and independent variables.
+The setup for the GAM model in this study involves specifying smooth functions of the predictors (likely year and number of bedrooms), allowing for the accommodation of non-linear trends that cannot be captured by traditional linear models.
+Unlike polynomial regression, which explicitly models non-linearity through polynomial terms, GAM uses smoothing functions that are fitted to the data in a non-parametric manner.
+This approach enables the model to adapt to the shape of the underlying trend without pre-specifying the form of the relationship, thus providing a tailored fit to the peculiarities in the data.
+
+$$
+TBD
+$$
+
+### Multiple Regression
+
+The Multiple Regression model reveals some intriguing insights about the housing market.
+It shows a substantial negative intercept, suggesting that starting from the model's baseline (which may be the earliest year in the data, though not directly interpretable), the house prices are initially low.
+The positive coefficient for 'Year' implies that there's a general trend of increasing house prices over time.
+Every passing year is associated with an increase in the average house price, as indicated by the coefficient of 6,958.769.
+Additionally, the 'NumBedroom' variable carries a positive coefficient of 59,897.327, which suggests that houses with more bedrooms tend to have higher prices.
+This model accounts for 44.5% of the variability in house prices (R-squared = 0.445), which is respectable but indicates that over half of the variance is explained by factors not included in the model.
+
+### Polynomial Regression
+
+Switching to the Polynomial Regression model, the positive intercept is considerably high, indicating that the baseline price for houses at the starting point of the model is substantial.
+The inclusion of polynomial terms for 'Year' and 'NumBedroom' indicates a non-linear relationship between these variables and house prices.
+The positive coefficients for the polynomial terms of 'Year' suggest an initial increase in prices, but as the magnitude of the coefficients for the quadratic terms is lower, this might indicate a slowing growth rate or a potential future downturn in prices.
+The coefficients for 'NumBedroom' follow a similar pattern, hinting that the increase in price associated with additional bedrooms becomes less pronounced as the number of bedrooms increases.
+This model fits the data slightly better than the Multiple Regression model, as evidenced by a higher R-squared value of 0.479.
+The lower AIC and BIC values suggest it is more efficient, and the reduced RMSE indicates improved prediction accuracy.
+
 ```{r}
@@
 ```
+
+### Generalized Additive Model (GAM)
+
+The results of the GAM model as shown in Table 6 indicate a significant positive intercept, similar to the polynomial regression model, suggesting a high base price for the houses at the starting point.
+The R-squared value of 0.506 is an improvement over the linear models, explaining approximately 50.6% of the variability in house prices.
+This is indicative of the GAM's superior capability to capture the variance in house prices with its non-linear approach.
+
+The AIC and BIC values are slightly lower than those in the polynomial regression model, which implies that the GAM provides a better model fit with less information loss.
+However, the reduction is not as substantial, suggesting that while the GAM improves upon the fit of linear models, it may not be drastically different in terms of information criteria.
+
+The RMSE of the GAM is slightly lower than that of the multiple regression but higher than that of the polynomial regression, indicating that the model's predictions are reasonably accurate, although there might be some room for improvement.
+This metric confirms that while the GAM has enhanced the model's predictive capability, it does so in a way that is comparable but not necessarily superior to the polynomial approach, potentially due to the complexity and variability inherent in housing market data.
 
@@
 ```
+
+### Further Justification
+
+In comparing the Multiple Regression, Polynomial Regression, and Generalized Additive Model (GAM) applied to the housing market data, each model serves a distinct purpose based on the complexity of the relationships it can capture between the predictors and the house prices.
+
+The Multiple Regression model, with its linear approach, is the most straightforward, offering a baseline model for comparison.
+It is the least complex and most transparent, which makes it a good starting point for analysis but also the most limited in capturing the nuanced behavior of the housing market.
+The Polynomial Regression model advances this analysis by introducing non-linearity through polynomial terms, providing a significantly better fit as evidenced by its improved R-squared value and lower information criteria scores.
+It acknowledges and captures the more intricate patterns in housing prices related to time and property features.
+
+GAM takes flexibility a step further, not being constrained by predefined polynomial terms, which allows for an even more nuanced understanding of the data.
+This model adapts to the data's inherent trends through the use of smoothing functions, leading to an improved R-squared value compared to the Multiple Regression model.
+However, when compared to Polynomial Regression, the gains in model performance are modest.
+While the GAM model offers a refined approach, it does not significantly outperform the Polynomial Regression in terms of the information criteria and RMSE, suggesting that the increased complexity of GAM may not be necessary for this particular dataset.
+
+Ultimately, the choice between these models should be informed by the balance between the need for model simplicity and the desire to capture complex data relationships.
+While the Polynomial Regression strikes a balance offering a good fit without overly complicating the model, the GAM provides the most flexibility, which could be essential for more intricate datasets or where the precise nature of relationships between variables is less understood.
 

### 2024-04-15 outputs/paper/paper.qmd
paper updated

@@
 #| fig-pos: 'h'
-#| label: fig-trend_overall
-#| fig-cap: Trend of Average House Price from 2000 to 2024
+#| label: fig-trend
+#| fig-subcap: ["Trend of Average House Price from 2000 to 2024", "Trend of Average House Price from 2000 to 2024 by House Type"]
+#| layout-ncol: 1
+#| layout-nrow: 2
 #| fig-width: 7
-#| fig-height: 3.5
+#| fig-height: 4
 
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-```
-
-@fig-trend_overall illustrates the average house price trend from 2000 to 2024.
-This graph highlights several inflection points in housing prices over the span of nearly a quarter of a century.
-A significant peak occurs in 2006, followed by a notable dip culminating in a trough in 2012.
-Subsequently, prices climb, reaching another prominent peak in 2022.
-The chart clearly marks these fluctuations, drawing attention to the cyclic nature of the housing market over time.
-When we ignore these brief highs or lows, the overall trend of housing prices is a clear upward trend.
-
-```{r}
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-pos: 'h'
-#| label: fig-trend_by_type
-#| fig-cap: Trend of Average House Price from 2000 to 2024 by House Type
-#| fig-width: 7
-#| fig-height: 3.5
 
@@
 
-@fig-trend_by_type segments the average house price trends from 2000 to 2024 by house type, from one-bedroom to houses with five or more bedrooms.
+@fig-trend-1 illustrates the average house price trend from 2000 to 2024.
+This graph highlights several inflection points in housing prices over the span of nearly a quarter of a century.
+A significant peak occurs in 2006, followed by a notable dip culminating in a trough in 2012.
+Subsequently, prices climb, reaching another prominent peak in 2022.
+The chart clearly marks these fluctuations, drawing attention to the cyclic nature of the housing market over time.
+When we ignore these brief highs or lows, the overall trend of housing prices has a clear upward trend.
+
+@fig-trend-2 segments the average house price trends from 2000 to 2024 by house type, from one-bedroom to houses with five or more bedrooms.
 The trajectory for each house type is distinct, with the five-plus-bedroom category reaching the highest average prices, particularly exhibiting a sharp rise after 2012.
@@
 #| fig-cap: Average Price by State in the US for All House Types
+#| fig-width: 7
+#| fig-height: 2.5
 
@@
 
-@fig-hm_by_type comprises a collection of heatmaps (labeled a through e) that delineate the average house prices by state within the United States, segmented by the number of bedrooms.
-In the one-bedroom category @fig-hm_by_type-1, the color distribution is relatively homogeneous, suggesting minor variations in average prices across the states.
-Moving to the two-bedroom category @fig-hm_by_type-2, there is a discernible increase in the range of colors, which implies a modest expansion in the price range across different states.
-
-As the number of bedrooms increases to three @fig-hm_by_type-3, the color contrast intensifies, indicative of a wider disparity in average prices among the states.
-This trend continues with the four-bedroom heatmap @fig-hm_by_type-4, where a richer variety of colors emerges, particularly with several states displaying notably darker hues, signaling higher average prices.
-
-The heatmap for homes with five or more bedrooms @fig-hm_by_type-5 showcases the most striking contrast in color intensity, with several states marked by significantly darker shades.
-This pattern demonstrates that there is a broad spectrum of average prices for larger homes, with some states showing considerably higher average prices than others, as evidenced by the more intense coloring.
-Each heatmap distinctly represents the average price range for each house category, visually emphasizing the range from lower to higher average prices across the states.
-
 ```{r}
@@
 #| layout-nrow: 3
-#| fig-width: 7
+#| fig-width: 6.5
 #| fig-height: 3
@@
 
+@fig-hm_by_type comprises a collection of heatmaps (labeled a through e) that delineate the average house prices by state within the United States, segmented by the number of bedrooms.
+In the one-bedroom category @fig-hm_by_type-1, the color distribution is relatively homogeneous, suggesting minor variations in average prices across the states.
+Moving to the two-bedroom category @fig-hm_by_type-2, there is a discernible increase in the range of colors, which implies a modest expansion in the price range across different states.
+
+As the number of bedrooms increases to three @fig-hm_by_type-3, the color contrast intensifies, indicative of a wider disparity in average prices among the states.
+This trend continues with the four-bedroom heatmap @fig-hm_by_type-4, where a richer variety of colors emerges, particularly with several states displaying notably darker hues, signaling higher average prices.
+
+The heatmap for homes with five or more bedrooms @fig-hm_by_type-5 showcases the most striking contrast in color intensity, with several states marked by significantly darker shades.
+This pattern demonstrates that there is a broad spectrum of average prices for larger homes, with some states showing considerably higher average prices than others, as evidenced by the more intense coloring.
+Each heatmap distinctly represents the average price range for each house category, visually emphasizing the range from lower to higher average prices across the states.
+
 ## Modeling
@@
 
+$$\text{Price} = \beta_0 + \beta_1 \times \text{Year} + \beta_2 \times \text{NumBedroom}$$
+
 The setup for the Multiple Regression model is straightforward, with direct relationships between the independent variables (Year and NumBedroom) and the dependent variable (house price).
@@
 
-$$
-TBD
-$$
+$$\text{Price} = \beta_0 + \beta_1 \times \text{Year} + \beta_2 \times \text{Year}^2 + \beta_3 \times \text{NumBedroom} + \beta_4 \times \text{NumBedroom}^2$$
 
@@
 
-$$
-TBD
-$$
+$$\text{Intercept for GAM} = \beta_0$$
 
 The Generalized Additive Model (GAM) represents a flexible approach to modeling complex, non-linear relationships between the dependent and independent variables.
-The setup for the GAM model in this study involves specifying smooth functions of the predictors (likely year and number of bedrooms), allowing for the accommodation of non-linear trends that cannot be captured by traditional linear models.
+This Model does not really have a fixed equation with coefficients due to its use of smoothing functions.
+The setup for the GAM model in this study involves specifying the smooth functions of the predictors, year and number of bedrooms, allowing for the accommodation of non-linear trends that cannot be captured by traditional linear models.
 Unlike polynomial regression, which explicitly models non-linearity through polynomial terms, GAM uses smoothing functions that are fitted to the data in a non-parametric manner.
 This approach enables the model to adapt to the shape of the underlying trend without pre-specifying the form of the relationship, thus providing a tailored fit to the peculiarities in the data.
-
-$$
-TBD
-$$
-
-### Multiple Regression
-
-The Multiple Regression model reveals some intriguing insights about the housing market.
-It shows a substantial negative intercept, suggesting that starting from the model's baseline (which may be the earliest year in the data, though not directly interpretable), the house prices are initially low.
-The positive coefficient for 'Year' implies that there's a general trend of increasing house prices over time.
-Every passing year is associated with an increase in the average house price, as indicated by the coefficient of 6,958.769.
-Additionally, the 'NumBedroom' variable carries a positive coefficient of 59,897.327, which suggests that houses with more bedrooms tend to have higher prices.
-This model accounts for 44.5% of the variability in house prices (R-squared = 0.445), which is respectable but indicates that over half of the variance is explained by factors not included in the model.
-
-### Polynomial Regression
-
-Switching to the Polynomial Regression model, the positive intercept is considerably high, indicating that the baseline price for houses at the starting point of the model is substantial.
-The inclusion of polynomial terms for 'Year' and 'NumBedroom' indicates a non-linear relationship between these variables and house prices.
-The positive coefficients for the polynomial terms of 'Year' suggest an initial increase in prices, but as the magnitude of the coefficients for the quadratic terms is lower, this might indicate a slowing growth rate or a potential future downturn in prices.
-The coefficients for 'NumBedroom' follow a similar pattern, hinting that the increase in price associated with additional bedrooms becomes less pronounced as the number of bedrooms increases.
-This model fits the data slightly better than the Multiple Regression model, as evidenced by a higher R-squared value of 0.479.
-The lower AIC and BIC values suggest it is more efficient, and the reduced RMSE indicates improved prediction accuracy.
 
@@
 
+After some understanding of model choices, @tbl-linear_model_result and @tbl-nonlinear_model_result displays the regression results, which will be explain with detials in later sections.
+
+### Multiple Regression
+
+$$\text{Price} = -13,969,628.019 + 6,958.769 \times \text{Year} + 59,897.327 \times \text{NumBedroom}$$
+
+The Multiple Regression model reveals some intriguing insights about the housing market.
+It shows a substantial negative intercept, suggesting that starting from the model's baseline (which may be the earliest year in the data, though not directly interpretable), the house prices are initially low.
+The positive coefficient for 'Year' implies that there's a general trend of increasing house prices over time.
+Every passing year is associated with an increase in the average house price, as indicated by the coefficient of 6,958.769.
+Additionally, the 'NumBedroom' variable carries a positive coefficient of 59,897.327, which suggests that houses with more bedrooms tend to have higher prices.
+This model accounts for 44.5% of the variability in house prices (R-squared = 0.445), which is respectable but indicates that over half of the variance is explained by factors not included in the model.
+
+### Polynomial Regression
+
+$$\text{Price} = 212,976.031 + 3,866,485.706 \times \text{Year} + 1,601,816.793 \times \text{Year}^2$$
+$$+ 6,591,537.312 \times \text{NumBedroom} + 1,365,656.076 \times \text{NumBedroom}^2$$
+
+Switching to the Polynomial Regression model, the positive intercept is considerably high, indicating that the baseline price for houses at the starting point of the model is substantial.
+The inclusion of polynomial terms for 'Year' and 'NumBedroom' indicates a non-linear relationship between these variables and house prices.
+The positive coefficients for the polynomial terms of 'Year' suggest an initial increase in prices, but as the magnitude of the coefficients for the quadratic terms is lower, this might indicate a slowing growth rate or a potential future downturn in prices.
+The coefficients for 'NumBedroom' follow a similar pattern, hinting that the increase in price associated with additional bedrooms becomes less pronounced as the number of bedrooms increases.
+This model fits the data slightly better than the Multiple Regression model, as evidenced by a higher R-squared value of 0.479.
+The lower AIC and BIC values suggest it is more efficient, and the reduced RMSE indicates improved prediction accuracy.
+
 ### Generalized Additive Model (GAM)
+
+$$\text{Intercept for GAM} = 212,976.031$$
 
@@
 
+Furthermore, two charts in @fig-GAM_fit each showing the smoothed relationship between the predictor and the response variable (house price), with a solid line representing the estimated function and dashed lines indicating the confidence interval around the estimate.
+
+The plot on the left shows the relationship between 'Year' and house prices.
+There’s a clear non-linear trend.
+The prices appear to have a cyclical pattern with a significant peak around 2006, followed by a sharp decline, which likely corresponds to the global financial crisis of 2008.
+After this trough, there’s a recovery with prices trending upwards again, but with a suggestion of leveling off towards the end of the period.
+This could indicate a stabilization in the market or potentially the beginning of another downturn.
+
+The plot on the right shows the relationship between 'NumBedroom' and house prices.
+This relationship is more straightforward: as the number of bedrooms increases, so does the house price.
+The upward trend appears to be exponential rather than linear, suggesting that additional bedrooms have an increasingly positive effect on house prices.
+The confidence interval widens with the number of bedrooms, indicating more variability in the data for houses with more bedrooms, which could be due to fewer observations in the higher bedroom categories or greater variance in prices for larger homes.
+
+Both plots reveal the flexibility of GAMs in modeling complex relationships, allowing for the identification of trends that simpler linear models could miss.
+The confidence intervals suggest the model is fairly confident about its estimates, especially for the 'NumBedroom' variable.
+However, the wider intervals at the extremes of the data range would call for cautious interpretation in those areas.
+
 ```{r}
@@
 #| fig-pos: 'h'
-#| label: fig-GAM
+#| label: fig-GAM_fit
 #| tbl-cap: GAM Regression Fit Result
@@
 While the GAM model offers a refined approach, it does not significantly outperform the Polynomial Regression in terms of the information criteria and RMSE, suggesting that the increased complexity of GAM may not be necessary for this particular dataset.
+
+However, the visual representation of data trends provided by @fig-GAM_fit is not just intuitively appealing but also offers practical insights that might be obscured in tabular regression outputs.
+For example, the cyclical pattern observed in the 'Year' graph aligns with known economic events that affected housing prices, demonstrating GAM's capacity to capture externalities and temporal dynamics beyond the reach of conventional linear models.
 

### 2024-04-15 outputs/paper/paper.qmd
paper and bib updated

@@
 
-## 1
-
-## 2
-
-## 3
-
-## 4
+## Global Trends
+
+The analysis presented in this paper has identified a consistent upward trend in housing prices that reflects global economic tendencies.
+This trend goes hand in hand with patterns observed in other developed nations, suggesting a widespread phenomenon rather than isolated occurrences.
+These global trends, demonstrated through extensive data, emphasize the housing market's sensitivity to worldwide economic conditions and highlight the necessity for international comparative studies to better understand localized impacts.
+
+The escalation of housing prices, as evidenced by the comprehensive data analysis, resonates with broader economic trends such as inflation rates and demographic shifts.
+
+## Potential Preference Shifts
+
+A notable aspect of the housing market dynamics is the significant rise in prices for larger homes, as highlighted by the data.
+This may be indicative of a potential shift in preferences, with more people possibly seeking spacious accommodations, perhaps influenced by recent global events such as the pandemic that emphasized the importance of personal space.
+Understanding these shifts is crucial, as they can have far-reaching effects on urban development, infrastructure requirements, and economic planning.
+
+## Geographical Position
+
+The geographical analysis, particularly the use of heat maps, has revealed sharp contrasts in price increments between regions.
+Coastal areas exhibit higher prices, which may be a consequence of urban sprawl, increased desirability, and limited supply.
+This geographical positioning underscores the multifaceted nature of real estate dynamics, where location continues to be a prime determinant of value.
+Such insights can guide infrastructural investments and real estate development to balance regional growth.
+
+## Possible Actions
+
+The implications of these findings are manifold.
+Policymakers and investors are provided with strategic insights that could guide investment decisions and policy frameworks.
+The public sector could leverage this knowledge to design interventions aimed at housing affordability, ensuring that market dynamics do not impede access to housing.
+Moreover, the identified price trends serve as a cautionary tale of widening economic disparities, necessitating proactive measures to ensure equitable housing opportunities across different societal strata.
 
@@
 
+The present study, while extensive, leaves room for further refinement.
+Future research could benefit from incorporating additional variables such as interest rates, housing supply metrics, and construction costs to explain price variability more comprehensively.
+Delving into microeconomic factors, such as individual income levels and the impact of local zoning laws, could also yield more granular insights into market dynamics.
+Additionally, longitudinal studies examining the long-term effects of policy changes on housing prices would be invaluable.
+Technological improvements in data collection, including real-time tracking of market listings and transactions, could provide a more immediate understanding of market fluctuations.
+Lastly, expanding the scope to include rental markets and affordable housing sectors would present a more holistic view of the housing ecosystem.
+
 # Conclusion {#sec-con}
+
+This paper has undertaken a detailed examination of the U.S.
+residential real estate market, revealing a complex interplay of factors influencing housing prices.
+The study has tracked price trends across various states and housing types, identifying significant price escalations over the last two decades.
+The analysis demonstrates that house prices are influenced not only by traditional factors such as location and property characteristics but also by broader economic indicators and potentially, shifts in societal preferences.
+
+Our findings suggest that while the U.S.
+housing market has certain unique attributes, it also shares common trends with global markets, particularly with regards to the upward trajectory of prices.
+The increasing prices for larger homes could be signaling a shift in consumer preferences, perhaps a post-pandemic effect, towards more spacious living accommodations.
+Additionally, the geographic analysis indicates a pronounced disparity in housing costs, with coastal regions exhibiting significantly higher prices, reflecting the enduring axiom of 'location, location, location' in property valuation.
+
+As comprehensive as this study is, there are avenues for improvement and expansion.
+Future studies should aim to integrate additional economic variables, such as interest rates and housing supply data, to provide a more complete picture of the market forces at play.
+Longitudinal analyses that track the impact of policy changes over time would also contribute to a deeper understanding of market dynamics.
+Technological advancements in data collection could facilitate a real-time analysis of market trends, providing an agile response to rapid market changes.
+Finally, expanding the scope of research to encompass rental markets and affordable housing will offer insights into the full spectrum of the housing market, a necessary step to ensure comprehensive housing strategies that serve all segments of society.
+
+In conclusion, the U.S.
+housing market is a dynamic and complex entity, influenced by a multitude of factors both local and global.
+This paper serves as a stepping stone towards a more profound understanding of these factors, providing a foundational analysis that stakeholders can build upon to make informed decisions that affect the fabric of communities across the nation.
 

### 2024-04-16 outputs/paper/paper.qmd
paper updated

--- 2024-04-15 outputs/paper/paper.qmd
@@
 date-format: long
-abstract: "This research explores the dynamics of the U.S.
-housing market.
-It provides a comprehensive analysis of price trends across different states and housing types, revealing significant escalations in recent years.
-The study elucidates how location, property characteristics, and external economic factors influence house prices."
+abstract: "This research examines the dynamics of the U.S.
+housing market, offering a detailed analysis of price trends across various states and types of housing, which have shown considerable increases in recent years.
+The study clarifies how factors such as location, property characteristics, and external economic influences impact house prices."
 format: pdf
@@
 
-This paper will focus on exploring trends in U.S.
-house prices, a pertinent issue as real estate markets worldwide have experienced significant increases.
-Cities like Toronto, Canada, and Beijing, China have witnessed substantial escalations in housing costs, reflecting a global phenomenon.
-This study shifts its focus to the United States, one of the most economically stable and developed countries in the world, to determine if similar trends persist in its real estate market.
-By examining the U.S.
-housing market, this research aims to uncover the principal factors influencing house prices.
+This paper concentrates on investigating trends in U.S.
+house prices, a relevant topic as real estate markets globally have undergone considerable increases [@citeGlobalHP].
+Cities such as Toronto, Canada [@citeCanadaHP], and Beijing, China [@citeChinaHP] have seen significant rises in housing costs, indicative of a widespread trend.
+This study redirects attention to the United States, among the most economically stable and developed nations globally, to ascertain whether similar trends are evident in its real estate market.
+Through an examination of the U.S.
+housing market, this research seeks to identify the primary factors that influence house prices.
 
 The U.S.
-real estate market offers a unique case study due to its diverse economic landscapes and varied housing markets across states and metropolitan areas.
-This analysis will consider variables such as location, number of bedrooms, and other key factors that are believed to significantly impact house pricing dynamics.
-Understanding these factors is crucial, as the findings will provide insights into how policy adjustments and economic changes can affect housing affordability.
-Ultimately, this research aims to serve as a guide for policymakers, investors, and the public, helping them make informed decisions regarding housing investments and urban planning.
+real estate market presents a distinctive case study due to its diverse economic conditions and varied housing markets across states and metropolitan areas [@CiteStatesHP].
+This analysis will examine variables such as location, number of bedrooms, and other critical factors that are considered to have a significant impact on house pricing dynamics.
+Understanding these factors is essential, as the findings will offer insights into how policy adjustments and economic changes can influence housing affordability.
+This research intends to serve as a resource for policymakers, investors, and the public, aiding them in making informed decisions about housing investments and urban planning.
 
@@
 
-Our study focuse on U.S.
-house price trends using the comprehensive datasets available from Zillow Research  [@citeZillowdata].
-This data repository offers an extensive array of real estate statistics, capturing the dynamics of house prices across various regions in the United States.
-As a cornerstone of transparency and public engagement, Zillow Research provides open access to these crucial real estate data, which are instrumental in understanding housing market behaviors.
-
-The dataset includes detailed information on house prices over several years, with breakdowns by state, city, and sometimes neighborhood levels, from median list prices to Zillow Home Value Index (ZHVI).
-This granularity allows for a nuanced analysis of house price trends and the identification of significant market dynamics, guiding potential real estate investments and policy formulations.
-
-We selected this dataset for its reliability, depth, and alignment with our study’s goals.
-As a trusted source in the real estate market, it provides accurate and timely data crucial for our analysis.
-Its comprehensive geographic and property-specific details allow for an in-depth examination of factors influencing house prices in the U.S.
-Additionally, the dataset’s structure supports efficient analysis and modeling of real estate trends.
+Our study focuses on U.S.
+house price trends using the extensive datasets available from Zillow Research [@citeZillowdata].
+This data repository provides a broad array of real estate statistics, capturing the dynamics of house prices across various regions in the United States.
+As a cornerstone of transparency and public engagement, Zillow Research offers open access to these vital real estate data, which are instrumental in understanding housing market behaviors.
+
+The dataset encompasses detailed information on house prices over several years, including breakdowns by state, city, and occasionally neighborhood levels, from median list prices to the Zillow Home Value Index (ZHVI).
+This level of detail facilitates a nuanced analysis of house price trends and the identification of significant market dynamics, which can guide potential real estate investments and policy formulations.
+
+We selected this dataset for its reliability, depth, and alignment with our study’s objectives.
+As a trusted source in the real estate market, it provides accurate and timely data essential for our analysis.
+Its extensive geographic and property-specific details enable an in-depth examination of factors influencing house prices in the U.S.
+Furthermore, the dataset’s structure facilitates efficient analysis and modeling of real estate trends.
 More details about this dataset can be found in the datasheet in @sec-app (Appendix).
@@
 
-Our analysis commenced with the essential task of refining the raw dataset downloaded from Zillow.
+Our analysis began with the critical task of refining the raw dataset downloaded from Zillow.
 The initial phase involved understanding the dataset’s structure.
-As the @tbl-raw_data showing, this dataset typically consists of multiple attributes such as Region ID, Region Type, location (states), year of transaction, house price, etc.
-Each record represents a unique property, providing a comprehensive snapshot of its characteristics.
+As shown in @tbl-raw_data, this dataset typically includes multiple attributes such as Region ID, Region Type, location (states), year of transaction, house price, etc.
+Each record represents a unique property, providing a detailed snapshot of its characteristics.
 
@@
 
-Then, we proceeded to clean and process the data as needed, setting the stage for deeper analysis.
-Our first step involved preprocessing the raw dataset from Zillow, which required several critical adjustments to ensure data quality and usability.
+We then proceeded to clean and process the data as required, setting the stage for more detailed analysis.
+Our first step involved preprocessing the raw dataset from Zillow, which necessitated several important adjustments to ensure data quality and usability.
 Initially, we removed the first row of the dataset, which represented national averages rather than state-specific data, to focus our analysis on regional trends.
-This step was crucial as the inclusion of national data could skew the results when examining state-level dynamics.
+This step was essential as the inclusion of national data could skew the results when examining state-level dynamics.
 
@@
 These columns were extraneous for our analysis, which aimed to concentrate on price fluctuations and trends across different states.
-By simplifying the dataset to whcih showing in @tbl-clean_data, we facilitated more focused and faster computations in subsequent steps.
+By simplifying the dataset as shown in @tbl-clean_data, we facilitated more focused and faster computations in subsequent steps.
 
@@
 
-After the initial cleanup, we performed a more detailed grouping and summarization of the data by state on a monthly basis showing in left table of @tbl-processed.
+After the initial cleanup, we conducted a more detailed grouping and summarization of the data by state on a monthly basis, as shown in the left table of @tbl-processed.
 For each state, we calculated the mean house price, excluding any missing values to ensure the accuracy of our results.
-This aggregation provided a clearer view of the monthly pricing trends across different states, serving as a foundational analysis for identifying patterns and anomalies.
-
-The data was further processed to derive yearly price trends from the monthly data showing in right table of @tbl-processed.
+This aggregation offered a clearer view of the monthly pricing trends across different states, serving as a foundational analysis for identifying patterns and anomalies.
+
+The data was further processed to derive yearly price trends from the monthly data, as shown in the right table of @tbl-processed.
 We transformed the dataset by pivoting the monthly columns into a single column of house prices, tagged by corresponding months.
 After converting these month labels into date formats and extracting the year, we calculated the average house price per state for each year.
-This yearly aggregation helped in examining longer-term trends and making more strategic conclusions about the housing market dynamics.
+This yearly aggregation aided in examining longer-term trends and making more strategic conclusions about the housing market dynamics.
 
@@
 
-Finally, as @tbl-hm_data shows, to create detailed heatmaps that visualize the distribution of average house prices across the United States, we enriched our dataset with additional geographic information.
-The subregion column are all NA since we do not need more detail than cities.
+Finally, as shown in @tbl-hm_data, to create detailed heatmaps that visualize the distribution of average house prices across the United States, we enriched our dataset with additional geographic information.
+The subregion column was marked as NA since more detail than cities was not required.
 This enhancement involved integrating region-specific data, along with precise longitude and latitude coordinates for each state.
@@
 By meticulously cleaning and structuring the data through these steps, we ensured that our dataset was not only more manageable but also primed for high-quality, reliable analysis.
-This thorough preparation was instrumental in supporting our subsequent charting, statistical analyses and predictive modeling efforts.
+This thorough preparation was instrumental in supporting our subsequent efforts in charting, statistical analyses, and predictive modeling.
 
 # Results {#sec-res}
-
-```{r}
-#| include: false
-
-#### Multiple Regression Model ####
-model_multi <- lm(AvgHousePrice ~ Year + NumBedroom, data = house_prices)
-
-#### Polynomial Regression Model ####
-model_poly <- lm(AvgHousePrice ~ poly(Year, 2) + poly(NumBedroom, 2), data = house_prices)
-
-#### Generalized Additive Models ####
-num_unique_years <- length(unique(house_prices$Year))
-num_unique_bedrooms <- length(unique(house_prices$NumBedroom))
-
-gam_model <- gam(AvgHousePrice ~ s(Year, k=num_unique_years-1) + s(NumBedroom, k=num_unique_bedrooms-1), data = house_prices)
-
-```
 
@@
 @fig-trend-1 illustrates the average house price trend from 2000 to 2024.
-This graph highlights several inflection points in housing prices over the span of nearly a quarter of a century.
-A significant peak occurs in 2006, followed by a notable dip culminating in a trough in 2012.
-Subsequently, prices climb, reaching another prominent peak in 2022.
-The chart clearly marks these fluctuations, drawing attention to the cyclic nature of the housing market over time.
-When we ignore these brief highs or lows, the overall trend of housing prices has a clear upward trend.
-
-@fig-trend-2 segments the average house price trends from 2000 to 2024 by house type, from one-bedroom to houses with five or more bedrooms.
-The trajectory for each house type is distinct, with the five-plus-bedroom category reaching the highest average prices, particularly exhibiting a sharp rise after 2012.
-The four-bedroom category shows substantial growth, albeit less sharply than the largest homes.
-One and two-bedroom homes reveal a more gradual increase, suggesting a different progression in price trends compared to larger properties.
-The disparities among the trends of different house types highlight the varied performance within the housing market.
-The overall price trend of different bedroom types is also showing a clear upward trend.
+This graph highlights several inflection points in housing prices over nearly a quarter of a century.
+A significant peak is observed in 2006, followed by a notable dip culminating in a trough in 2012.
+Subsequently, prices ascend, reaching another prominent peak in 2022.
+The chart clearly marks these fluctuations, emphasizing the cyclic nature of the housing market over time.
+Excluding these brief highs or lows, the overall trend of housing prices shows a clear upward trajectory.
+
+@fig-trend-2 segments the average house price trends from 2000 to 2024 by house type, ranging from one-bedroom to houses with five or more bedrooms.
+Each house type displays a distinct trajectory, with the category for homes with five or more bedrooms achieving the highest average prices, particularly showing a sharp rise after 2012.
+The four-bedroom category also demonstrates substantial growth, though less pronounced than that of the largest homes.
+One and two-bedroom homes exhibit a more gradual increase, indicating a different progression in price trends compared to larger properties.
+The disparities among the trends of different house types underscore the varied performance within the housing market.
+Despite these differences, the overall price trend across different bedroom types consistently shows a clear upward trajectory.
 
@@
 
-The heat map in @fig-hm_overall illustrates the average house prices across different states in the United States.
-The color gradient represents the average house price range, with the scale on the right indicating higher values in warmer colors (yellows and reds) and lower values in cooler colors (purples).
-The states colored with the warmer end of the spectrum exhibit higher average house prices, whereas those in the cooler end show lower average prices.
-There appears to be a geographical pattern where certain regions are predominantly one color, indicating a regional similarity in housing prices.
-States in the central part of the map are mostly in the mid-range of the color scale, neither the highest nor the lowest in average house price.
-The variation in color intensity among the states suggests a diversity in average house pricing across the country.
-Further, in the @fig-hm_overall, we can observe that there are a few states, particularly on the east and west coasts, such as California and New York, that are shaded in these most yellowish colour.
-These states stand out from the rest, suggesting they have notably higher average house prices compared to states with cooler colors (purples and dark blues).
+The heatmap in @fig-hm_overall illustrates the average house prices across different states in the United States.
+The color gradient signifies the range of average house prices, with warmer colors (yellows and reds) indicating higher values and cooler colors (purples) representing lower values.
+States depicted with warmer colors exhibit higher average house prices, while those in cooler colors show lower prices.
+A noticeable geographical pattern emerges, with certain regions displaying a uniform color, suggesting regional similarities in housing prices.
+States in the central part of the map typically fall in the mid-range of the color scale, neither the highest nor the lowest in average house price.
+The variation in color intensity across the states highlights the diversity in average house pricing throughout the country.
+Additionally, in @fig-hm_overall, states on the east and west coasts, such as California and New York, are shaded in the most intense yellowish colors, distinguishing them as having notably higher average house prices compared to states depicted with cooler colors.
 
@@
 
-@fig-hm_by_type comprises a collection of heatmaps (labeled a through e) that delineate the average house prices by state within the United States, segmented by the number of bedrooms.
-In the one-bedroom category @fig-hm_by_type-1, the color distribution is relatively homogeneous, suggesting minor variations in average prices across the states.
-Moving to the two-bedroom category @fig-hm_by_type-2, there is a discernible increase in the range of colors, which implies a modest expansion in the price range across different states.
-
-As the number of bedrooms increases to three @fig-hm_by_type-3, the color contrast intensifies, indicative of a wider disparity in average prices among the states.
-This trend continues with the four-bedroom heatmap @fig-hm_by_type-4, where a richer variety of colors emerges, particularly with several states displaying notably darker hues, signaling higher average prices.
-
-The heatmap for homes with five or more bedrooms @fig-hm_by_type-5 showcases the most striking contrast in color intensity, with several states marked by significantly darker shades.
-This pattern demonstrates that there is a broad spectrum of average prices for larger homes, with some states showing considerably higher average prices than others, as evidenced by the more intense coloring.
-Each heatmap distinctly represents the average price range for each house category, visually emphasizing the range from lower to higher average prices across the states.
+@fig-hm_by_type consists of a series of heatmaps (labeled a through e) that outline the average house prices by state within the United States, segmented by the number of bedrooms.
+In the one-bedroom category @fig-hm_by_type-1, the color distribution is relatively uniform, indicating minor variations in average prices across the states.
+As we progress to the two-bedroom category @fig-hm_by_type-2, there is a noticeable increase in the diversity of colors, suggesting a modest expansion in the price range across different states.
+
+As the number of bedrooms increases to three, as shown in @fig-hm_by_type-3, the color contrast intensifies, indicative of a wider disparity in average prices among the states.
+This trend continues with the four-bedroom heatmap @fig-hm_by_type-4, where a richer variety of colors emerges.
+Several states display notably darker hues, signaling higher average prices.
+This progression highlights the increasing variability in house prices as the size of the properties increases, reflecting differing market dynamics and housing demands across the states.
+
+The heatmap for homes with five or more bedrooms, as depicted in @fig-hm_by_type-5, showcases the most striking contrast in color intensity, with several states marked by significantly darker shades.
+This pattern indicates a broad spectrum of average prices for larger homes, with some states showing considerably higher average prices than others, as evidenced by the more intense coloring.
+Each heatmap distinctly represents the average price range for each house category, visually emphasizing the spectrum from lower to higher average prices across the states.
+This visual distinction underscores the significant variation in market conditions for larger properties across different regions.
 
@@
 
-The setup for the Multiple Regression model is straightforward, with direct relationships between the independent variables (Year and NumBedroom) and the dependent variable (house price).
-This model assumes that these relationships are linear and additive.
+The setup for the Multiple Regression model is straightforward, positing direct linear and additive relationships between the independent variables (Year and NumBedroom) and the dependent variable (house price).
+This model assumes that changes in the year and the number of bedrooms are directly proportional to changes in house prices, with each independent variable contributing additively to the predicted price.
 
@@
 
-In contrast, the Polynomial Regression model requires a more intricate setup.
-It involves creating polynomial transformations of the independent variables to allow the model to fit non-linear trends.
-Specifically, the second-degree polynomial terms are created to model the curvature in the data, accounting for the possibility that the effect of time and the number of bedrooms on house prices changes at different rates.
-This setup acknowledges that increases in house prices may accelerate or decelerate over time and that the price premium for additional bedrooms may not be constant but could increase or decrease with the number of bedrooms.
+The Polynomial Regression model involves a more complex setup, incorporating polynomial transformations of the independent variables to accommodate non-linear trends in the data.
+Specifically, second-degree polynomial terms are introduced to model curvature, reflecting that the impacts of time and the number of bedrooms on house prices may vary at different rates.
+This approach recognizes that increases in house prices could either accelerate or decelerate over time, and that the price premium associated with additional bedrooms may not be constant, potentially increasing or decreasing with the number of bedrooms.
+This model thereby allows for a more nuanced understanding of how these variables interact to influence house prices.
 
@@
 
-The Generalized Additive Model (GAM) represents a flexible approach to modeling complex, non-linear relationships between the dependent and independent variables.
-This Model does not really have a fixed equation with coefficients due to its use of smoothing functions.
-The setup for the GAM model in this study involves specifying the smooth functions of the predictors, year and number of bedrooms, allowing for the accommodation of non-linear trends that cannot be captured by traditional linear models.
-Unlike polynomial regression, which explicitly models non-linearity through polynomial terms, GAM uses smoothing functions that are fitted to the data in a non-parametric manner.
-This approach enables the model to adapt to the shape of the underlying trend without pre-specifying the form of the relationship, thus providing a tailored fit to the peculiarities in the data.
+The Generalized Additive Model (GAM) provides a flexible approach for modeling complex, non-linear relationships between the dependent and independent variables.
+Distinct from models that use fixed equations with coefficients, GAM utilizes smoothing functions, enabling it to dynamically adapt to the data without a predefined formula.
+Additionally, GAM has been shown to fit well in the housing market context [@citeGAM].
+In this study, the setup for the GAM involves specifying smooth functions for the predictors, specifically year and number of bedrooms.
+This approach effectively accommodates non-linear trends that traditional linear models cannot adequately capture.
+
+Unlike polynomial regression, which directly incorporates non-linearity through polynomial terms, GAM utilizes smoothing functions that are fitted to the data in a non-parametric way.
+This method enables the model to conform to the underlying trend's shape, providing a tailored fit to specific patterns and anomalies in the data.
+This adaptability makes GAM particularly useful for exploring and understanding the nuanced behaviors of variables within complex datasets.
+
+After gaining some understanding of the model choices, @tbl-linear_model_result and @tbl-nonlinear_model_result display the regression results.
+These will be explained in detail in later sections.
+This structured approach helps in comparing the outcomes of different modeling strategies, illustrating how each model handles the complexity of the data and the specific characteristics of the variables involved.
+
+```{r}
+#| include: false
+
+#### Multiple Regression Model ####
+model_multi <- lm(AvgHousePrice ~ Year + NumBedroom, data = house_prices)
+
+#### Polynomial Regression Model ####
+model_poly <- lm(AvgHousePrice ~ poly(Year, 2) + poly(NumBedroom, 2), data = house_prices)
+
+#### Generalized Additive Models ####
+num_unique_years <- length(unique(house_prices$Year))
+num_unique_bedrooms <- length(unique(house_prices$NumBedroom))
+
+gam_model <- gam(AvgHousePrice ~ s(Year, k=num_unique_years-1) + s(NumBedroom, k=num_unique_bedrooms-1), data = house_prices)
+
+```
 
@@
 
-After some understanding of model choices, @tbl-linear_model_result and @tbl-nonlinear_model_result displays the regression results, which will be explain with detials in later sections.
-
 ### Multiple Regression
@@
 
-The Multiple Regression model reveals some intriguing insights about the housing market.
-It shows a substantial negative intercept, suggesting that starting from the model's baseline (which may be the earliest year in the data, though not directly interpretable), the house prices are initially low.
-The positive coefficient for 'Year' implies that there's a general trend of increasing house prices over time.
-Every passing year is associated with an increase in the average house price, as indicated by the coefficient of 6,958.769.
-Additionally, the 'NumBedroom' variable carries a positive coefficient of 59,897.327, which suggests that houses with more bedrooms tend to have higher prices.
-This model accounts for 44.5% of the variability in house prices (R-squared = 0.445), which is respectable but indicates that over half of the variance is explained by factors not included in the model.
+The Multiple Regression model provides some intriguing insights into the housing market.
+It reveals a substantial negative intercept, suggesting that starting from the model's baseline (possibly the earliest year in the dataset), house prices are initially low.
+The positive coefficient for 'Year' indicates a general trend of increasing house prices over time, with each passing year associated with an average increase in house price of \$6,958.769.
+Additionally, the 'NumBedroom' variable has a positive coefficient of \$59,897.327, indicating that houses with more bedrooms tend to command higher prices.
+This model explains 44.5% of the variability in house prices (R-squared = 0.445), which is respectable but also suggests that over half of the variance is due to factors not included in the model.
+This highlights the need to consider other variables or more complex models to better capture the dynamics affecting house prices.
 
@@
 
-Switching to the Polynomial Regression model, the positive intercept is considerably high, indicating that the baseline price for houses at the starting point of the model is substantial.
-The inclusion of polynomial terms for 'Year' and 'NumBedroom' indicates a non-linear relationship between these variables and house prices.
-The positive coefficients for the polynomial terms of 'Year' suggest an initial increase in prices, but as the magnitude of the coefficients for the quadratic terms is lower, this might indicate a slowing growth rate or a potential future downturn in prices.
-The coefficients for 'NumBedroom' follow a similar pattern, hinting that the increase in price associated with additional bedrooms becomes less pronounced as the number of bedrooms increases.
-This model fits the data slightly better than the Multiple Regression model, as evidenced by a higher R-squared value of 0.479.
-The lower AIC and BIC values suggest it is more efficient, and the reduced RMSE indicates improved prediction accuracy.
+The Polynomial Regression model reveals notable dynamics within the housing market with its considerably high positive intercept, suggesting that the baseline price for houses at the model's inception is substantial.
+The incorporation of polynomial terms for 'Year' and 'NumBedroom' illustrates a non-linear relationship between these variables and house prices.
+Specifically, the positive coefficients for the polynomial terms of 'Year' suggest an initial increase in prices.
+However, the lower magnitude of the coefficients for the quadratic terms might indicate a slowing growth rate or a potential future downturn in prices.
+
+Similarly, the coefficients for 'NumBedroom' follow a pattern that hints at diminishing returns; the increase in price associated with additional bedrooms becomes less pronounced as the number of bedrooms increases.
+This model provides a slightly better fit to the data compared to the Multiple Regression model, as evidenced by a higher R-squared value of 0.479.
+Furthermore, lower AIC and BIC values suggest greater efficiency in model selection, and the reduced RMSE points to improved prediction accuracy.
+This indicates that the Polynomial Regression model might be more adept at capturing the complexities of the housing market than a simpler linear approach.
 
@@
 
-The results of the GAM model as shown in Table 6 indicate a significant positive intercept, similar to the polynomial regression model, suggesting a high base price for the houses at the starting point.
-The R-squared value of 0.506 is an improvement over the linear models, explaining approximately 50.6% of the variability in house prices.
-This is indicative of the GAM's superior capability to capture the variance in house prices with its non-linear approach.
-
-The AIC and BIC values are slightly lower than those in the polynomial regression model, which implies that the GAM provides a better model fit with less information loss.
-However, the reduction is not as substantial, suggesting that while the GAM improves upon the fit of linear models, it may not be drastically different in terms of information criteria.
-
-The RMSE of the GAM is slightly lower than that of the multiple regression but higher than that of the polynomial regression, indicating that the model's predictions are reasonably accurate, although there might be some room for improvement.
-This metric confirms that while the GAM has enhanced the model's predictive capability, it does so in a way that is comparable but not necessarily superior to the polynomial approach, potentially due to the complexity and variability inherent in housing market data.
+The results of the Generalized Additive Model (GAM) as presented in Table 6 show a significant positive intercept, mirroring the polynomial regression model, which suggests a high base price for houses at the outset of the analysis.
+The R-squared value of 0.506 represents an improvement over the linear models, explaining approximately 50.6% of the variability in house prices.
+This indicates the GAM's enhanced ability to capture the variance in house prices through its non-linear approach.
+
+The Akaike Information Criterion (AIC) and Bayesian Information Criterion (BIC) values for the GAM are slightly lower than those observed in the polynomial regression model, suggesting that the GAM provides a better model fit with less information loss.
+However, the marginal reduction in these values implies that while the GAM enhances the fit of the linear models, the improvement in terms of information criteria is not dramatically significant.
+
+The Root Mean Square Error (RMSE) of the GAM is lower than that of the multiple regression but higher than that observed in the polynomial regression, indicating that the model's predictions are reasonably accurate, yet there may be some potential for enhancement.
+This metric confirms that the GAM has improved the predictive capability of the model, providing an accuracy that is comparable, but not necessarily superior, to the polynomial approach.
+This performance might reflect the complexities and variabilities inherent in the housing market data, which can challenge even sophisticated models like the GAM.
 
@@
 
-Furthermore, two charts in @fig-GAM_fit each showing the smoothed relationship between the predictor and the response variable (house price), with a solid line representing the estimated function and dashed lines indicating the confidence interval around the estimate.
-
-The plot on the left shows the relationship between 'Year' and house prices.
-There’s a clear non-linear trend.
-The prices appear to have a cyclical pattern with a significant peak around 2006, followed by a sharp decline, which likely corresponds to the global financial crisis of 2008.
-After this trough, there’s a recovery with prices trending upwards again, but with a suggestion of leveling off towards the end of the period.
-This could indicate a stabilization in the market or potentially the beginning of another downturn.
-
-The plot on the right shows the relationship between 'NumBedroom' and house prices.
+The two charts in @fig-GAM_fit each depict the smoothed relationship between the predictor and the response variable (house price), with a solid line representing the estimated function and dashed lines indicating the confidence intervals around the estimate.
+
+The plot on the left illustrates the relationship between 'Year' and house prices, revealing a clear non-linear trend.
+The prices exhibit a cyclical pattern with a notable peak around 2006, followed by a sharp decline, which likely corresponds to the global financial crisis of 2008.
+After this trough, there is a recovery, with prices trending upwards again, but with indications of leveling off towards the end of the period.
+This pattern could signify a stabilization in the market or potentially the onset of another downturn.
+
+The plot on the right demonstrates the relationship between 'NumBedroom' and house prices.
 This relationship is more straightforward: as the number of bedrooms increases, so does the house price.
 The upward trend appears to be exponential rather than linear, suggesting that additional bedrooms have an increasingly positive effect on house prices.
-The confidence interval widens with the number of bedrooms, indicating more variability in the data for houses with more bedrooms, which could be due to fewer observations in the higher bedroom categories or greater variance in prices for larger homes.
-
-Both plots reveal the flexibility of GAMs in modeling complex relationships, allowing for the identification of trends that simpler linear models could miss.
-The confidence intervals suggest the model is fairly confident about its estimates, especially for the 'NumBedroom' variable.
-However, the wider intervals at the extremes of the data range would call for cautious interpretation in those areas.
+The confidence interval widens with the number of bedrooms, indicating more variability in the data for houses with more bedrooms.
+This increased variability could be due to fewer observations in the higher bedroom categories or greater variance in prices for larger homes.
+
+Both plots exemplify the flexibility of GAMs in modeling complex relationships, enabling the identification of trends that simpler linear models might overlook.
+The confidence intervals demonstrate that the model is fairly confident in its estimates, particularly for the 'NumBedroom' variable.
+However, the wider intervals at the extremes of the data range suggest a need for cautious interpretation in those areas, reflecting potential uncertainty or the influence of outlier values.
 
@@
 
-In comparing the Multiple Regression, Polynomial Regression, and Generalized Additive Model (GAM) applied to the housing market data, each model serves a distinct purpose based on the complexity of the relationships it can capture between the predictors and the house prices.
-
-The Multiple Regression model, with its linear approach, is the most straightforward, offering a baseline model for comparison.
-It is the least complex and most transparent, which makes it a good starting point for analysis but also the most limited in capturing the nuanced behavior of the housing market.
-The Polynomial Regression model advances this analysis by introducing non-linearity through polynomial terms, providing a significantly better fit as evidenced by its improved R-squared value and lower information criteria scores.
-It acknowledges and captures the more intricate patterns in housing prices related to time and property features.
-
-GAM takes flexibility a step further, not being constrained by predefined polynomial terms, which allows for an even more nuanced understanding of the data.
-This model adapts to the data's inherent trends through the use of smoothing functions, leading to an improved R-squared value compared to the Multiple Regression model.
-However, when compared to Polynomial Regression, the gains in model performance are modest.
-While the GAM model offers a refined approach, it does not significantly outperform the Polynomial Regression in terms of the information criteria and RMSE, suggesting that the increased complexity of GAM may not be necessary for this particular dataset.
-
-However, the visual representation of data trends provided by @fig-GAM_fit is not just intuitively appealing but also offers practical insights that might be obscured in tabular regression outputs.
-For example, the cyclical pattern observed in the 'Year' graph aligns with known economic events that affected housing prices, demonstrating GAM's capacity to capture externalities and temporal dynamics beyond the reach of conventional linear models.
-
-Ultimately, the choice between these models should be informed by the balance between the need for model simplicity and the desire to capture complex data relationships.
-While the Polynomial Regression strikes a balance offering a good fit without overly complicating the model, the GAM provides the most flexibility, which could be essential for more intricate datasets or where the precise nature of relationships between variables is less understood.
+In comparing the Multiple Regression, Polynomial Regression, and Generalized Additive Model (GAM) applied to housing market data, each model serves a distinct purpose based on the complexity of the relationships it can capture between the predictors and house prices.
+
+The Multiple Regression model, with its linear approach, offers a straightforward baseline for comparison.
+It is the simplest and most transparent, making it an ideal starting point for analysis.
+However, it is also the most limited in capturing the nuanced behaviors of the housing market.
+The Polynomial Regression model extends this analysis by introducing non-linearity through polynomial terms, which significantly improves the fit as evidenced by its higher R-squared value and lower information criteria scores.
+This model captures more intricate patterns in housing prices related to time and property features.
+
+The GAM further enhances flexibility, not being constrained by predefined polynomial terms, which allows for an even more nuanced understanding of data trends.
+This model adapts to inherent trends in the data through the use of smoothing functions, leading to an improved R-squared value over the Multiple Regression model.
+However, the gains in model performance when compared to the Polynomial Regression are modest.
+While the GAM offers a refined approach, it does not significantly outperform the Polynomial Regression in terms of information criteria and RMSE, suggesting that the increased complexity of the GAM may not be necessary for this particular dataset.
+
+Nevertheless, the visual representation of data trends provided by @fig-GAM_fit offers practical insights that might be obscured in tabular regression outputs.
+For example, the cyclical pattern observed in the 'Year' graph aligns with known economic events that affected housing prices, demonstrating GAM's ability to capture externalities and temporal dynamics beyond what conventional linear models can offer.
+
+The choice between these models should be informed by the balance between the need for simplicity and the desire to capture complex data relationships.
+While the Polynomial Regression strikes a balance, offering a good fit without overly complicating the model, the GAM provides the most flexibility, which could be crucial for analyzing more intricate datasets or where the precise nature of relationships between variables is less understood.
 
@@
 
-## 1
-
-## 2
-
-## 3
-
-## 4
+## Global Trends
+
+The consistent rise in housing prices documented in this research parallels global economic patterns observed across developed nations [@citeGlobalHP].
+This upward trend is not an isolated phenomenon but part of a broader pattern evident across international borders, highlighting a confluence of macroeconomic and microeconomic factors.
+This research emphasizes the deep interconnection between the housing market and global economic conditions, illustrating the intrinsic relationship between real estate values and global economic health.
+
+The escalation in prices may coincide with inflationary trends affecting various economic sectors, suggesting that housing markets are susceptible to the impacts of global monetary policies and financial climates.
+These inflationary pressures can distort consumer purchasing power, creating a complex scenario where rising prices may not necessarily reflect actual increases in property value.
+The data also suggests that demographic shifts, particularly changes in family structures with more members per household, significantly influence demand and, consequently, prices in the housing market.
+
+Furthermore, the continued increase in house prices in the US has spurred an intriguing area of study.
+As other research indicates, the GDP of the US has not seen clear growth for an extended period [@citeUSGDP].
+It would typically be assumed that without increased earnings, consumer purchasing would not rise; however, the rising house prices contradict this expectation.
+This discrepancy poses critical questions about the factors driving housing market dynamics, independent of direct economic growth.
+
+## Potential Preference Shifts
+
+The increase in demand for larger homes, as indicated by the data, suggests a potential shift in housing preferences, possibly spurred by the recent pandemic, which emphasized the need for multifunctional living spaces that combine comfort with functionality.
+This shift reflects not only a preference for larger living spaces but also an adaptation to new realities—homes that can accommodate remote work, virtual schooling, and indoor recreation.
+
+This evolving preference could have profound implications, affecting urban planning and infrastructure development.
+Urban areas, known for high-density living, may need to reconsider zoning laws and development guidelines to meet the growing demand for larger homes.
+This trend might also influence the architecture and design industries, inspiring innovations in home layouts, building materials, and community planning to cater to new living standards.
+
+Additionally, the preference for larger homes could intensify economic disparities, as the ability to afford such spaces is closely linked to economic status.
+This dynamic presents a significant challenge for economic planners and policymakers who must ensure that the housing market remains accessible and that housing policies do not exclude those unable to afford larger homes.
+
+Economically, this trend could boost industries related to home construction, renovation, and furnishing as consumers seek to enhance or acquire properties that match their new preferences.
+However, it also calls for a careful evaluation of the environmental impacts, as larger homes generally require more resources and energy, posing challenges to sustainability goals [@citeUKPollution].
+This balance between accommodating consumer preferences and maintaining sustainability will be crucial for future policy and economic strategies in the housing sector.
+
+## Geographical Position
+
+The geographical analysis of housing price trends, illustrated through heat maps, reveals significant disparities across different regions, depicting an uneven economic landscape.
+Particularly, coastal regions display elevated housing prices, embodying the enduring principle in real estate that location is paramount.
+These regions, often centers of economic and cultural activity, have experienced increasing demand due to urban expansion and the attractive notion of coastal living.
+This demand, coupled with the inherently limited availability of land, leads to a competitive market with steadily rising prices.
+
+The insights derived from the geographical distribution of housing markets highlight opportunities for strategic infrastructural investments.
+Identifying areas with rapid price growth allows stakeholders to pinpoint locations primed for development or in need of policy adjustments to stabilize the market.
+Additionally, this data facilitates the targeted allocation of resources to underserved regions, promoting a more balanced economic distribution and enhancing overall economic equity.
+
+## Possible Actions
+
+The insights from this analysis provide a valuable toolset for policymakers, urban planners, and investors, enabling informed decision-making aimed at stabilizing and guiding the housing market.
+Strategic policy interventions, designed based on the disparities and dynamics revealed by the data, can ensure that the housing market contributes positively to economic stability and community welfare.
+
+For policymakers, the rising housing prices and regional disparities highlight the pressing need for robust housing policies that enhance affordability and accessibility.
+Government initiatives could include subsidies for first-time homebuyers, incentives for developers to build more affordable housing units, and stricter regulations on speculative real estate investments that inflate prices.
+Moreover, tax incentives or revisions to zoning laws could promote the development of underused areas, helping to relieve pressure on overheated markets.
+
+The measures implemented in response to these findings must be adaptable and responsive to the shifting patterns of the market.
+It will be essential to regularly reassess housing policies and investment strategies to respond to new data and changing economic conditions, ensuring that the housing market continues to serve as a pillar of economic stability and a source of equitable opportunities.
+This proactive approach will help mitigate future market volatilities and foster a robust, inclusive housing market.
 
@@
 
+The current study provides a detailed analysis, yet there is scope for further refinement.
+Future research could benefit from incorporating additional variables such as interest rates, housing supply metrics, and construction costs, which could provide a more detailed explanation of price variability.
+Examining microeconomic factors such as individual income levels and the impact of local zoning laws could yield deeper insights into market dynamics.
+Additionally, longitudinal studies that assess the long-term effects of policy changes on housing prices would be highly valuable.
+Technological advancements in data collection, including real-time tracking of market listings and transactions, could offer more immediate insights into market fluctuations.
+Expanding the research to include rental markets and the affordable housing sector would enhance the understanding of the broader housing ecosystem.
+
 # Conclusion {#sec-con}
+
+This paper has conducted an in-depth analysis of the U.S.
+residential real estate market, uncovering the intricate factors that influence housing prices.
+The study tracked price trends across various states and types of housing, noting significant increases over the past two decades.
+It demonstrates that house prices are affected not only by traditional factors such as location and property characteristics but also by broader economic indicators and potential shifts in societal preferences.
+
+Our findings indicate that while the U.S.
+housing market possesses unique attributes, it also aligns with global market trends, especially regarding the rising prices of larger homes.
+This increase may indicate a shift in consumer preferences, possibly a reaction to the pandemic, towards more spacious living arrangements.
+Moreover, the geographical analysis reveals a stark disparity in housing costs, with coastal regions showing considerably higher prices, underscoring the critical importance of location in property valuation.
+
+While this study is detailed, there is room for further enhancement and expansion.
+Future research should incorporate additional economic variables, such as interest rates and housing supply data, to offer a more comprehensive understanding of the market forces involved.
+Longitudinal studies that examine the effects of policy changes over time could also enrich our understanding of market dynamics.
+Advances in data collection technology could enable real-time analysis of market trends, allowing for a more responsive approach to rapid changes in the market.
+Additionally, broadening the scope to include rental markets and affordable housing would provide insights into the entire housing spectrum, crucial for developing holistic housing strategies that benefit all societal segments.
+
+In summary, the U.S.
+housing market is dynamic and complex, shaped by a wide range of local and global factors.
+This paper lays the groundwork for a deeper understanding of these factors, offering a fundamental analysis that stakeholders can utilize to make informed decisions that impact community life across the nation.
 

### 2024-04-17 scripts/01-data_processing.R
scripts updated

@@ -1,0 +1,308 @@
+#### Preamble ####
+# Purpose: This code loads, cleans, aggregates, and saves various datasets of U.S. house prices, removing irrelevant rows and columns, and summarizing the data by state on a monthly and yearly basis, then stores the processed data for further analysis.
+# Author: Ziheng Zhong
+# Date: 02 April 2024
+# Contact: ziheng.zhong@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: none
+
+
+
+#### Workspace setup ####
+library(dplyr)
+library(tidyverse)
+library(lubridate)
+
+
+
+#### load data ####
+all_house_price <- read_csv("../../inputs/data/all_house_price.csv")
+
+
+
+#### Cleaning ####
+clean_house_price <- all_house_price %>% # Delete the first row since it cause issues in further analysis (we focus on state but the first row is data for U.S. as a whole)
+  slice(-1)
+
+clean_house_price <- clean_house_price %>% # Remove specific columns
+  select(-c(SizeRank, RegionID, RegionType, RegionName))
+
+#### Further Cleaning ####
+# By State - monthly
+state_house_price_by_month <- clean_house_price %>%
+  group_by(StateName) %>%
+  summarise(across(where(is.numeric), mean, na.rm = TRUE)) # Calculating means house price of each state excluding NA
+
+# By State - yearly
+state_house_price_by_year <- state_house_price_by_month %>%
+  pivot_longer( # Gather the monthly columns into key-value pairs, convert the keys to Dates, extract the year, and then calculate the mean
+    cols = starts_with("20"), # Assuming all your date columns start with '20'
+    names_to = "Month",
+    values_to = "HousePrice"
+  ) %>%
+  mutate(Year = year(ymd(Month))) %>%
+  group_by(StateName, Year) %>%
+  summarise(AvgHousePrice = mean(HousePrice, na.rm = TRUE), .groups = 'drop') # Calculating mean house price by year for each state
+
+
+
+#### save data ####
+write_csv(
+  x = clean_house_price,
+  file = "outputs/data/clean_house_price.csv"
+)
+
+write_csv(
+  x = state_house_price_by_month,
+  file = "outputs/data/state_house_price_by_month.csv"
+)
+
+write_csv(
+  x = state_house_price_by_year,
+  file = "outputs/data/state_house_price_by_year.csv"
+)
+
+
+
+#### load data ####
[…]

### 2024-04-17 scripts/01_00-data_cleaning.R

@@ -1,63 +1,0 @@
-#### Preamble ####
-# Purpose: This code is aimed at cleaning and restructuring a comprehensive dataset of house prices, focusing on aggregating the data by state both on a monthly and yearly basis, followed by saving these cleaned and aggregated datasets for further analysis or visualization.
-# Author: Ziheng Zhong
-# Date: 02 April 2024
-# Contact: ziheng.zhong@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: none
-
-
-
-#### Workspace setup ####
-library(dplyr)
-library(tidyverse)
-library(lubridate)
-
-
-
-#### load data ####
-raw_data <- read_csv("../../inputs/data/all_house_price.csv")
-
-
-
-#### Cleaning ####
-clean_house_price <- raw_data %>% # Delete the first row since it cause issues in further analysis (we focus on state but the first row is data for U.S. as a whole)
-  slice(-1)
-
-clean_house_price <- clean_house_price %>% # Remove specific columns
-  select(-c(SizeRank, RegionID, RegionType, RegionName))
-
-#### Further Cleaning ####
-# By State - monthly
-state_house_price_by_month <- clean_house_price %>%
-  group_by(StateName) %>%
-  summarise(across(where(is.numeric), mean, na.rm = TRUE)) # Calculating means house price of each state excluding NA
-
-# By State - yearly
-state_house_price_by_year <- state_house_price_by_month %>%
-  pivot_longer( # Gather the monthly columns into key-value pairs, convert the keys to Dates, extract the year, and then calculate the mean
-    cols = starts_with("20"), # Assuming all your date columns start with '20'
-    names_to = "Month",
-    values_to = "HousePrice"
-  ) %>%
-  mutate(Year = year(ymd(Month))) %>%
-  group_by(StateName, Year) %>%
-  summarise(AvgHousePrice = mean(HousePrice, na.rm = TRUE), .groups = 'drop') # Calculating mean house price by year for each state
-
-
-
-#### save data ####
-write_csv(
-  x = clean_house_price,
-  file = "outputs/data/clean_house_price.csv"
-)
-
-write_csv(
-  x = state_house_price_by_month,
-  file = "outputs/data/state_house_price_by_month.csv"
-)
-
-write_csv(
-  x = state_house_price_by_year,
-  file = "outputs/data/state_house_price_by_year.csv"
-)

### 2024-04-17 scripts/01_01-data_cleaning.R

@@ -1,63 +1,0 @@
-#### Preamble ####
-# Purpose: This code is aimed at cleaning and restructuring a comprehensive dataset of house prices, focusing on aggregating the data by state both on a monthly and yearly basis, followed by saving these cleaned and aggregated datasets for further analysis or visualization.
-# Author: Ziheng Zhong
-# Date: 02 April 2024
-# Contact: ziheng.zhong@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: none
-
-
-
-#### Workspace setup ####
-library(dplyr)
-library(tidyverse)
-library(lubridate)
-
-
-
-#### load data ####
-raw_data <- read_csv("inputs/data/1b_house_price.csv")
-
-
-
-#### Cleaning ####
-clean_house_price <- raw_data %>% # Delete the first row since it cause issues in further analysis (we focus on state but the first row is data for U.S. as a whole)
-  slice(-1)
-
-clean_house_price <- clean_house_price %>% # Remove specific columns
-  select(-c(SizeRank, RegionID, RegionType, RegionName))
-
-#### Further Cleaning ####
-# By State - monthly
-state_house_price_by_month <- clean_house_price %>%
-  group_by(StateName) %>%
-  summarise(across(where(is.numeric), mean, na.rm = TRUE)) # Calculating means house price of each state excluding NA
-
-# By State - yearly
-state_house_price_by_year <- state_house_price_by_month %>%
-  pivot_longer( # Gather the monthly columns into key-value pairs, convert the keys to Dates, extract the year, and then calculate the mean
-    cols = starts_with("20"), # Assuming all your date columns start with '20'
-    names_to = "Month",
-    values_to = "HousePrice"
-  ) %>%
-  mutate(Year = year(ymd(Month))) %>%
-  group_by(StateName, Year) %>%
-  summarise(AvgHousePrice = mean(HousePrice, na.rm = TRUE), .groups = 'drop') # Calculating mean house price by year for each state
-
-
-
-#### save data ####
-write_csv(
-  x = clean_house_price,
-  file = "outputs/data/clean_1b_house_price.csv"
-)
-
-write_csv(
-  x = state_house_price_by_month,
-  file = "outputs/data/state_1b_house_price_by_month.csv"
-)
-
-write_csv(
-  x = state_house_price_by_year,
-  file = "outputs/data/state_1b_house_price_by_year.csv"
-)

### 2024-04-17 scripts/01_02-data_cleaning.R

@@ -1,63 +1,0 @@
-#### Preamble ####
-# Purpose: This code is aimed at cleaning and restructuring a comprehensive dataset of house prices, focusing on aggregating the data by state both on a monthly and yearly basis, followed by saving these cleaned and aggregated datasets for further analysis or visualization.
-# Author: Ziheng Zhong
-# Date: 02 April 2024
-# Contact: ziheng.zhong@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: none
-
-
-
-#### Workspace setup ####
-library(dplyr)
-library(tidyverse)
-library(lubridate)
-
-
-
-#### load data ####
-raw_data <- read_csv("inputs/data/2b_house_price.csv")
-
-
-
-#### Cleaning ####
-clean_house_price <- raw_data %>% # Delete the first row since it cause issues in further analysis (we focus on state but the first row is data for U.S. as a whole)
-  slice(-1)
-
-clean_house_price <- clean_house_price %>% # Remove specific columns
-  select(-c(SizeRank, RegionID, RegionType, RegionName))
-
-#### Further Cleaning ####
-# By State - monthly
-state_house_price_by_month <- clean_house_price %>%
-  group_by(StateName) %>%
-  summarise(across(where(is.numeric), mean, na.rm = TRUE)) # Calculating means house price of each state excluding NA
-
-# By State - yearly
-state_house_price_by_year <- state_house_price_by_month %>%
-  pivot_longer( # Gather the monthly columns into key-value pairs, convert the keys to Dates, extract the year, and then calculate the mean
-    cols = starts_with("20"), # Assuming all your date columns start with '20'
-    names_to = "Month",
-    values_to = "HousePrice"
-  ) %>%
-  mutate(Year = year(ymd(Month))) %>%
-  group_by(StateName, Year) %>%
-  summarise(AvgHousePrice = mean(HousePrice, na.rm = TRUE), .groups = 'drop') # Calculating mean house price by year for each state
-
-
-
-#### save data ####
-write_csv(
-  x = clean_house_price,
-  file = "outputs/data/clean_2b_house_price.csv"
-)
-
-write_csv(
-  x = state_house_price_by_month,
-  file = "outputs/data/state_2b_house_price_by_month.csv"
-)
-
-write_csv(
-  x = state_house_price_by_year,
-  file = "outputs/data/state_2b_house_price_by_year.csv"
-)

### 2024-04-17 scripts/01_03-data_cleaning.R

@@ -1,63 +1,0 @@
-#### Preamble ####
-# Purpose: This code is aimed at cleaning and restructuring a comprehensive dataset of house prices, focusing on aggregating the data by state both on a monthly and yearly basis, followed by saving these cleaned and aggregated datasets for further analysis or visualization.
-# Author: Ziheng Zhong
-# Date: 02 April 2024
-# Contact: ziheng.zhong@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: none
-
-
-
-#### Workspace setup ####
-library(dplyr)
-library(tidyverse)
-library(lubridate)
-
-
-
-#### load data ####
-raw_data <- read_csv("inputs/data/3b_house_price.csv")
-
-
-
-#### Cleaning ####
-clean_house_price <- raw_data %>% # Delete the first row since it cause issues in further analysis (we focus on state but the first row is data for U.S. as a whole)
-  slice(-1)
-
-clean_house_price <- clean_house_price %>% # Remove specific columns
-  select(-c(SizeRank, RegionID, RegionType, RegionName))
-
-#### Further Cleaning ####
-# By State - monthly
-state_house_price_by_month <- clean_house_price %>%
-  group_by(StateName) %>%
-  summarise(across(where(is.numeric), mean, na.rm = TRUE)) # Calculating means house price of each state excluding NA
-
-# By State - yearly
-state_house_price_by_year <- state_house_price_by_month %>%
-  pivot_longer( # Gather the monthly columns into key-value pairs, convert the keys to Dates, extract the year, and then calculate the mean
-    cols = starts_with("20"), # Assuming all your date columns start with '20'
-    names_to = "Month",
-    values_to = "HousePrice"
-  ) %>%
-  mutate(Year = year(ymd(Month))) %>%
-  group_by(StateName, Year) %>%
-  summarise(AvgHousePrice = mean(HousePrice, na.rm = TRUE), .groups = 'drop') # Calculating mean house price by year for each state
-
-
-
-#### save data ####
-write_csv(
-  x = clean_house_price,
-  file = "outputs/data/clean_3b_house_price.csv"
-)
-
-write_csv(
-  x = state_house_price_by_month,
-  file = "outputs/data/state_3b_house_price_by_month.csv"
-)
-
-write_csv(
-  x = state_house_price_by_year,
-  file = "outputs/data/state_3b_house_price_by_year.csv"
-)

### 2024-04-17 scripts/01_04-data_cleaning.R

@@ -1,63 +1,0 @@
-#### Preamble ####
-# Purpose: This code is aimed at cleaning and restructuring a comprehensive dataset of house prices, focusing on aggregating the data by state both on a monthly and yearly basis, followed by saving these cleaned and aggregated datasets for further analysis or visualization.
-# Author: Ziheng Zhong
-# Date: 02 April 2024
-# Contact: ziheng.zhong@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: none
-
-
-
-#### Workspace setup ####
-library(dplyr)
-library(tidyverse)
-library(lubridate)
-
-
-
-#### load data ####
-raw_data <- read_csv("inputs/data/4b_house_price.csv")
-
-
-
-#### Cleaning ####
-clean_house_price <- raw_data %>% # Delete the first row since it cause issues in further analysis (we focus on state but the first row is data for U.S. as a whole)
-  slice(-1)
-
-clean_house_price <- clean_house_price %>% # Remove specific columns
-  select(-c(SizeRank, RegionID, RegionType, RegionName))
-
-#### Further Cleaning ####
-# By State - monthly
-state_house_price_by_month <- clean_house_price %>%
-  group_by(StateName) %>%
-  summarise(across(where(is.numeric), mean, na.rm = TRUE)) # Calculating means house price of each state excluding NA
-
-# By State - yearly
-state_house_price_by_year <- state_house_price_by_month %>%
-  pivot_longer( # Gather the monthly columns into key-value pairs, convert the keys to Dates, extract the year, and then calculate the mean
-    cols = starts_with("20"), # Assuming all your date columns start with '20'
-    names_to = "Month",
-    values_to = "HousePrice"
-  ) %>%
-  mutate(Year = year(ymd(Month))) %>%
-  group_by(StateName, Year) %>%
-  summarise(AvgHousePrice = mean(HousePrice, na.rm = TRUE), .groups = 'drop') # Calculating mean house price by year for each state
-
-
-
-#### save data ####
-write_csv(
-  x = clean_house_price,
-  file = "outputs/data/clean_4b_house_price.csv"
-)
-
-write_csv(
-  x = state_house_price_by_month,
-  file = "outputs/data/state_4b_house_price_by_month.csv"
-)
-
-write_csv(
-  x = state_house_price_by_year,
-  file = "outputs/data/state_4b_house_price_by_year.csv"
-)

### 2024-04-17 scripts/01_05-data_cleaning.R

@@ -1,63 +1,0 @@
-#### Preamble ####
-# Purpose: This code is aimed at cleaning and restructuring a comprehensive dataset of house prices, focusing on aggregating the data by state both on a monthly and yearly basis, followed by saving these cleaned and aggregated datasets for further analysis or visualization.
-# Author: Ziheng Zhong
-# Date: 02 April 2024
-# Contact: ziheng.zhong@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: none
-
-
-
-#### Workspace setup ####
-library(dplyr)
-library(tidyverse)
-library(lubridate)
-
-
-
-#### load data ####
-raw_data <- read_csv("inputs/data/5bplus_house_price.csv")
-
-
-
-#### Cleaning ####
-clean_house_price <- raw_data %>% # Delete the first row since it cause issues in further analysis (we focus on state but the first row is data for U.S. as a whole)
-  slice(-1)
-
-clean_house_price <- clean_house_price %>% # Remove specific columns
-  select(-c(SizeRank, RegionID, RegionType, RegionName))
-
-#### Further Cleaning ####
-# By State - monthly
-state_house_price_by_month <- clean_house_price %>%
-  group_by(StateName) %>%
-  summarise(across(where(is.numeric), mean, na.rm = TRUE)) # Calculating means house price of each state excluding NA
-
-# By State - yearly
-state_house_price_by_year <- state_house_price_by_month %>%
-  pivot_longer( # Gather the monthly columns into key-value pairs, convert the keys to Dates, extract the year, and then calculate the mean
-    cols = starts_with("20"), # Assuming all your date columns start with '20'
-    names_to = "Month",
-    values_to = "HousePrice"
-  ) %>%
-  mutate(Year = year(ymd(Month))) %>%
-  group_by(StateName, Year) %>%
-  summarise(AvgHousePrice = mean(HousePrice, na.rm = TRUE), .groups = 'drop') # Calculating mean house price by year for each state
-
-
-
-#### save data ####
-write_csv(
-  x = clean_house_price,
-  file = "outputs/data/clean_5bplus_house_price.csv"
-)
-
-write_csv(
-  x = state_house_price_by_month,
-  file = "outputs/data/state_5bplus_house_price_by_month.csv"
-)
-
-write_csv(
-  x = state_house_price_by_year,
-  file = "outputs/data/state_5bplus_house_price_by_year.csv"
-)

### 2024-04-17 scripts/04-charts.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: The code utilizes R libraries for data manipulation, visualization, and mapping to generate heat maps of average house prices by state in the US, categorized by the number of bedrooms, using datasets on yearly house prices.
+# Purpose: This code loads multiple datasets of U.S. house prices, converts state abbreviations to full names, generates heat maps for different types of houses, and plots the overall and bedroom-specific price trends by month.
 # Author: Ziheng Zhong
 # Date: 02 April 2024
 # Contact: ziheng.zhong@mail.utoronto.ca
@@ -179,3 +179,65 @@ ggplot() + # Generating heat map
        subtitle = "Data: Your Dataset",
        caption = "Visualization by Example") +
   theme(legend.position = "right")
+
+
+
+#### 5 Types of Houses Price Trend by Month ####
+process_data <- function(data, label) { # Function to process each dataset
+  data %>% 
+    select(-StateName) %>%
+    summarise_all(mean, na.rm = TRUE) %>%
+    pivot_longer(everything(), names_to = "Date", values_to = "AverageHousePrice") %>%
+    mutate(Trend = label,
+           Date = as.Date(Date))
+}
+
+data_list <- list(
+  `1B` = state_1b_house_price_by_month,
+  `2B` = state_2b_house_price_by_month,
+  `3B` = state_3b_house_price_by_month,
+  `4B` = state_4b_house_price_by_month,
+  `5B+` = state_5bplus_house_price_by_month
+)
+
+processed_data <- lapply(names(data_list), function(name) process_data(data_list[[name]], name))
+
+combined_data <- bind_rows(processed_data)
+
+ggplot(combined_data, aes(x = Date, y = AverageHousePrice, color = Trend)) +
+  geom_line() + 
+  labs(x = "Year", y = "Average House Price ($)") +
+  theme_minimal() + 
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_color_viridis_d(name = "Number of Bedrooms")
+
+
+
+#### All Types of House Price Trend by Month ####
+state_all <- state_house_price_by_month %>% select(-StateName) %>%  # Processing datasets
+  summarise_all(mean, na.rm = TRUE) %>%
+  pivot_longer(everything(), names_to = "Date", values_to = "AverageHousePrice")
+
[…]

### 2024-04-17 scripts/04_01-charts.R

@@ -1,92 +1,0 @@
-#### Preamble ####
-# Purpose: The code is designed to process and visualize trends in average house prices by month for different house types using R, integrating data manipulation and visualization libraries to prepare and plot the data.
-# Author: Ziheng Zhong
-# Date: 02 April 2024
-# Contact: ziheng.zhong@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: none
-
-
-
-#### Workspace setup ####
-library(maps)
-library(mgcv)
-library(ggmap)
-library(readr)
-library(tidyr)
-library(dplyr)
-library(ggplot2)
-library(viridis)
-library(lubridate)
-
-
-
-#### load data ####
-state_house_price_by_month <- read_csv("outputs/data/state_house_price_by_month.csv")
-state_1b_house_price_by_month <- read_csv("outputs/data/state_1b_house_price_by_month.csv")
-state_2b_house_price_by_month <- read_csv("outputs/data/state_2b_house_price_by_month.csv")
-state_3b_house_price_by_month <- read_csv("outputs/data/state_3b_house_price_by_month.csv")
-state_4b_house_price_by_month <- read_csv("outputs/data/state_4b_house_price_by_month.csv")
-state_5bplus_house_price_by_month <- read_csv("outputs/data/state_5bplus_house_price_by_month.csv")
-
-
-
-#### 5 Types of Houses Price Trend by Month ####
-process_data <- function(data, label) { # Function to process each dataset
-  data %>% 
-    select(-StateName) %>%
-    summarise_all(mean, na.rm = TRUE) %>%
-    pivot_longer(everything(), names_to = "Date", values_to = "AverageHousePrice") %>%
-    mutate(Trend = label,
-           Date = as.Date(Date))
-}
-
-data_list <- list(
-  `1B` = state_1b_house_price_by_month,
-  `2B` = state_2b_house_price_by_month,
-  `3B` = state_3b_house_price_by_month,
-  `4B` = state_4b_house_price_by_month,
-  `5B+` = state_5bplus_house_price_by_month
-)
-
-processed_data <- lapply(names(data_list), function(name) process_data(data_list[[name]], name))
-
-combined_data <- bind_rows(processed_data)
-
-ggplot(combined_data, aes(x = Date, y = AverageHousePrice, color = Trend)) +
-  geom_line() + 
-  labs(x = "Year", y = "Average House Price ($)") +
-  theme_minimal() + 
[…]

### 2024-04-17 scripts/02-data_merging.R
scripts updated

@@ -9,6 +9,7 @@
 
 
 #### Workspace setup ####
+library(arrow)
 library(dplyr)
 library(tidyverse)
 library(lubridate)
@@ -16,11 +17,13 @@ library(lubridate)
 
 
 #### load data ####
-state_1b_house_price_by_year <- read_csv("../../outputs/data/state_1b_house_price_by_year.csv")
-state_2b_house_price_by_year <- read_csv("../../outputs/data/state_2b_house_price_by_year.csv")
-state_3b_house_price_by_year <- read_csv("../../outputs/data/state_3b_house_price_by_year.csv")
-state_4b_house_price_by_year <- read_csv("../../outputs/data/state_4b_house_price_by_year.csv")
-state_5bplus_house_price_by_year <- read_csv("../../outputs/data/state_5bplus_house_price_by_year.csv")
+setwd("/Users/ijustin/Library/CloudStorage/Dropbox/Code/STA302/Research_Papers/House_Price")
+
+state_1b_house_price_by_year <- read_csv("outputs/data/state_1b_house_price_by_year.csv")
+state_2b_house_price_by_year <- read_csv("outputs/data/state_2b_house_price_by_year.csv")
+state_3b_house_price_by_year <- read_csv("outputs/data/state_3b_house_price_by_year.csv")
+state_4b_house_price_by_year <- read_csv("outputs/data/state_4b_house_price_by_year.csv")
+state_5bplus_house_price_by_year <- read_csv("outputs/data/state_5bplus_house_price_by_year.csv")
 
 
 
@@ -49,8 +52,7 @@ state_5bplus_house_price_by_year <- state_5bplus_house_price_by_year %>%
 merge_data <- bind_rows(state_1b_house_price_by_year, state_2b_house_price_by_year, state_3b_house_price_by_year, state_4b_house_price_by_year, state_5bplus_house_price_by_year)
 merge_data <- na.omit(merge_data)
 
-#### save data ####
-write_csv(
-  x = merge_data,
-  file = "outputs/data/merged_house_price.csv"
-)
+
+
+#### save data as Parquet ####
+write_parquet(merge_data, "outputs/data/merged_house_price.parquet")

### 2024-04-17 scripts/03-modeling.R

@@ -19,7 +19,7 @@ library(modelsummary)
 
 
 #### load data ####
-house_prices <- read_csv("../../outputs/data/merged_house_price.csv")
+house_prices <- read_csv("../../outputs/data/merged_house_price.parquet")
 
 
 

### 2024-04-17 outputs/paper/paper.qmd
paper updated

@@
 # load packages
-packages <- c("tidyverse", "ggplot2", "janitor", "readr", "knitr", "modelsummary", "testthat", "kableExtra", "viridis", "lubridate", "maps", "mgcv")
+packages <- c("tidyverse", "ggplot2", "janitor", "readr", "knitr", "modelsummary", "testthat", "kableExtra", "viridis", "lubridate", "maps", "mgcv", "arrow")
 
@@
 all_house_price <- read_csv("../../inputs/data/all_house_price.csv")
-house_prices <- read_csv("../../outputs/data/merged_house_price.csv")
+house_prices <- read_parquet("../../outputs/data/merged_house_price.parquet")
 
@@
 Data used in this paper was cleaned, processed and tested with the programming language R [@citeR].
-Also with support of additional packages in R: `tidyverse` [@citeTidyverse], `ggplot2` [@citeGgplot], `janitor` [@citeJanitor], `readr` [@citeReadr], `knitr` [@citeKnitr], `modelsummary` [@citeModelsummary], `testthat` [@citeTestthat], `KableExtra` [@citeKableEx], `viridis` [@citeViridis], `lubridate` [@citeLubridate], `maps` [@citeMaps], `mgcv` [@citeMgcv].
+Also with support of additional packages in R: `tidyverse` [@citeTidyverse], `ggplot2` [@citeGgplot], `janitor` [@citeJanitor], `readr` [@citeReadr], `knitr` [@citeKnitr], `modelsummary` [@citeModelsummary], `testthat` [@citeTestthat], `KableExtra` [@citeKableEx], `viridis` [@citeViridis], `lubridate` [@citeLubridate], `maps` [@citeMaps], `mgcv` [@citeMgcv], `Arrow` [@citeArrow].
 
@@
 
-The results of the Generalized Additive Model (GAM) as presented in Table 6 show a significant positive intercept, mirroring the polynomial regression model, which suggests a high base price for houses at the outset of the analysis.
+The results of the Generalized Additive Model (GAM) as presented in @tbl-nonlinear_model_result show a significant positive intercept, mirroring the polynomial regression model, which suggests a high base price for houses at the outset of the analysis.
 The R-squared value of 0.506 represents an improvement over the linear models, explaining approximately 50.6% of the variability in house prices.
@@
 
-The two charts in @fig-GAM_fit each depict the smoothed relationship between the predictor and the response variable (house price), with a solid line representing the estimated function and dashed lines indicating the confidence intervals around the estimate.
+In @fig-GAM_fit, the numbers in the parentheses, such as 10.83 and 2.74, represent the smoothing parameters or effective degrees of freedom.
+These parameters control the flexibility of the spline, balancing the fit of the model against the risk of overfitting by adjusting the "wiggliness" of the curve.
+The values on the y-axis, which range from negative to positive, quantify the contribution of each smooth term to the predicted value of the response variable, such as house prices.
+For example, these y-axis values illustrate how changes in the year or the number of bedrooms influence the predicted house price, relative to a certain baseline.
+
+The solid line depicted in the plot is the estimated smooth curve, which shows the modeled relationship between the predictor and the response variable.
+The dashed lines surrounding the solid line typically represent confidence intervals, providing a range within which the true effect is likely to lie with a certain level of confidence.
+This visualization helps in understanding the certainty of the model predictions and the potential variability in the effect of each predictor on the response variable.
 

### 2024-04-17 outputs/paper/paper.qmd
paper updated

@@
 date-format: long
-abstract: "This research examines the dynamics of the U.S.
-housing market, offering a detailed analysis of price trends across various states and types of housing, which have shown considerable increases in recent years.
-The study clarifies how factors such as location, property characteristics, and external economic influences impact house prices."
+abstract: "This research conducts a detailed analysis of U.S.
+residential real estate price trends over the past two decades, using data from Zillow to explore influences from geographic location, property characteristics, and economic indicators.
+By employing multiple regression models, polynomial regression, and Generalized Additive Models (GAM), the study identifies complex interactions that affect housing prices.
+It highlights a notable shift towards larger homes, potentially driven by changes in work and lifestyle patterns due to recent global events."
 format: pdf

### 2024-04-17 outputs/paper/paper.qmd
paper final draft

@@
 
-Data used in this paper was cleaned, processed and tested with the programming language R [@citeR].
+Data used in this paper was cleaned, processed, modeled and tested with the programming language R [@citeR].
 Also with support of additional packages in R: `tidyverse` [@citeTidyverse], `ggplot2` [@citeGgplot], `janitor` [@citeJanitor], `readr` [@citeReadr], `knitr` [@citeKnitr], `modelsummary` [@citeModelsummary], `testthat` [@citeTestthat], `KableExtra` [@citeKableEx], `viridis` [@citeViridis], `lubridate` [@citeLubridate], `maps` [@citeMaps], `mgcv` [@citeMgcv], `Arrow` [@citeArrow].
@@
 This data repository provides a broad array of real estate statistics, capturing the dynamics of house prices across various regions in the United States.
+The dataset encompasses detailed information on house prices from 2000 to 2024, including breakdowns by state, city, and occasionally neighborhood levels, from median list prices to the Zillow Home Value Index (ZHVI).
 As a cornerstone of transparency and public engagement, Zillow Research offers open access to these vital real estate data, which are instrumental in understanding housing market behaviors.
-
-The dataset encompasses detailed information on house prices over several years, including breakdowns by state, city, and occasionally neighborhood levels, from median list prices to the Zillow Home Value Index (ZHVI).
-This level of detail facilitates a nuanced analysis of house price trends and the identification of significant market dynamics, which can guide potential real estate investments and policy formulations.
 
@@
 
+## Measurement
+
+In this study, the conversion of real-world real estate transactions into quantifiable data entries is meticulously executed using Zillow's extensive datasets.
+Each real-world transaction is represented as a data point in our analysis, capturing essential attributes such as sale price, property location, and characteristics like the number of bedrooms and bathrooms.
+For example, a house sale in Miami with three bedrooms and two bathrooms, sold at a specific price, is recorded with precise values for each attribute, including the sale date.
+This method ensures that each entry in our dataset accurately reflects an actual event, maintaining the integrity and accuracy of the data.
+The rigorous methodology employed in documenting these transactions allows for precise, data-driven insights into market dynamics, providing a solid foundation for analyzing trends and drawing conclusions about factors influencing real estate prices.
+
 ## Method
 
-Our analysis began with the critical task of refining the raw dataset downloaded from Zillow.
+Our analysis began with the essential task of refining the raw dataset downloaded from Zillow.
+The original data was largely retained because it covers a time span that aligns well with the needs of this study.
 The initial phase involved understanding the dataset’s structure.
-As shown in @tbl-raw_data, this dataset typically includes multiple attributes such as Region ID, Region Type, location (states), year of transaction, house price, etc.
-Each record represents a unique property, providing a detailed snapshot of its characteristics.
+As demonstrated in @tbl-raw_data, this dataset typically includes multiple attributes such as Region ID, Region Type, location (states), year of transaction, house price, etc.
+Each record represents a unique property, offering a detailed snapshot of its characteristics.
 
@@
 
-$$\text{Price} = \beta_0 + \beta_1 \times \text{Year} + \beta_2 \times \text{NumBedroom}$$
+$$\text{Price} = \beta_0 + \beta_1 \times \text{Year} + \beta_2 \times \text{NumBedroom} + \varepsilon$$
 
@@
 
-$$\text{Price} = \beta_0 + \beta_1 \times \text{Year} + \beta_2 \times \text{Year}^2 + \beta_3 \times \text{NumBedroom} + \beta_4 \times \text{NumBedroom}^2$$
+$$\text{Price} = \beta_0 + \beta_1 \times \text{Year} + \beta_2 \times \text{Year}^2 + \beta_3 \times \text{NumBedroom} + \beta_4 \times \text{NumBedroom}^2 + \varepsilon$$
 
@@
 
-$$\text{Price} = -13,969,628.019 + 6,958.769 \times \text{Year} + 59,897.327 \times \text{NumBedroom}$$
+$$\hat{\text{Price}}  = -13,969,628.019 + 6,958.769 \times \text{Year} + 59,897.327 \times \text{NumBedroom}$$
 
@@
 
-$$\text{Price} = 212,976.031 + 3,866,485.706 \times \text{Year} + 1,601,816.793 \times \text{Year}^2$$
+$$\hat{\text{Price}}  = 212,976.031 + 3,866,485.706 \times \text{Year} + 1,601,816.793 \times \text{Year}^2$$
 $$+ 6,591,537.312 \times \text{NumBedroom} + 1,365,656.076 \times \text{NumBedroom}^2$$
@@
 
+The modeling efforts in this paper underscore the complexity of the housing market's response to economic forces and suggest that straightforward economic models might not always suffice in capturing the real-world dynamics of real estate pricing.
 The escalation in prices may coincide with inflationary trends affecting various economic sectors, suggesting that housing markets are susceptible to the impacts of global monetary policies and financial climates.
@@
 
-The increase in demand for larger homes, as indicated by the data, suggests a potential shift in housing preferences, possibly spurred by the recent pandemic, which emphasized the need for multifunctional living spaces that combine comfort with functionality.
+The increase in demand for larger homes, as indicated in previous sections, suggests a potential shift in housing preferences, possibly spurred by the recent pandemic, which emphasized the need for multifunctional living spaces that combine comfort with functionality.
 This shift reflects not only a preference for larger living spaces but also an adaptation to new realities—homes that can accommodate remote work, virtual schooling, and indoor recreation.
@@
 
-The insights derived from the geographical distribution of housing markets highlight opportunities for strategic infrastructural investments.
-Identifying areas with rapid price growth allows stakeholders to pinpoint locations primed for development or in need of policy adjustments to stabilize the market.
+The insights derived from the geographical distribution of housing markets underscore opportunities for strategic infrastructural investments.
+The correct utilization of the Generalized Additive Model (GAM) was instrumental in identifying subtle patterns, such as the impact of microeconomic factors like local zoning laws or the availability of amenities on housing prices.
+Identifying areas with rapid price growth enables stakeholders to pinpoint locations primed for development or in need of policy adjustments to stabilize the market.
 Additionally, this data facilitates the targeted allocation of resources to underserved regions, promoting a more balanced economic distribution and enhancing overall economic equity.

### 2024-04-17 outputs/paper/paper.qmd
paper final draft

@@
 This metric confirms that the GAM has improved the predictive capability of the model, providing an accuracy that is comparable, but not necessarily superior, to the polynomial approach.
-This performance might reflect the complexities and variabilities inherent in the housing market data, which can challenge even sophisticated models like the GAM.
+This performance might reflect the complexities and variabilities inherent in the housing market data, which can challenge even complex models like the GAM.
 
@@
 The choice between these models should be informed by the balance between the need for simplicity and the desire to capture complex data relationships.
-While the Polynomial Regression strikes a balance, offering a good fit without overly complicating the model, the GAM provides the most flexibility, which could be crucial for analyzing more intricate datasets or where the precise nature of relationships between variables is less understood.
+While the Polynomial Regression strikes a balance, offering a good fit without overly complicating the model, the GAM provides the most flexibility, which could be important for analyzing more intricate datasets or where the precise nature of relationships between variables is less understood.
 
@@
 However, it also calls for a careful evaluation of the environmental impacts, as larger homes generally require more resources and energy, posing challenges to sustainability goals [@citeUKPollution].
-This balance between accommodating consumer preferences and maintaining sustainability will be crucial for future policy and economic strategies in the housing sector.
+This balance between accommodating consumer preferences and maintaining sustainability will be important  for future policy and economic strategies in the housing sector.
 
@@
 While this study is detailed, there is room for further enhancement and expansion.
-Future research should incorporate additional economic variables, such as interest rates and housing supply data, to offer a more comprehensive understanding of the market forces involved.
+Future research should incorporate additional economic variables, such as interest rates and housing supply data, to offer a better understanding of the market forces involved.
 Longitudinal studies that examine the effects of policy changes over time could also enrich our understanding of market dynamics.
 Advances in data collection technology could enable real-time analysis of market trends, allowing for a more responsive approach to rapid changes in the market.
-Additionally, broadening the scope to include rental markets and affordable housing would provide insights into the entire housing spectrum, crucial for developing holistic housing strategies that benefit all societal segments.
+Additionally, broadening the scope to include rental markets and affordable housing would provide insights into the entire housing spectrum, important for developing holistic housing strategies that benefit all societal segments.
 

### 2024-04-17 scripts/02-data_merging.R
script updated

@@ -17,8 +17,6 @@ library(lubridate)
 
 
 #### load data ####
-setwd("/Users/ijustin/Library/CloudStorage/Dropbox/Code/STA302/Research_Papers/House_Price")
-
 state_1b_house_price_by_year <- read_csv("outputs/data/state_1b_house_price_by_year.csv")
 state_2b_house_price_by_year <- read_csv("outputs/data/state_2b_house_price_by_year.csv")
 state_3b_house_price_by_year <- read_csv("outputs/data/state_3b_house_price_by_year.csv")

### 2024-04-17 scripts/05-tests.R

@@ -9,6 +9,7 @@
 
 
 #### Workspace setup ####
+library(arrow)
 library(readr)
 library(testthat)
 library(tidyverse)
@@ -17,7 +18,7 @@ library(tidyverse)
 
 #### merged_house_price testing ####
 test_that("merged_house_price testing", {
-  merged_house_price<- readr::read_csv("../../House_Price/outputs/data/merged_house_price.csv")
+  merged_house_price <- read_parquet("../../House_Price/outputs/data/merged_house_price.parquet")
   
   # Test if 'Year' values are positive
   expect_true(all(merged_house_price$Year > 0),
